---
title: "bin code"
author: "Vera Weisbecker"
date: "`r Sys.Date()`"
output: html_document
---


### Prepare evolutionary allometry information (verbatim from AmNat paper) for Fig 1
```{r}
# Run Procrustes ANOVA 
pgls.gdf <- geomorph.data.frame(coords = mean.shapes, size = info.means$MeanCsize, clade = info.means$Clade)

#pgls as per previous paper
fit_pgls_allo_evo <- procD.pgls(coords ~ log(size)*clade , aus.tree, iter = 500, data = pgls.gdf, print.progress = FALSE)

#plain gls
fit_lm_allo <- procD.lm(coords ~ log(size) , iter = 500, data = pgls.gdf, print.progress = FALSE)


# Find Reg Score values for plotting purposes
evo_allo_rs <- plotAllometry(fit_pgls_allo_evo, size = info.means$MeanCsize, method = "RegScore")

# Also find Reg Score values for lm for plotting purposes
allo_rs <- plotAllometry(fit_lm_allo, size = info.means$MeanCsize, method = "RegScore")

```

### Export Figure 1 to pdf
Finally, we compare the full shape PCA to the shape residuals PCAs. All PCAs are relevant to Figure 2 heatmaps.
```{r}
# Set up Notomys abline
not.index <- which(info.means$Genus.y == "Notomys")
not.Reg <- evo_allo_rs$RegScore[not.index]
not.Cs <- info.means$MeanCsize[not.index]

# Set up for export and for multiple panels
part.coords <- c(-0.117, 0.095)  # plotting parameters for A, C, D, E placement

#setEPS()
#postscript("../Data/Results/Figure1_PCAs_6plot.eps")
mat <- matrix(c(1:6), 3, 2, byrow = TRUE)  # 6 plots, 3 rows, 2 columns
layout(mat, widths = rep(1, dim(mat)[2]), heights = rep(0.5, dim(mat)[1]))   

# A) FULL SHAPE PCA - diet & locomotion legend
par(mar = c(4.5, 4.5, 2, 1))
plot(x = -mean.PCA$x[, 1],  # Flip for consistency
                 y = mean.PCA$x[, 2],  
                 xlab = "PC1 (52.3%)", 
                 ylab = "PC2 (14.5%)",
                 xlim = c(-0.12, 0.12),
                 ylim = c(-0.1, 0.1),
                 col = col.diet.means, 
                 pch = pch.gld.means, 
                 bg = col.diet.means,
                 cex = 1.5,
                 cex.axis = 1.1, 
                 cex.lab = 1.3)
text(part.coords[1], part.coords[2], "a", cex = 1.5)
text(-0.045, 0.092, "full shape dataset (ecology)")
text(-0.078, 0.075, expression(paste("Marcy ", italic("et al. "), "2020")), cex = 0.8)

legend(-0.131, 0.067, legend = c("Arboreal", "Hopping", "Terrestrial", "Semiaquatic", "", "Carnivore", "Frugivore", "Granivore", "Folivore", "Omnivore"), col = c(rep("black", 4), "white", col.diet.5), border = NULL, pch = c(24:21, rep(16, 6)), cex = 0.8, bty = "n")

# B) Evolutionary Allometry plot - diet & locomotion legend
par(mar = c(4, 4, 1, 1))
plot(x = log(info.means$MeanCsize),
     y = -evo.allo.rs$RegScore,
     xlim = c(4.75, 6),
     col = col.diet.means, 
     pch = pch.gld.means, 
     bg = col.diet.means,
     main = " ",
     xlab = "Log centroid size", 
     ylab = "Shape  (Reg Score)",
     bty = "n")
text(4.8, 0.006, "b", cex = 1.5)
text(5.1, 0.0058, "evolutionary allometry")
abline(lm(evo_allo_rs$RegScore ~ log(info.means$MeanCsize)), col = "dark grey", lwd = 1)  # common evolutionary allometry line
#text(x = log(info.means$MeanCsize), y = -evo.allo.rs$RegScore, labels = info.means$Genus)  # labels genera on plot points

# C) FULL SHAPE PCA - genus colors
par(mar = c(4.5, 4.5, 2, 1))
plot(x = -mean.PCA$x[, 1],  # flip for consistency
                 y = mean.PCA$x[, 2],  
                 xlab = "PC1 (52.3%)", 
                 ylab = "PC2 (14.5%)",
                 xlim = c(-0.12, 0.12),
                 ylim = c(-0.1, 0.1),
                 col = alpha.col,
                 pch = 16,
                 bg = alpha.col,
                 cex = 1.5,
                 cex.axis = 1.1, 
                 cex.lab = 1.3)
text(part.coords[1], part.coords[2], "c", cex = 1.5)
text(-0.046, 0.092, "full shape dataset (genera)")

legend("bottomleft", legend = genera.phylo.full, col = col.phylo, pch = 16, cex = 0.68, pt.cex = 0.85, bty = "n", text.font = 3)

# D) shape residuals PCA - genus colors
par(mar = c(4.5, 4.5, 2, 1))
plot(x = res.PCA$x[, 1],  # no flip for consistency
                 y = -res.PCA$x[, 2],  
                 xlab = "PC1 (26.8%)", 
                 ylab = "PC2 (18.6%)",
                 xlim = c(-0.12, 0.12),
                 ylim = c(-0.1, 0.1),
                 col = alpha.col,
                 pch = 16,
                 bg = alpha.col,
                 cex = 1.5,
                 cex.axis = 1.1, 
                 cex.lab = 1.3)
text(part.coords[1], part.coords[2], "d", cex = 1.5)
text(-0.035, 0.092, "shape residual dataset (genera)")

# E) No Notomys shape residuals PCA - genus colors
par(mar = c(4.5, 4.5, 2, 1))
plot(x = wrap.PCA.r.n$x[, 1],  
                 y = wrap.PCA.r.n$x[, 2],  # no flip for consistency
                 xlab = "PC1 (30.1%)", 
                 ylab = "PC1 (16.7%)",
                 xlim = c(-0.12, 0.12),
                 ylim = c(-0.1, 0.1),
                 col = alpha.col.no, 
                 pch = 16, 
                 bg = alpha.col.no,
                 cex = 1.5,
                 cex.axis = 1.1, 
                 cex.lab = 1.3)
text(part.coords[1], part.coords[2], "e", cex = 1.5)
text(-0.034, 0.092, expression(paste("shape residual without ", italic("Notomys"))))

# F) shape Residuals PCA - ecology
par(mar = c(4.5, 4.5, 2, 1))
plot(x = res.PCA$x[, 1],  # no flip for consistency
                 y = -res.PCA$x[, 2],  
                 xlab = "PC1 (26.8%)", 
                 ylab = "PC1 (18.6%)",
                 xlim = c(-0.12, 0.12),
                 ylim = c(-0.1, 0.1),
                 col = col.diet.means, 
                 pch = pch.gld.means, 
                 bg = col.diet.means,
                 cex = 1.5,
                 cex.axis = 1.1, 
                 cex.lab = 1.3)
text(part.coords[1], part.coords[2], "f", cex = 1.5)
text(-0.053, 0.092, "shape residual (ecology)")

#dev.off()
```

#two-panel additional supplementary figure to compare allometry with and without phylogenetic correction

```{r}
setEPS()
postscript("../Data/Results/Supp_X_allometries.eps")

par(mfrow=c(2,1))

#evolutionary
par(mar = c(4, 4, 1, 1))
plot(x = log(info.means$MeanCsize),
     y = -evo.allo.rs$RegScore,
     xlim = c(4.75, 6),
     col = col.diet.means, 
     pch = pch.gld.means, 
     bg = col.diet.means,
     main = " ",
     xlab = "Log centroid size", 
     ylab = "Shape  (Reg Score)",
     bty = "n")
text(4.8, 0.001, "A", cex = 1.5)
text(5.05, 0.001, "evolutionary allometry")
abline(lm(-evo_allo_rs$RegScore ~ log(info.means$MeanCsize)), col = "dark grey", lwd = 1)  # common evolutionary allometry line


#lm
par(mar = c(4, 4, 1, 1))
plot(x = log(info.means$MeanCsize),
     y = allo_rs$RegScore,
     xlim = c(4.75, 6),
     col = col.diet.means, 
     pch = pch.gld.means, 
     bg = col.diet.means,
     main = " ",
     xlab = "Log centroid size", 
     ylab = "Shape  (Reg Score)",
     bty = "n")
text(4.8, 0.1, "B", cex = 1.5)
text(5.2, 0.1, "allometry in non-phylogenetic linear model")
abline(lm(allo_rs$RegScore ~ log(info.means$MeanCsize)), col = "dark grey", lwd = 1)  # common evolutionary allometry line

dev.off()


```



## 2) Compare the full shape and shape residual datasets
Mantel tests to compare full shape morphospace to the shape residual morphospace
```{r}
#A quick check of the correlation betwen full shape PC2 and residual shape PC1:
cor(mean.PCA$x[,2], res.PCA$x[,1])


# Make distance matrices
d.full.pca <- dist(mean.PCA$x)
d.res.pca <- dist(res.PCA$x)

# Mantel test for all PC scores
man.all <- mantel(d.full.pca, d.res.pca, method = "pearson", permutations = 1000) 
man.all$statistic
man.all$signif
```

### Mantel tests to compare just the 3 most important PCs
```{r}
# Make distance matrices
d.full.pc <- dist(mean.PCA$x[, 1:3])
d.res.pc <- dist(res.PCA$x[, 1:3])

# Mantel test for all PC scores
man.pc13 <- mantel(d.full.pc, d.res.pc, method = "pearson", permutations = 1000) 
man.pc13$statistic
man.pc13$signif
```

### Comparison of morphospaces without allometric PC1
```{r}
# Make distance matrices
d.full.wo1 <- dist(mean.PCA$x[, 2:dim(mean.PCA$x)[2]])
d.res.woL <- dist(res.PCA$x[, 1:dim(res.PCA$x)[2]-1])

# Mantel test for all PC scores
man.pc.wo1 <- mantel(d.full.wo1, d.res.woL, method = "pearson", permutations = 1000) 
man.pc.wo1$statistic
man.pc.wo1$signif
```


### Correlation of residual PC1 to full shape PC2 and vice versa
```{r}
cor.pc1 <- cor(mean.PCA$x[, 2], -res.PCA$x[, 1])
cor.pc1

cor.pc2 <- cor(mean.PCA$x[, 1], res.PCA$x[, 2])
cor.pc2

cor.pc2v3 <- cor(mean.PCA$x[, 3], res.PCA$x[, 2])
cor.pc2v3
```


## 3) Export skreeplots for PC importance scores for 3 separate datasets
Set up 3 datasets for screeplots
```{r}
# Full shape
full.PC.impt <- data.frame(PC = 1:(length(summary(mean.PCA)$PC.summary)), impt = unlist(summary(mean.PCA)[[1]][2,]))

# Shape residual
res.PC.impt <- data.frame(PC = 1:(length(summary(res.PCA)$PC.summary)), impt = unlist(summary(res.PCA)[[1]][2,]))

# Shape residual without Notomys wrap.PCA.r.n
rnot.PC.impt <- data.frame(PC = 1:(length(summary(wrap.PCA.r.n)$PC.summary)), impt = unlist(summary(wrap.PCA.r.n)[[1]][2,]))

```

### Create supplementary figure 1
```{r}
# Make simple multipanel figure
pdf("../Data/Results/SFigure1_Screeplots.pdf")
mat <- matrix(c(1:3), 3, 1, byrow = TRUE)  # 3 plots, 3 rows, 1 column
layout(mat, widths = rep(0.5, dim(mat)[2]), heights = rep(0.25, dim(mat)[1])) 

# 1) 
barplot(full.PC.impt[, 2], ylim = c(0, 0.5))
abline(h = 0.1, col = 'deeppink4')
mtext("A) full shape dataset", 3, line = 1.5, cex = 0.9, adj = 0, col = "black")
mtext("% of variance", 2, line = 2.5, cex = 0.83, col = "black")

# 2)
barplot(res.PC.impt[, 2], ylim = c(0, 0.5))
abline(h = 0.1, col = 'deeppink4')
mtext("B) shape residual dataset", 3, line = 1, cex = 0.9, adj = 0, col = "black")
mtext("% of variance", 2, line = 2.5, cex = 0.83, col = "black")

# 3)
barplot(rnot.PC.impt[, 2], ylim = c(0, 0.5))
abline(h = 0.1, col = 'deeppink4')
mtext(expression(paste("C) shape residual dataset without ", italic(Notomys))), 3, line = 1, cex = 0.9, adj = 0, col = "black")
mtext("% of variance", 2, line = 2.5, cex = 0.83, col = "black")
mtext("Principal component", 1, line = 1, cex = 0.83, col = "black")

dev.off()
```

