---
title: "02-calculate-user-error"
author: "Ariel Marcy"
date: "2018-09-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = '../eco-rodents')
```

# Calculate user error
Geometric morphometrics usually requires a lot of user-generated data and it's important to assess how much user error is created during landmarking. 

First, we visualize whether any outliers can be explained by mistakes in landmark placement (e.g. a point "falling through" the nasal to land on the inner basicranium). *This was already done for this dataset in allometry-rodents repo and associated paper.*

Second, we compare the shape variation between 20 individuals, from a variety of different species, landmarked twice to calculate the intra-observer (i.e. digitizer) error.

Third, we'll finalize the specimens to be included in future analyses.

### Load packages, functions, and data from previous steps
```{r message = FALSE}
library(stringr)    # CRAN v1.5.0
library(stringi)    # CRAN v1.7.12
library(data.table) # CRAN v1.14.8
library(geomorph)   # CRAN v4.0.6

# This sets the wd to local
library(rstudioapi) # Safely Access the RStudio API
setwd(dirname(getActiveDocumentContext()$path))

source("../Functions/utilities.R")  # custom functions
load(file = "../Data/Processed/01-bilat-data-w-ERR.rda")
```

## 1) Check for outliers and fix landmarking errors
Done in previous manuscript (see [allometry-rodents repo](https://github.com/miracleray/allometry-rodents))

### Remove outliers from dataset
Three specimens had major landmarking errors -- landmarks consistently placed in the wrong places. Therefore, we opted to remove these specimens.
```{r}
# Find indices (same for both big and small datasets)
rat.out <- which(sm.sp.info$CatNum == "C11381")  # Rattus norvegicus
uro.out <- which(sm.sp.info$CatNum == "J17610")  # Uromys caudimaculatus

# Remove from both big and small datasets
sm.sp.info <- sm.sp.info[-c(rat.out, uro.out), ]  # metadata
big.sp.info <- big.sp.info[-c(rat.out, uro.out), ]
sm.bY.shape <- sm.bY.shape[, , -c(rat.out, uro.out)]  # shape data
big.bY.shape <- big.bY.shape[, , -c(rat.out, uro.out)]

#now for the modules

for (i in 1:length(module_list)){
  
  module_bilat[[i]]<-module_bilat[[i]]$coor [-c(rat.out, uro.out), ]
  
}

```

## 2) Calculate intra-observer error
Now that the shape data have been processed via GPA + bilateral symmetry and corrected for obvious landmarking errors, we can separate the data intended to measure user error from the main data for the remaining analyses. 

### Make the datasets to be used for the rest of the analyses.
```{r}
main <- which(big.sp.info$Rep == "non")  # this vector of non-replicate indices is the same for big and small patch protocols, tested with all.equal()

big.shape <- big.bY.shape[, , main]  # big patch
big.info <- big.sp.info[main, ]

sm.shape <- sm.bY.shape[, , main]  # small patch
sm.info <- sm.sp.info[main, ]
```

### Create error measurement datasets for each protocol 
Includes replicates and the specimens they duplicated (i.e. each CatNum should have two entries, one labeled "ERR", and one labled "non").
```{r}
# Find index for both versions of specimens with replicates using rep.spec list made in script 01
rep.CatNums <- as.character(str_sub(rep.spec, 9, -1))  # isolate CatNum
error.data <- which(sm.sp.info$CatNum %in% rep.CatNums)  # error.data is the same for small and big patch protocols, tested with all.equal()

# Subset shape and metadata tables for each protcol
big.err.shape <- big.bY.shape[, , error.data]
big.err.info <- big.sp.info[error.data, ]

sm.err.shape <- sm.bY.shape[, , error.data]
sm.err.info <- sm.sp.info[error.data, ]
```

### Calculate user error aka Repeatability
Calculated for each patch protocol from formulas 1-3 in Fruciano 2016. Custom function, `RepAbility()` was modified from Marcy et al. 2018.
```{r}
# n.rep = 2 in this case because we measured each specimen twice
RepAbility(big.err.shape, big.err.info$CatNum, n.Rep = 2)  
RepAbility(sm.err.shape, sm.err.info$CatNum, n.Rep = 2)
```

### Save intermediate data 
Main datasets for both protocols without replicates.
```{r}
save(sm.shape, sm.info, sm.pt.names, file = "../Data/Processed/02-no-error-data-sm.rda")
save(big.shape, big.info, big.pt.names, file = "../Data/Processed/02-no-error-data-big.rda")
```
