---
title: "02-calculate-user-error"
author: "Ariel Marcy"
date: "2018-09-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '../eco-rodents')
```
# Calculate user error
Geometric morphometric studies usually require a lot of user-generated data and it's important to assess how much user error is created during the landmarking process. 

First, we visualize whether any outliers can be explained by mistakes in landmark placement (e.g. a point "falling through" the nasal to land on the inner basicranium). This was already done for this dataset in allometry-rodents repo and associated paper.

Second, we compare the shape variation between 20 individuals, from a variety of different species, landmarked twice to calculate the intra-observer (i.e. digitizer) error.

Third, we'll finalize the specimens to be included in future analyses.

### Load packages, functions, and data from previous steps
```{r message = FALSE}
library(stringr)
library(stringi)
library(data.table)
library(geomorph)
source("../Functions/utilities.R")  # custom functions defined in utilities.R
load(file = "../Data/Processed/01-bilat-data-w-ERR.rda")
```

## 1) Check for outliers and fix landmarking errors
Done in previous manuscript (see allometry-rodents repo)

### Remove outliers from dataset
Three specimens had major landmarking errors -- landmarks consistently placed in the wrong places. Therefore, we opted to remove these specimens.
```{r}
# Find indices (same for both big and small datasets)
mel.out <- which(sm.sp.info$CatNum == "J22537")  # Melomys burtoni
rat.out <- which(sm.sp.info$CatNum == "C11381")  # Rattus norvegicus
uro.out <- which(sm.sp.info$CatNum == "J17610")  # Uromys caudimaculatus

# Remove from both big and small datasets
sm.sp.info <- sm.sp.info[-c(mel.out, rat.out, uro.out), ]
big.sp.info <- big.sp.info[-c(mel.out, rat.out, uro.out), ]
sm.bY.shape <- sm.bY.shape[, , -c(mel.out, rat.out, uro.out)]
big.bY.shape <- big.bY.shape[, , -c(mel.out, rat.out, uro.out)]
```

## 2) Calculate intra-observer error
Now that the shape data have been processed via GPA + bilateral symmetry and corrected for obvious landmarking errors, we can separate the data intended to measure user error from the main data for the remaining analyses. 

Make the datasets to be used for the rest of the analyses.
```{r}
main <- which(big.sp.info$Rep == "non")
big.shape <- big.bY.shape[, , main]
big.info <- big.sp.info[main, ]

sm.shape <- sm.bY.shape[, , main]  # main is the same for big and small patch protocols, tested with all.equal()
sm.info <- sm.sp.info[main, ]
```

Create an error measurement dataset for both protocols that includes replicates and the specimens they duplicated (i.e. each CatNum should have two entries, one labeled "ERR", and one labled "non"). These data will be used to calculate user error. 
```{r}
# Find index for both versions of specimens with replicates using rep.spec list make in script 01.
rep.CatNums <- as.character(str_sub(rep.spec, 9, -1))  # isolate CatNum
error.data <- which(sm.sp.info$CatNum %in% rep.CatNums)  # error.data is the same for small and big patch protocols, tested with all.equal()

# Subset shape and metadata tables for each protcol
big.err.shape <- big.bY.shape[, , error.data]
big.err.info <- big.sp.info[error.data, ]

sm.err.shape <- sm.bY.shape[, , error.data]
sm.err.info <- sm.sp.info[error.data, ]
```

### Calculate user error aka Repeatability
Here, we calculate repeatability for each patch protocol from formulas 1-3 in Fruciano 2016. The function `RepAbility()` was modified from Marcy et al. 2018.
```{r}
# Note: n.rep = 2 in this case because we measured each specimen twice
RepAbility(big.err.shape, big.err.info$CatNum, n.Rep = 2)  
RepAbility(sm.err.shape, sm.err.info$CatNum, n.Rep = 2)
```

### Calculate repeatability without patch points
Patch points could artificially inflate our repeatability numbers since they are placed automatically by the Viewbox software. To test this, we calculated repeatability for each protocol with patch points removed. 

Note that removing patch points will remove 16 more points from the big patch protocol than from the small patch protocol. 

<Done in previous manuscript, see allometry-rodents repo>

Taking out patch points decreases the repeatability by less than 2% for each protocol:
Big Patch is 93.0% with patches, 91.3% without patches (140 points removed)
Small Patch: 93.0% with patches, 91.5% without patches (124 points removed) 

### Save intermediate data 
These are the main datasets for both protocols without replicates
```{r}
save(big.shape, big.info, sm.shape, sm.info, file = "../Data/Processed/02-main-data-both-protocols.rda")
```