---
title: "04-plot-exploratory-PCAs"
author: "Ariel Marcy"
date: "2018-10-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '../eco-rodents')
```

# Plotting exploratory PCAs
PCAs provide a human-friendly way to visualize shape data. They are a good way to understand what's going on in the data quickly. However, because they artificially collapse the shape variation into orthogonal PC axes, they must be considered exploratory analyses only.

### Load packages, functions, and data
```{r message = FALSE}
library(stringr)
library(data.table)
library(geomorph)
library(colorspace)  # colors
source("../Functions/utilities.R")  # custom functions
load(file = "../Data/Processed/03-sm-PCA-with-color.rda")
```

### Exploratory analyses with PCAs
Here, we explore the variation in rodent skulls by coloring the PCA plots by genus, by traits, by biome region.

Since we are going to add several columns to the metadata table, we'll remove columns no longer needed past the bilateral symmetry and replicate handling stages. 
```{r}
# Prep metadata for column additions
info <- info[, c(-4, -5, -7)]  # removes now-unnecessary columns
info$Order <- c(1:dim(info)[1])  # preserves specimen order
```

Trait information was gathered from the book, Native Mice and Rats by Dr.s Bill Breed and Fred Ford (2007). 
```{r}
# Add trait data to the metadata table
traits <- read.csv("../Data/Processed/in_ex_traits.csv", header = TRUE)
traits$Taxa <- paste(str_sub(traits$Genus, 1, 1), str_sub(traits$Species, 1, 3), sep = "_")  # make matching Taxa column
info.traits <- merge(info, traits, by = "Taxa", sort = F)  # merge
info.traits <- info.traits[order(info.traits$Order), ]  # sort by Order

## Rename Genus and Species columns 
# short names (e.g. Pse or mus) have original column names
# long names (e.g. Pseudomys or musculus) have a ".y" appended
names(info.traits)[names(info.traits) == 'Genus.x'] <- 'Genus'
names(info.traits)[names(info.traits) == 'Species.x'] <- 'Species'

# Make effective taxa column to assign Mastacomys to genus Pseudomys
info.traits$ETaxa <- info.traits$Taxa
info.traits$ETaxa[which(info.traits$Taxa == "M_fus")] <- "P_fus"
```

## Plot PCA by genus for small patch dataset
The function `PlotPCA()` quickly plots a PCA with specimens colored by groups and optionally given different point shapes for other attributes. The function also reports PC axis variation in %. 
```{r}
# Plot PCA by genus and export as .eps into ../Data/Results folder
### setEPS()
### postscript("../Data/Results/PCA_genus.eps")
PlotPCA(PCA, 1, 2, sp.col.gen)
legend(0.144, 0.056, legend = genera.phylo, col = col.phylo, pch = 16, cex = 0.78)
### dev.off()
```

## PCA plot by diet and guild
This PCA colors specimens by diet and assigns point shape based on guild, i.e. where they spend most of their time.
```{r}
# Set up pallete for diet
warm <- sequential_hcl(5, "LaJolla")  
col.carn <- warm[4]  # meat colored red
col.frug <- warm[3]  # fruit colored orange
col.gran <- warm[2]  # grain colored gold
col.herb <- sequential_hcl(5, "Greens 3")[2]  # green colored grass
col.omni <- sequential_hcl(5, "Grays")[3]  # gray for anything goes

# Define color pallete for diets of all specimens
col.diet <- c(col.carn, col.frug, col.gran, col.herb, col.omni)
col.diet <- PlotByGroup(info.traits, "Feeding", col.diet)

# Points: Aquatic (circle), Arboreal (triangle), Terrestrial (square)
pch.gld <- c(19, 24, 22)
pch.gld <- PlotByGroup(info.traits, "Guild", pch.gld)
```

Plot PC1 vs PC2 and add custom legends.
```{r}
PlotPCA(PCA, 1, 2, col.diet, pch.gld)
legend("bottomright", legend = c("Aquatic", "Arboreal", "Terrest."), col = "black", border = NULL, pch = c(21, 24, 22))
text(-0.01, -0.09, "Carnivore", col = col.carn)
text(0.11, -0.02, "Frugivore", col = col.frug)
text(0.08, 0.045, "Herbivore", col = col.herb)
text(-0.07, -0.06, "Omnivore", col = col.omni)
text(-0.117, 0.01, "Granivore", col = col.gran)
```

## Morphospace partitioning by habitat
So far, most of the variation in the morphospace could be explained by size and genus affiliation, with different genera of similar sizes overlapping quite a bit. This looks like there isn't much differentiation among Australian rodents, but perhaps the differentiation is hidden by geographical region or partitioning of the environment. 

### Write functions to subset and visualize species morphospace by region
First we will subset the PCA scores by species which inhabit the 7 different ecological regions of Australia. Regions and species distributions therein taken from Breed & Ford 2007. `FoundInRegion()` subsets existing PCA data by species present in given region(s).

Next, we wrote a function, `PointOutDiffSpp()` to give plotting points (pch values) by unique species within each genus.

Finally, we'll combine these functions with a loop to plot PCAs with color by taxa for each region, unique species within each genera having a unique point shape, and a legend for only the species present in each region. 
```{r}
region.names <- c("Savannah", "Arid Zone", "NE Wet Forest", "NE Dry Forest", "Pilbara", "South West", "South East")
pc.scores <- subset(PCA$pc.scores, select = 1:3)  # only 1st 3 PCs
col.egen <- col.gen[-4]  # take out Mas color

# Loop through all 7 regions to plot morphospace of species found in each
for (i in 1:7) {
        # Subset info.traits, PC scores, and color key by region
        in.region <- FoundInRegion(info.traits, regions = i)
        info.region <- info.traits[in.region, ]
        pc.region <- as.data.frame(pc.scores[in.region, ])
        col.taxa.region <- sp.col.gen[in.region]
        
        # Assign points by unique species in a genus
        pch.tax.legend <- PointOutDiffSpp(info.region)
        pch.tax.region <- PlotByGroup(info.region, "ETaxa", pch.tax.legend)
        
        # Plot with fancy (i.e. verbose) legend by:
        info.region$Legend <- paste(str_sub(info.region$EGenus, 1, 3), str_sub(info.region$Species, 1, 3))  # Mus musculus -> "Mus mus"

        # 1) Making list of unique binomial names
        unique.taxa <- sort(unique(info.region$Legend))  # alphabetical order
        legend.genus <- substr(unique.taxa, 1, 3)  # makes string of genera names with repeats when there are unique species in a genus
        
        # 2) Making color key for legend
        genera.in.region <- which(levels(info.traits$EGenus) %in% sort(unique(info.region$EGenus)))
        col.gen.reg <- col.egen[genera.in.region]  # subset col.egen for region
        names(col.gen.reg) <- unique(legend.genus)  # make genus color key with names of unique genera
        col.gen.legend <- col.gen.reg[match(legend.genus, names(col.gen.reg))]
        
        # Write x and y labels with proportion of variance for PCx and PCy
        PCs <- PCA$pc.summary$importance
        PCx.per <- round(PCs[2, 1] * 100, digits = 1)  # % with 1 decimal
        PCx.lab <- paste("PC1 (", PCx.per, "%)", sep = "")
        PCy.per <- round(PCs[2, 2] * 100, digits = 1)
        PCy.lab <- paste("PC2 (", PCy.per, "%)", sep = "")
        
        # Plot PCA with legend and labels
        PCA.plot <- plot(x = pc.region$PC1,
                         y = pc.region$PC2,
                         main = paste(region.names[i], "Region Morphospace" ),
                         xlab = PCx.lab,
                         ylab = PCy.lab,
                         xlim = c(-0.11, 0.18),
                         ylim = c(-0.095, 0.05),
                         col = col.taxa.region, 
                         pch = pch.tax.region, 
                         bg = col.taxa.region,
                         cex = 1.5,
                         cex.axis = 1.3, 
                         cex.lab = 1.3)
        legend(0.147, 0.058, legend = unique.taxa, col = col.gen.legend, border = NULL, pch = pch.tax.legend, pt.bg = col.gen.legend, cex = 0.64)
}
```

### Save intermediate data
```{r}
save(info.traits, col.diet, pch.gld, col.egen, file = "../Data/Processed/04-sm-data-with-traits.rda")
```