---
title: "04-plot-allometry-PCAs"
author: "Ariel Marcy"
date: "2018-10-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '../eco-rodents')
```

# Plotting exploratory PCAs
PCAs provide a human-friendly way to visualize shape data. They are a good way to understand what's going on in the data quickly. However, because they artificially collapse the shape variation into orthogonal PC axes, they must be considered exploratory analyses only.

Here, we explore the variation in rodent skulls by coloring the PCA plots by genus and by ecologically-relevant traits.

### Load packages, functions, and data
```{r message = FALSE}
library(stringr)
library(data.table)
library(geomorph)
library(geiger)
library(colorspace)  # colors
source("../Functions/utilities.R")  # custom functions
load(file = "../Data/Processed/03-main-data.rda")
load(file = "../Data/Processed/03-graphing-vectors.rda")
```

## 1) Plot PCA by genus for small patch dataset
The custom function `PlotPCA()` quickly plots a PCA with specimens colored by groups and optionally given different point shapes for other attributes. The function also reports PC axis variation in %. 
```{r}
PCA <- plotTangentSpace(shape)

# Plot PCA with all data by genus
PlotPCA(PCA, 1, 2, sp.col.gen, pch.spp)
legend("bottomright", legend = phylo.names, col = col.unique.spp, pch = pch.unique.spp, pt.bg = col.unique.spp, cex = .35, pt.cex = .7, ncol = 2)
```

## 2) PCA plot by ancestor group
There are two main tribes on Australia: Hydromini and Rattini, which represent two different lineages within Murinae to make it from Asia to Sahul. 

Within Hydromini, there are 5 groups with unique ancestors and arrivals from New Guinea to Australia. Most had unique radiations that centered on different biomes. 
```{r}
# Set up colors, using distinct colors from above. 
col.Hyd <- col.unique.spp[which(phylo.names == "H chr")]  # Hydromys 
col.invR <- "black"  # to distinguish between native and invasive Rattus rats 
col.Mes <- col.unique.spp[which(phylo.names == "M gou")]  # Mesembriomys
col.mus <- col.unique.spp[which(phylo.names == "M mus")]  # Mus musculus
col.Pog <- col.unique.spp[which(phylo.names == "P mol")]  # Pogonomys
col.Pse <- col.unique.spp[which(phylo.names == "L for")]  # Pseudomys
col.Rat <- col.unique.spp[which(phylo.names == "R vil")]  # native Rattus
col.Uro <- col.unique.spp[which(phylo.names == "M bur")]  # Uromys

col.clade <- c(col.Hyd, col.invR, col.Mes, col.mus, col.Pog, col.Pse, col.Rat, col.Uro)

# Colors matching each specimen in order of shape dataset
col.by.clade <- PlotByGroup(info, "Clade", col.clade)

# PCA 
PlotPCA(PCA, 1, 2, col.by.clade)
```

## 3) Check for disparity differences between old and new endemics
Check to see if differences seen in PCA (which squashes multi-variate variation) occur when all axes of variation are taken into account. Here we are testing the difference between the space occupied by the old endemics versus the new endemics
```{r}
# Remove invasives
end.shape <- shape[, , -(which(info$Endemic == "Inv"))]
end.info <- info[-(which(info$Endemic == "Inv")), ]

# Force shape into class "gpagen"
gpagen.shape <- list()
gpagen.shape$coords <- end.shape
gpagen.shape$endemic <- end.info
class(gpagen.shape) <- "gpagen"

# Make into geomorph data frame - need to use "coords" and "factors" 
traitspace.gdf <- geomorph.data.frame(coords = gpagen.shape$coords, factors = gpagen.shape$endemic)

# Ordinate
traitspace <- geomorph.ordination(traitspace.gdf)

# This computes the distance of both groups from the old endemics' center; summary gives the stats, plot gives the boxplot
dist.from.old <- dispRity(traitspace, metric = pairwise.dist)

summary(dist.from.old)
plot(dist.from.old, ylim = c(0.01, 0.26))

# This t-test shows that the new species are significantly further away from the old centroid than the old species
disp.t.test <- test.dispRity(dist.from.old, test = t.test, alternative = "less")
```

## 4) PCA plot by diet and guild
This PCA colors specimens by diet and assigns point shape based on guild, i.e. where they spend most of their time.
```{r}
# Set up pallete for 5 diet types
warm <- sequential_hcl(5, "LaJolla")  
col.carn <- warm[4]  # meat colored red
col.frug <- warm[3]  # fruit colored orange
col.gran <- warm[2]  # grain colored gold
col.herb <- sequential_hcl(5, "Greens 3")[2]  # green colored grass
col.omni <- sequential_hcl(5, "Grays")[3]  # gray for anything goes

# Define color pallete for diets of all specimens
col.diet.5 <- c(col.carn, col.frug, col.gran, col.herb, col.omni)
col.diet <- PlotByGroup(info, "Feeding", col.diet.5)

# Points for 4 locomotion types: Aquatic (circle), Arboreal (triangle), Terrestrial (square), hopping (diamond)
pch.gld.4 <- c(16, 24, 23, 22)
pch.gld <- PlotByGroup(info, "Guild", pch.gld.4)
```

Plot PC1 vs PC2 and add custom legends.
```{r}
PlotPCA(PCA, 1, 2, col.diet, pch.gld)
legend("bottomright", legend = c("Aquatic", "Arboreal", "Terrest.", "Hopping"), col = "black", border = NULL, pch = c(1, 24, 22, 23))
text(0.02, -0.09, "Carnivore", col = col.carn)
text(0.11, -0.02, "Frugivore", col = col.frug)
text(0.08, 0.045, "Herbivore", col = col.herb)
text(-0.06, -0.05, "Omnivore", col = col.omni)
text(-0.11, 0.01, "Granivore", col = col.gran)
```

## 5) Extra plots for powerpoint
Created for presentation with different formatting

### Residual by genera
```{r}
plot(x = -res.PCA$pc.scores[, 1],  # flip for consistency
                 y = -res.PCA$pc.scores[, 2],  
                 xlab = PCx.lab, 
                 ylab = PCy.lab,
                 asp = F,
                 xlim = c(-0.12, 0.15),
                 ylim = c(-0.15, 0.12),
                 col = alpha.col,
                 pch = alpha.pch,
                 bg = alpha.col,
                 cex = 1.5,
                 cex.axis = 1.1, 
                 cex.lab = 1.3)
legend("bottomright", legend = genera.phylo, col = col.phylo, pch = 16, pt.bg = col.unique.spp, cex = 0.75, bty = "n", text(font = 3))
```

### Residual by locomotion and diet
```{r}
plot(x = -res.PCA$pc.scores[, 1],  # flip for consistency
                 y = -res.PCA$pc.scores[, 2],  
                 xlab = PCx.lab, 
                 ylab = PCy.lab,
                 asp = F,
                 xlim = c(-0.12, 0.15),
                 ylim = c(-0.15, 0.12),
                 col = col.diet.means, 
                 pch = pch.gld.means, 
                 bg = col.diet.means,
                 cex = 1.5,
                 cex.axis = 1.1, 
                 cex.lab = 1.3)
legend("topright", legend = c("Aquatic", "Arboreal", "Terrestrial", "Hopping", "", "Carnivore", "Frugivore", "Granivore", "Herbivore", "Omnivore"), col = c(rep("black", 4), "white", col.diet.5), border = NULL, pch = c(21, 24, 22, 23, rep(15, 6)), cex = 0.76, bty = "n", inset = c(0, 0))
```

### Residual phylomorphospace
```{r}
source(file = "../Functions/plotGMPhyloMorphoSpace_plotmod.R")
library(phytools)

plotGMPhyloMorphoSpace_plotmod(aus.tree, residuals, tip.labels = FALSE, node.labels = FALSE, xaxis = -1, yaxis = 2, plot.param = list(xlim = c(-0.15, 0.12), ylim = c(-0.12, 0.15), t.bg = col.diet.means, t.pch = pch.gld.means, l.col = "dark gray", n.cex = 1, t.cex = 1.5, xlab = "PC1 (30%)", ylab = "PC2 (21.1%)"))
```