---
title: "04-compare_PCAs"
author: "Ariel Marcy & Vera Weisbecker"
date: "8/15/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = '../eco-rodents')
```

# Plot PCAs for allometric and shape residual datasets
Plot PCAs for allometric shape and shape residual, provide coloration and pch symbols by locomotion and diet, and visualize changes in shape residual PCA when *Notomys* species are removed. These plots are relevant to the heatplots created in script 05 / shown in Figure 2. 

Creates multipanel Figure 1.

Compares morphospace similarity of the two datasets with Mantel tests and a correlation test.

Creates supplementary Figure 1, a screeplot of the the PCs for each dataset.

Produces shape residual data used in later analyses, stored in rda file 04.

### Load packages, functions, and data
Tree, mean shape, and metadata for phylogenetic analyses come from the allometry rodents script 04. 
```{r message = FALSE}
library(geomorph)   
library(vegan)      
library(dplyr)      
library(colorspace) 
library(phytools) 
library(geiger)      
library(mvMORPH)  
library(qpcR)


# This sets the wd to local
library(rstudioapi) # Safely Access the RStudio API
setwd(dirname(getActiveDocumentContext()$path))

source("../Functions/utilities.R")  # custom functions
load(file = "../Data/Processed/03-main-data.rda")
load(file = "../Data/Processed/03-tree-data.rda")
load(file = "../Data/Processed/03-graphing-vectors.rda")
load(file="../Data/Processed/allometry_residuals.rda")

```

#Phylogenetic signal of shape and size
```{r}


physignal(mean.shapes, aus.tree )

physignal(log(info.means$MeanCsize), aus.tree )

```

#Now we want to know if allometry affects all modules equally. 

```{r}

#add empty colums for values
Allometry <- matrix(ncol=7, nrow=length(names(mean_modules)), dimnames=list(c(names(mean_modules)), c("Df", "SS","MS" ,"Rsq", "F", "Z","P")) )


for (i in 1:length(mean_modules)){
  
temp <- procD.pgls( mean_modules[[i]]$shape ~ log(mean_modules[[i]]$csize), phy=aus.tree)

Allometry[i, 1:7] <- unlist( temp$aov.table[1,])

}


write.csv(Allometry, file="../Data/Results/Module_allometry.csv")

#it is also worth checking whether this pattern holds up if we use the overall csize of the skull, since the datasets are all aligned.

#add empty colums for values
Allometry_fullcsize <- matrix(ncol=7, nrow=length(names(mean_modules)), dimnames=list(c(names(mean_modules)), c("Df", "SS","MS" ,"Rsq", "F", "Z","P")) )


for (i in 1:length(mean_modules)){
  
temp <- procD.pgls( mean_modules[[i]]$shape ~ log(info.means$MeanCsize), phy=aus.tree)

Allometry_fullcsize[i, 1:7] <- unlist( temp$aov.table[1,])

}


```


### Prepare evolutionary allometry information (from AmNat paper) for Fig 1
```{r}
# Run Procrustes ANOVA with interaction
pgls.gdf <- geomorph.data.frame(coords = mean.shapes, size = info.means$MeanCsize, clade = info.means$Clade)

#pgls as per previous paper
fit_pgls_allo_evo <- procD.pgls(coords ~ log(size) , aus.tree, iter = 500, data = pgls.gdf, print.progress = FALSE)

#plain gls
fit_lm_allo <- procD.lm(coords ~ log(size) , iter = 500, data = pgls.gdf, print.progress = FALSE)


# Find Reg Score values for plotting purposes
evo_allo_rs <- plotAllometry(fit_pgls_allo_evo, size = info.means$MeanCsize, method = "RegScore")

# Also find Reg Score values for lm for plotting purposes
allo_rs <- plotAllometry(fit_lm_allo, size = info.means$MeanCsize, method = "RegScore")

```

### Export Figure 1 to pdf
Finally, we compare the full shape PCA to the shape residuals PCAs. All PCAs are relevant to Figure 2 heatmaps.
```{r}
# Set up Notomys abline
not.index <- which(info.means$Genus.y == "Notomys")
not.Reg <- evo_allo_rs$RegScore[not.index]
not.Cs <- info.means$MeanCsize[not.index]

# Set up for export and for multiple panels
part.coords <- c(-0.117, 0.095)  # plotting parameters for A, C, D, E placement

#setEPS()
#postscript("../Data/Results/Figure1_PCAs_6plot.eps")
mat <- matrix(c(1:6), 3, 2, byrow = TRUE)  # 6 plots, 3 rows, 2 columns
layout(mat, widths = rep(1, dim(mat)[2]), heights = rep(0.5, dim(mat)[1]))   

# A) FULL SHAPE PCA - diet & locomotion legend
par(mar = c(4.5, 4.5, 2, 1))
plot(x = -mean.PCA$x[, 1],  # Flip for consistency
                 y = mean.PCA$x[, 2],  
                 xlab = "PC1 (52.3%)", 
                 ylab = "PC2 (14.5%)",
                 xlim = c(-0.12, 0.12),
                 ylim = c(-0.1, 0.1),
                 col = col.diet.means, 
                 pch = pch.gld.means, 
                 bg = col.diet.means,
                 cex = 1.5,
                 cex.axis = 1.1, 
                 cex.lab = 1.3)
text(part.coords[1], part.coords[2], "a", cex = 1.5)
text(-0.045, 0.092, "full shape dataset (ecology)")
text(-0.078, 0.075, expression(paste("Marcy ", italic("et al. "), "2020")), cex = 0.8)

legend(-0.131, 0.067, legend = c("Arboreal", "Hopping", "Terrestrial", "Semiaquatic", "", "Carnivore", "Frugivore", "Granivore", "Folivore", "Omnivore"), col = c(rep("black", 4), "white", col.diet.5), border = NULL, pch = c(24:21, rep(16, 6)), cex = 0.8, bty = "n")

# B) Evolutionary Allometry plot - diet & locomotion legend
par(mar = c(4, 4, 1, 1))
plot(x = log(info.means$MeanCsize),
     y = -evo.allo.rs$RegScore,
     xlim = c(4.75, 6),
     col = col.diet.means, 
     pch = pch.gld.means, 
     bg = col.diet.means,
     main = " ",
     xlab = "Log centroid size", 
     ylab = "Shape  (Reg Score)",
     bty = "n")
text(4.8, 0.006, "b", cex = 1.5)
text(5.1, 0.0058, "evolutionary allometry")
abline(lm(evo_allo_rs$RegScore ~ log(info.means$MeanCsize)), col = "dark grey", lwd = 1)  # common evolutionary allometry line
#text(x = log(info.means$MeanCsize), y = -evo.allo.rs$RegScore, labels = info.means$Genus)  # labels genera on plot points

# C) FULL SHAPE PCA - genus colors
par(mar = c(4.5, 4.5, 2, 1))
plot(x = -mean.PCA$x[, 1],  # flip for consistency
                 y = mean.PCA$x[, 2],  
                 xlab = "PC1 (52.3%)", 
                 ylab = "PC2 (14.5%)",
                 xlim = c(-0.12, 0.12),
                 ylim = c(-0.1, 0.1),
                 col = alpha.col,
                 pch = 16,
                 bg = alpha.col,
                 cex = 1.5,
                 cex.axis = 1.1, 
                 cex.lab = 1.3)
text(part.coords[1], part.coords[2], "c", cex = 1.5)
text(-0.046, 0.092, "full shape dataset (genera)")

legend("bottomleft", legend = genera.phylo.full, col = col.phylo, pch = 16, cex = 0.68, pt.cex = 0.85, bty = "n", text.font = 3)

# D) shape residuals PCA - genus colors
par(mar = c(4.5, 4.5, 2, 1))
plot(x = res.PCA$x[, 1],  # no flip for consistency
                 y = -res.PCA$x[, 2],  
                 xlab = "PC1 (26.8%)", 
                 ylab = "PC2 (18.6%)",
                 xlim = c(-0.12, 0.12),
                 ylim = c(-0.1, 0.1),
                 col = alpha.col,
                 pch = 16,
                 bg = alpha.col,
                 cex = 1.5,
                 cex.axis = 1.1, 
                 cex.lab = 1.3)
text(part.coords[1], part.coords[2], "d", cex = 1.5)
text(-0.035, 0.092, "shape residual dataset (genera)")

# E) No Notomys shape residuals PCA - genus colors
par(mar = c(4.5, 4.5, 2, 1))
plot(x = wrap.PCA.r.n$x[, 1],  
                 y = wrap.PCA.r.n$x[, 2],  # no flip for consistency
                 xlab = "PC1 (30.1%)", 
                 ylab = "PC1 (16.7%)",
                 xlim = c(-0.12, 0.12),
                 ylim = c(-0.1, 0.1),
                 col = alpha.col.no, 
                 pch = 16, 
                 bg = alpha.col.no,
                 cex = 1.5,
                 cex.axis = 1.1, 
                 cex.lab = 1.3)
text(part.coords[1], part.coords[2], "e", cex = 1.5)
text(-0.034, 0.092, expression(paste("shape residual without ", italic("Notomys"))))

# F) shape Residuals PCA - ecology
par(mar = c(4.5, 4.5, 2, 1))
plot(x = res.PCA$x[, 1],  # no flip for consistency
                 y = -res.PCA$x[, 2],  
                 xlab = "PC1 (26.8%)", 
                 ylab = "PC1 (18.6%)",
                 xlim = c(-0.12, 0.12),
                 ylim = c(-0.1, 0.1),
                 col = col.diet.means, 
                 pch = pch.gld.means, 
                 bg = col.diet.means,
                 cex = 1.5,
                 cex.axis = 1.1, 
                 cex.lab = 1.3)
text(part.coords[1], part.coords[2], "f", cex = 1.5)
text(-0.053, 0.092, "shape residual (ecology)")

#dev.off()
```

#two-panel additional supplementary figure to compare allometry with and without phylogenetic correction

```{r}
setEPS()
postscript("../Data/Results/Supp_X_allometries.eps")

par(mfrow=c(2,1))

#evolutionary
par(mar = c(4, 4, 1, 1))
plot(x = log(info.means$MeanCsize),
     y = -evo.allo.rs$RegScore,
     xlim = c(4.75, 6),
     col = col.diet.means, 
     pch = pch.gld.means, 
     bg = col.diet.means,
     main = " ",
     xlab = "Log centroid size", 
     ylab = "Shape  (Reg Score)",
     bty = "n")
text(4.8, 0.001, "A", cex = 1.5)
text(5.05, 0.001, "evolutionary allometry")
abline(lm(-evo_allo_rs$RegScore ~ log(info.means$MeanCsize)), col = "dark grey", lwd = 1)  # common evolutionary allometry line


#lm
par(mar = c(4, 4, 1, 1))
plot(x = log(info.means$MeanCsize),
     y = allo_rs$RegScore,
     xlim = c(4.75, 6),
     col = col.diet.means, 
     pch = pch.gld.means, 
     bg = col.diet.means,
     main = " ",
     xlab = "Log centroid size", 
     ylab = "Shape  (Reg Score)",
     bty = "n")
text(4.8, 0.1, "B", cex = 1.5)
text(5.2, 0.1, "allometry in non-phylogenetic linear model")
abline(lm(allo_rs$RegScore ~ log(info.means$MeanCsize)), col = "dark grey", lwd = 1)  # common evolutionary allometry line

dev.off()


```



## 2) Compare the full shape and shape residual datasets
Mantel tests to compare full shape morphospace to the shape residual morphospace
```{r}
#A quick check of the correlation betwen full shape PC2 and residual shape PC1:
cor(mean.PCA$x[,2], res.PCA$x[,1])


# Make distance matrices
d.full.pca <- dist(mean.PCA$x)
d.res.pca <- dist(res.PCA$x)

# Mantel test for all PC scores
man.all <- mantel(d.full.pca, d.res.pca, method = "pearson", permutations = 1000) 
man.all$statistic
man.all$signif
```

### Mantel tests to compare just the 3 most important PCs
```{r}
# Make distance matrices
d.full.pc <- dist(mean.PCA$x[, 1:3])
d.res.pc <- dist(res.PCA$x[, 1:3])

# Mantel test for all PC scores
man.pc13 <- mantel(d.full.pc, d.res.pc, method = "pearson", permutations = 1000) 
man.pc13$statistic
man.pc13$signif
```

### Comparison of morphospaces without allometric PC1
```{r}
# Make distance matrices
d.full.wo1 <- dist(mean.PCA$x[, 2:dim(mean.PCA$x)[2]])
d.res.woL <- dist(res.PCA$x[, 1:dim(res.PCA$x)[2]-1])

# Mantel test for all PC scores
man.pc.wo1 <- mantel(d.full.wo1, d.res.woL, method = "pearson", permutations = 1000) 
man.pc.wo1$statistic
man.pc.wo1$signif
```


### Correlation of residual PC1 to full shape PC2 and vice versa
```{r}
cor.pc1 <- cor(mean.PCA$x[, 2], -res.PCA$x[, 1])
cor.pc1

cor.pc2 <- cor(mean.PCA$x[, 1], res.PCA$x[, 2])
cor.pc2

cor.pc2v3 <- cor(mean.PCA$x[, 3], res.PCA$x[, 2])
cor.pc2v3
```


## 3) Export skreeplots for PC importance scores for 3 separate datasets
Set up 3 datasets for screeplots
```{r}
# Full shape
full.PC.impt <- data.frame(PC = 1:(length(summary(mean.PCA)$PC.summary)), impt = unlist(summary(mean.PCA)[[1]][2,]))

# Shape residual
res.PC.impt <- data.frame(PC = 1:(length(summary(res.PCA)$PC.summary)), impt = unlist(summary(res.PCA)[[1]][2,]))

# Shape residual without Notomys wrap.PCA.r.n
rnot.PC.impt <- data.frame(PC = 1:(length(summary(wrap.PCA.r.n)$PC.summary)), impt = unlist(summary(wrap.PCA.r.n)[[1]][2,]))

```

### Create supplementary figure 1
```{r}
# Make simple multipanel figure
pdf("../Data/Results/SFigure1_Screeplots.pdf")
mat <- matrix(c(1:3), 3, 1, byrow = TRUE)  # 3 plots, 3 rows, 1 column
layout(mat, widths = rep(0.5, dim(mat)[2]), heights = rep(0.25, dim(mat)[1])) 

# 1) 
barplot(full.PC.impt[, 2], ylim = c(0, 0.5))
abline(h = 0.1, col = 'deeppink4')
mtext("A) full shape dataset", 3, line = 1.5, cex = 0.9, adj = 0, col = "black")
mtext("% of variance", 2, line = 2.5, cex = 0.83, col = "black")

# 2)
barplot(res.PC.impt[, 2], ylim = c(0, 0.5))
abline(h = 0.1, col = 'deeppink4')
mtext("B) shape residual dataset", 3, line = 1, cex = 0.9, adj = 0, col = "black")
mtext("% of variance", 2, line = 2.5, cex = 0.83, col = "black")

# 3)
barplot(rnot.PC.impt[, 2], ylim = c(0, 0.5))
abline(h = 0.1, col = 'deeppink4')
mtext(expression(paste("C) shape residual dataset without ", italic(Notomys))), 3, line = 1, cex = 0.9, adj = 0, col = "black")
mtext("% of variance", 2, line = 2.5, cex = 0.83, col = "black")
mtext("Principal component", 1, line = 1, cex = 0.83, col = "black")

dev.off()
```

### Save intermediate data
```{r}
save(mean.PCA, res.PCA, file = "../Data/Processed/04-PCAs.rda")
```
