---
title: "05-heatmap-both-datasets"
author: "Ariel Marcy"
date: "9/18/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '../eco-rodents')
```

# Heatmaps both with-allometry and without-allometry 3D coordinates
1) Make heatmaps with shape dataset including allometry.
        A) Smallest to mean to largest
        B) Smallest projected shape to mean to largest projected shape
        C) PC1 minimum to mean to PC maximum

Produces multipanel Figure 2.

2) Make heatmaps for residuals aka size-less shape dataset.
        A) PC1 min to mean to PC1 max
        B) PC1 min on subset without Notomys to mean to PC1 max of this subset
        C) PC2 min to mean to PC2 max

Produces multipanel Figure 3.

### Load packages and data
Including installing libraries for landvR package, which creates the heatmaps.
```{r message = FALSE}
library(geomorph)
library(devtools)
if(!require(dispRity)) install.packages("dispRity")  
if(!require(landvR)) install_github("TGuillerme/landvR")  # needed for heatmaps
library(landvR)
library(png)  # needed to work with screenshots
load(file = "../Data/Processed/03-main-data.rda")
load(file = "../Data/Processed/03-tree-data.rda")
load(file = "../Data/Processed/04-residual-data.rda")
```

## 1) Create heatmap figures
This script is dominated by the code needed to generate, export, and clean the heatmap images which are tiled in Figures 2 and 3. 

### Make heatmaps involving the smallest and largest native species 
We specify native species because technically *Mus musculus* has smallest centroid size but since it is not native, it is less relevant than *P. delicatulus*, which is only slightly larger than it.
```{r}
# Find mean configuration - i.e. mean shape in dataset
mean.config <- (select.procrustes(mean.shapes, selector = mean))[[1]]

# Measure spherical coordinates differences from the mean for each specimen for each landmark
diff.mean <- coordinates.difference(coordinates = mean.shapes, reference = mean.config, type = "spherical")

# Get index for smallest and largest native species 
sm.index <- which(info.means$Species == "del")
big.index <- which(info.means$MeanCsize == max(info.means$MeanCsize))

# Designate mean shapes for individual species with indices
smallest.mean <- mean.shapes[, , sm.index]
biggest.mean <- mean.shapes[, , big.index]
```

### Export the heatmaps for individual allometric shape changes
Export images for plots
```{r}
# Lateral view of mean to smallest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(mean.config, smallest.mean, col = heat.colors, col.val = diff.mean$Pseudomys_delicatulus[, "radius"])
view3d(phi = 90)
rgl.snapshot("../Data/Processed/Heatmaps/IND_mean-sm_lat.png")

# Lateral view of mean to biggest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(mean.config, biggest.mean, col = heat.colors, col.val = diff.mean$Uromys_caudimaculatus[, "radius"])
view3d(phi = 90)
rgl.snapshot("../Data/Processed/Heatmaps/IND_mean-bg_lat.png")

# Dorsal view of mean to smallest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(mean.config, smallest.mean, col = heat.colors, col.val = diff.mean$Pseudomys_delicatulus[, "radius"]) 
view3d(phi = 0)  # flip cranium
rgl.snapshot("../Data/Processed/Heatmaps/IND_mean-sm_dor.png")

# Dorsal view of mean to biggest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(mean.config, biggest.mean, col = heat.colors, col.val = diff.mean$Uromys_caudimaculatus[, "radius"])
view3d(phi = 0)  # flip cranium
rgl.snapshot("../Data/Processed/Heatmaps/IND_mean-bg_dor.png")
```

## Make heatmaps involving smallest and largest fitted allometric shapes
First find mean specimen, then smallest and largest. Uses Dr Thomas Guillerme's new package, `landvR` to see how landmarks vary between these specimens.
```{r}
# Fit specimens
allom.fit <- procD.lm(mean.shapes ~ log(info.means$MeanCsize))
fit.shapes <- arrayspecs(allom.fit$fitted, 325, 3)

# Find mean configuration - i.e. mean shape in dataset
allom.mean <- (select.procrustes(fit.shapes, selector = mean))[[1]]

# Measure spherical coordinates differences from the mean for each specimen for each landmark
allom.diff.mean <- coordinates.difference(coordinates = fit.shapes, reference = allom.mean, type = "spherical")

# Find largest and smallest native species from mean centroid size
smallest.shape <- fit.shapes[, , sm.index]
biggest.shape <- fit.shapes[, , big.index]
```

### Export the variation for fitted allometry shape change
Export images for plots
```{r}
# Lateral view of smallest to mean
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(allom.mean, smallest.shape, col = heat.colors, col.val = allom.diff.mean$Pseudomys_delicatulus[, "radius"])
view3d(phi = 90)
rgl.snapshot("../Data/Processed/Heatmaps/ALL_mean-sm_lat.png")

# Lateral view of mean to biggest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(allom.mean, biggest.shape, col = heat.colors, col.val = allom.diff.mean$Uromys_caudimaculatus[, "radius"])
view3d(phi = 90)
rgl.snapshot("../Data/Processed/Heatmaps/ALL_mean-bg_lat.png")

# Dorsal view of smallest to mean
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(allom.mean, smallest.shape, col = heat.colors, col.val = allom.diff.mean$Pseudomys_delicatulus[, "radius"]) 
view3d(phi = 0)  # flip cranium
rgl.snapshot("../Data/Processed/Heatmaps/ALL_mean-sm_dor.png")

# Dorsal view of mean to biggest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(allom.mean, biggest.shape, col = heat.colors, col.val = allom.diff.mean$Uromys_caudimaculatus[, "radius"])
view3d(phi = 0)  # flip cranium
rgl.snapshot("../Data/Processed/Heatmaps/ALL_mean-bg_dor.png")
```

### Make heatmaps for PC1 with allometry
Using hypothetical shapes for min and max
```{r}
# Find mean configuration - i.e. mean shape in dataset
PC.allom.mean <- (select.procrustes(mean.shapes, selector = mean))[[1]]

# Measure spherical coordinates differences from the mean for each specimen for each landmark
PC.allom.diff.mean <- coordinates.difference(coordinates = mean.shapes, reference = PC.allom.mean, type = "spherical")

# Ordinate the data
twoD.shape <- two.d.array(mean.shapes)
ordination <- stats::prcomp(twoD.shape)

# Force the symmetric component of shape into class "gpagen" (required for variation.range)
gpagen.shape <- list()
gpagen.shape$coords <- mean.shapes
gpagen.shape$consensus <- allom.mean
class(gpagen.shape) <- "gpagen"

# Measure extremes of variation from mean on PC1
PC1.var <- variation.range(gpagen.shape, return.ID = FALSE, axis = 1, ordination = ordination, type = "spherical")

# Wrap specimens on the tangent space
wrap.PCA <- gm.prcomp(mean.shapes)

# Select extremes (the hypotheticals that make the warp-meshes)
allom.PC1.min.h <- wrap.PCA$shapes[[1]]$max  # switched b/c axes are flipped
allom.PC1.max.h <- wrap.PCA$shapes[[1]]$min
```

### Export the PC1 variation for with allometry shape change
Export images for plots
```{r}
# Lateral view of mean to smallest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(PC.allom.mean, allom.PC1.min.h, col = heat.colors, col.val = PC1.var[, "radius"])
view3d(phi = 90)
rgl.snapshot("../Data/Processed/Heatmaps/PCA_mean-sm_lat.png")

# Lateral view of mean to biggest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(PC.allom.mean, allom.PC1.max.h, col = heat.colors, col.val = PC1.var[, "radius"])
view3d(phi = 90)
rgl.snapshot("../Data/Processed/Heatmaps/PCA_mean-bg_lat.png")

# Dorsal view of mean to smallest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(PC.allom.mean, allom.PC1.min.h, col = heat.colors, col.val = PC1.var[, "radius"]) 
view3d(phi = 0)  # flip cranium
rgl.snapshot("../Data/Processed/Heatmaps/PCA_mean-sm_dor.png")

# Dorsal view of mean to biggest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(PC.allom.mean, allom.PC1.max.h, col = heat.colors, col.val = PC1.var[, "radius"])
view3d(phi = 0)  # flip cranium
rgl.snapshot("../Data/Processed/Heatmaps/PCA_mean-bg_dor.png")
```

### Make heatmaps for PC1 with residual coordinates
Steps as above.
```{r}
# Find mean configuration - i.e. mean shape in dataset
res.mean <- (select.procrustes(residuals, selector = mean))[[1]]

# Measure spherical coordinates differences from the mean for each specimen for each landmark
res.diff.mean <- coordinates.difference(coordinates = residuals, reference = res.mean, type = "spherical")

# Ordinate the data
twoD.shape.r <- two.d.array(residuals)
ordination.r <- stats::prcomp(twoD.shape.r)

# Force the symmetric component of shape into class "gpagen" (required for variation.range)
gpagen.shape <- list()
gpagen.shape$coords <- residuals
gpagen.shape$consensus <- res.mean
class(gpagen.shape) <- "gpagen"

# Measure extremes of variation from mean on PC1
PC1.var.r <- variation.range(gpagen.shape, return.ID = FALSE, axis = 1, ordination = ordination.r, type = "spherical")

# Wrap specimens on the tangent space
wrap.PCA.r <- gm.prcomp(residuals)

# Select extreme specimens (the ones that make the warp-meshes)
res.PC1.min.h <- wrap.PCA.r$shapes[[1]]$min
res.PC1.max.h <- wrap.PCA.r$shapes[[1]]$max
```

### Export the variation for shape residual change
```{r}
# Lateral view of mean to smallest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean, res.PC1.min.h, col = heat.colors, col.val = PC1.var.r[, "radius"])
view3d(phi = 90)
rgl.snapshot("../Data/Processed/Heatmaps/RES_mean-sm_lat.png")

# Lateral view of mean to biggest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean, res.PC1.max.h, col = heat.colors, col.val = PC1.var.r[, "radius"])
view3d(phi = 90)
rgl.snapshot("../Data/Processed/Heatmaps/RES_mean-bg_lat.png")

# Dorsal view of mean to smallest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean, res.PC1.min.h, col = heat.colors, col.val = PC1.var.r[, "radius"]) 
view3d(phi = 0)  # flip cranium
rgl.snapshot("../Data/Processed/Heatmaps/RES_mean-sm_dor.png")

# Dorsal view of mean to biggest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean, res.PC1.max.h, col = heat.colors, col.val = PC1.var.r[, "radius"])
view3d(phi = 0)  # flip cranium
rgl.snapshot("../Data/Processed/Heatmaps/RES_mean-bg_dor.png")
```

### Make heatmaps for PC2 residual coordinates
Steps as above but for PC2; hypothetical shapes
```{r}
# Measure extremes of variation from mean on PC1
PC1.var.r2 <- variation.range(gpagen.shape, return.ID = FALSE, axis = 2, ordination = ordination.r, type = "spherical")

# Select extreme specimens (the ones that make the warp-meshes)
res.PC2.min.h <- wrap.PCA.r$shapes[[2]]$max  # switch for consistency
res.PC2.max.h <- wrap.PCA.r$shapes[[2]]$min

# Calculate distances between shapes to ensure correct point color
min.coords <- cbind(res.mean, res.PC2.min.h)
min.dist <- apply(min.coords, 1, function(x) dist(matrix(x, nrow = 2, byrow = TRUE)))
max.coords <- cbind(res.mean, res.PC2.max.h)
max.dist <- apply(max.coords, 1, function(x) dist(matrix(x, nrow = 2, byrow = TRUE)))
```

### Export the variation for shape residual change
```{r}
# Lateral view of mean to smallest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean, res.PC2.min.h, col = heat.colors, col.val = min.dist)
view3d(phi = 90)
rgl.snapshot("../Data/Processed/Heatmaps/RP2_mean-sm_lat.png")

# Lateral view of mean to biggest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean, res.PC2.max.h, col = heat.colors, col.val = min.dist)
view3d(phi = 90)
rgl.snapshot("../Data/Processed/Heatmaps/RP2_mean-bg_lat.png")

# Dorsal view of mean to smallest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean, res.PC2.min.h, col = heat.colors, col.val = max.dist) 
view3d(phi = 0)  # flip cranium
rgl.snapshot("../Data/Processed/Heatmaps/RP2_mean-sm_dor.png")

# Dorsal view of mean to biggest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean, res.PC2.max.h, col = heat.colors, col.val = max.dist)
view3d(phi = 0)  # flip cranium
rgl.snapshot("../Data/Processed/Heatmaps/RP2_mean-bg_dor.png")
```

### Remove Notomys and find hypothetical shapes
```{r}
# Remove Notomys from dataset
index.not <- which(info.means$Genus == "Not")
residuals.no <- residuals[, , -index.not]
info.means.no <- info.means[-index.not, ]

# Find new mean
res.mean.no <- (select.procrustes(residuals.no, selector = mean))[[1]]

# Measure spherical coordinates differences from the mean for each specimen for each landmark
res.diff.mean.no <- coordinates.difference(coordinates = residuals.no, reference = res.mean.no, type = "spherical")

# Ordinate the data
twoD.shape.r.n <- two.d.array(residuals.no)
ordination.r.n <- stats::prcomp(twoD.shape.r.n)

# Force the symmetric component of shape into class "gpagen" (required for variation.range)
gpagen.shape.no <- list()
gpagen.shape.no$coords <- residuals.no
gpagen.shape.no$consensus <- res.mean.no
class(gpagen.shape.no) <- "gpagen"

# Measure extremes of variation from mean on PC1
PC1.var.r.n <- variation.range(gpagen.shape.no, return.ID = FALSE, axis = 1, ordination = ordination.r, type = "spherical")

# Wrap specimens on the tangent space
wrap.PCA.r.n <- gm.prcomp(residuals.no)

# Select extreme specimens (the ones that make the warp-meshes)
res.PC1.min.h.n <- wrap.PCA.r.n$shapes[[2]]$min 
res.PC1.max.h.n <- wrap.PCA.r.n$shapes[[2]]$max
```

### Export residual no Notomys heatmaps
```{r}
# Lateral view of mean to smallest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean.no, res.PC1.min.h.n, col = heat.colors, col.val = PC1.var.r.n[, "radius"])
view3d(phi = 90)
rgl.snapshot("../Data/Processed/Heatmaps/NOt_mean-sm_lat.png")

# Lateral view of mean to biggest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean.no, res.PC1.max.h.n, col = heat.colors, col.val = PC1.var.r.n[, "radius"])
view3d(phi = 90)
rgl.snapshot("../Data/Processed/Heatmaps/NOt_mean-bg_lat.png")

# Dorsal view of mean to smallest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean.no, res.PC1.min.h.n, col = heat.colors, col.val = PC1.var.r.n[, "radius"]) 
view3d(phi = 0)  # flip cranium
rgl.snapshot("../Data/Processed/Heatmaps/NOt_mean-sm_dor.png")

# Dorsal view of mean to biggest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean.no, res.PC1.max.h.n, col = heat.colors, col.val = PC1.var.r.n[, "radius"])
view3d(phi = 0)  # flip cranium
rgl.snapshot("../Data/Processed/Heatmaps/NOt_mean-bg_dor.png")
```

### Find no Notomys PC2 shapes 
```{r}
# Select extreme specimens (the ones that make the warp-meshes)
res.PC2.min.h.n2 <- wrap.PCA.r.n$shapes[[2]]$min
res.PC2.max.h.n2 <- wrap.PCA.r.n$shapes[[2]]$max

# Calculate distances between shapes to ensure correct point color
min.coords.n2 <- cbind(res.mean.no, res.PC2.min.h.n2)
min.dist.n2 <- apply(min.coords.n2, 1, function(x) dist(matrix(x, nrow = 2, byrow = TRUE)))
max.coords.n2 <- cbind(res.mean.no, res.PC2.max.h.n2)
max.dist.n2 <- apply(max.coords.n2, 1, function(x) dist(matrix(x, nrow = 2, byrow = TRUE)))
```

### Export residual no Notomys heatmaps
```{r}
# Lateral view of mean to smallest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean.no, res.PC2.min.h.n2, col = heat.colors, col.val = min.dist.n2)
view3d(phi = 90)
rgl.snapshot("../Data/Processed/Heatmaps/NO2_mean-sm_lat.png")

# Lateral view of mean to biggest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean.no, res.PC2.max.h.n2, col = heat.colors, col.val = min.dist.n2)
view3d(phi = 90)
rgl.snapshot("../Data/Processed/Heatmaps/NO2_mean-bg_lat.png")

# Dorsal view of mean to smallest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean.no, res.PC2.min.h.n2, col = heat.colors, col.val = max.dist.n2) 
view3d(phi = 0)  # flip cranium
rgl.snapshot("../Data/Processed/Heatmaps/NO2_mean-sm_dor.png")

# Dorsal view of mean to biggest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean.no, res.PC2.max.h.n2, col = heat.colors, col.val = max.dist.n2)
view3d(phi = 0)  # flip cranium
rgl.snapshot("../Data/Processed/Heatmaps/NO2_mean-bg_dor.png")
```

### Load in all heatmaps
```{r}
# Remove white space from all heatmaps (requires ImageMagik installed on PC)
system("mogrify ../Data/Processed/Heatmaps/*.png -trim ../Data/Processed/Heatmaps/*.png")
```

```{r}
# Load in with allometry smallest to largest
indiv.sm.lat <- readPNG("../Data/Processed/Heatmaps/IND_mean-sm_lat.png") 
indiv.big.lat <- readPNG("../Data/Processed/Heatmaps/IND_mean-bg_lat.png")
indiv.sm.dor <- readPNG("../Data/Processed/Heatmaps/IND_mean-sm_dor.png")
indiv.big.dor <- readPNG("../Data/Processed/Heatmaps/IND_mean-bg_dor.png")

# Load in with fitted allometry smallest to largest
allom.sm.lat <- readPNG("../Data/Processed/Heatmaps/ALL_mean-sm_lat.png")
allom.big.lat <- readPNG("../Data/Processed/Heatmaps/ALL_mean-bg_lat.png")
allom.sm.dor <- readPNG("../Data/Processed/Heatmaps/ALL_mean-sm_dor.png")
allom.big.dor <- readPNG("../Data/Processed/Heatmaps/ALL_mean-bg_dor.png")

# Load in with allometry PC1
allPC.sm.lat <- readPNG("../Data/Processed/Heatmaps/PCA_mean-sm_lat.png")
allPC.big.lat <- readPNG("../Data/Processed/Heatmaps/PCA_mean-bg_lat.png")
allPC.sm.dor <- readPNG("../Data/Processed/Heatmaps/PCA_mean-sm_dor.png")
allPC.big.dor <- readPNG("../Data/Processed/Heatmaps/PCA_mean-bg_dor.png")

# Load in without allometry residuals PC1
resid.sm.lat <- readPNG("../Data/Processed/Heatmaps/RES_mean-sm_lat.png")
resid.big.lat <- readPNG("../Data/Processed/Heatmaps/RES_mean-bg_lat.png")
resid.sm.dor <- readPNG("../Data/Processed/Heatmaps/RES_mean-sm_dor.png")
resid.big.dor <- readPNG("../Data/Processed/Heatmaps/RES_mean-bg_dor.png")

# Load in without allometry residuals PC2
resP2.sm.lat <- readPNG("../Data/Processed/Heatmaps/RP2_mean-sm_lat.png")
resP2.big.lat <- readPNG("../Data/Processed/Heatmaps/RP2_mean-bg_lat.png")
resP2.sm.dor <- readPNG("../Data/Processed/Heatmaps/RP2_mean-sm_dor.png")
resP2.big.dor <- readPNG("../Data/Processed/Heatmaps/RP2_mean-bg_dor.png")

# Load in without allometry residuals PC1 - no Notomys
resid.sm.lat.no <- readPNG("../Data/Processed/Heatmaps/NOt_mean-sm_lat.png")
resid.big.lat.no <- readPNG("../Data/Processed/Heatmaps/NOt_mean-bg_lat.png")
resid.sm.dor.no <- readPNG("../Data/Processed/Heatmaps/NOt_mean-sm_dor.png")
resid.big.dor.no <- readPNG("../Data/Processed/Heatmaps/NOt_mean-bg_dor.png")

# Load in without allometry residuals PC2 - no Notomys
resid.sm.lat.n2 <- readPNG("../Data/Processed/Heatmaps/NO2_mean-sm_lat.png")
resid.big.lat.n2 <- readPNG("../Data/Processed/Heatmaps/NO2_mean-bg_lat.png")
resid.sm.dor.n2 <- readPNG("../Data/Processed/Heatmaps/NO2_mean-sm_dor.png")
resid.big.dor.n2 <- readPNG("../Data/Processed/Heatmaps/NO2_mean-bg_dor.png")
```

## 2) Make multi-panel Figure 2, heatmaps with allometry
```{r}
# Set up multi-panel figure; lateral 
pdf("../Data/Results/Figure2_Heatmap_Allometry.pdf")
mat <- matrix(c(1:12), 4, 3, byrow = FALSE)  # 12 plots, 4 rows, 3 columns
layout(mat, widths = rep(1, dim(mat)[2]), heights = rep(0.5, dim(mat)[1]))

## Fitted allometry
# 1) Mean to Small, with allometry, lateral view
par(mar = c(0, 0, 0, 0))  # sets the margins
plot(c(0, dim(allom.sm.lat)[2]), c(0, dim(allom.sm.lat)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(allom.sm.lat, 0, 0, dim(allom.sm.lat)[2], dim(allom.sm.lat)[1])
text(0, dim(allom.sm.lat)[1]-10, "a", cex = 2)
legend("topright", inset = c(0.28, 0), legend = "Fitted Allometry", bty = "n", cex = 1.2)
legend("bottomleft", legend = expression(paste("Mean to ", italic("P. delicatulus"), " fitted shape")), bty = "n", inset = c(0.03, 0.07))

# 2) Mean to Big, with allometry, lateral view
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(allom.big.lat)[2]), c(0, dim(allom.big.lat)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(allom.big.lat, 0, 0, dim(allom.big.lat)[2], dim(allom.big.lat)[1])
text(0, dim(allom.big.lat)[1]-10, "b", cex = 2)
legend("bottomleft", legend = expression(paste("Mean to ", italic("U. caudimaculatus"), " fitted shape")), bty = "n", inset = c(0.03, 0.07))

# 3) Mean to Small, with Allometry, dorsal view
par(mar = c(0, 0, 0, 0))  # sets the margins
plot(c(0, dim(allom.big.dor)[2]), c(0, dim(allom.big.dor)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(allom.sm.dor, 0, 0, dim(allom.big.dor)[2], dim(allom.big.dor)[1])
text(0, dim(allom.big.dor)[1]-10, "c", cex = 2)
legend("bottomleft", "Dorsal view of a", bty = "n", inset = c(0.2, 0.05))

# 4) Mean to Big, with allometry, dorsal view
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(allom.big.dor)[2]), c(0, dim(allom.big.dor)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(allom.big.dor, 0, 0, dim(allom.big.dor)[2], dim(allom.big.dor)[1])
text(0, dim(allom.big.dor)[1]-10, "d", cex = 2)
legend("bottomleft", "Dorsal view of b", bty = "n", inset = c(0.2, 0.05))

## Individuals
# 5) Mean to Small, with allometry, lateral view
par(mar = c(0, 0, 0, 0))  # sets the margins
plot(c(0, dim(indiv.sm.lat)[2]), c(0, dim(indiv.sm.lat)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(indiv.sm.lat, 0, 0, dim(indiv.sm.lat)[2], dim(indiv.sm.lat)[1])
text(0, dim(indiv.sm.lat)[1]-10, "e", cex = 2)
legend("topleft", inset = c(.02, 0), legend = "Smallest/Largest Species", bty = "n", cex = 1.2)
legend("bottomleft", legend = expression(paste("Mean to ", italic("P. delicatulus"))), bty = "n", inset = c(0.1, 0.07))

# 6) Mean to Big, with allometry, lateral view
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(indiv.big.lat)[2]), c(0, dim(indiv.big.lat)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(indiv.big.lat, 0, 0, dim(indiv.big.lat)[2], dim(indiv.big.lat)[1])
text(0, dim(indiv.big.lat)[1]-10, "f", cex = 2)
legend("bottomleft", legend = expression(paste("Mean to ", italic("U. caudimaculatus"))), bty = "n", inset = c(0.1, 0.07))

# 7) Mean to Small, with Allometry, dorsal view
par(mar = c(0, 0, 0, 0))  # sets the margins
plot(c(0, dim(indiv.big.dor)[2]), c(0, dim(indiv.big.dor)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(indiv.sm.dor, 0, 0, dim(indiv.big.dor)[2], dim(indiv.big.dor)[1])
text(0, dim(indiv.big.dor)[1]-10, "g", cex = 2)
legend("bottomleft", "Dorsal view of e", bty = "n", inset = c(0.2, 0.05))

# 8) Mean to Big, with allometry, dorsal view
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(indiv.big.dor)[2]), c(0, dim(indiv.big.dor)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(indiv.big.dor, 0, 0, dim(indiv.big.dor)[2], dim(indiv.big.dor)[1])
text(0, dim(indiv.big.dor)[1]-10, "h", cex = 2)
legend("bottomleft", "Dorsal view of f", bty = "n", inset = c(0.2, 0.05))

## PC1 
# 9) Mean to Small, PCA allometry, lateral view
par(mar = c(0, 0, 0, 0))  # sets the margins
plot(c(0, dim(allPC.sm.lat)[2]), c(0, dim(allPC.sm.lat)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(allPC.sm.lat, 0, 0, dim(allPC.sm.lat)[2], dim(allPC.sm.lat)[1])
text(0, dim(allPC.sm.lat)[1]-10, "i", cex = 2)
legend("topright", inset = c(0.16, 0), legend = "Principal Component 1", bty = "n", cex = 1.2)
legend("bottomleft", "Mean to hypothetical minimum", bty = "n", inset = c(0.1, 0.07))

# 10) Mean to Big, PCA allometry, lateral view
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(allPC.big.lat)[2]), c(0, dim(allPC.big.lat)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(allPC.big.lat, 0, 0, dim(allPC.big.lat)[2], dim(allPC.big.lat)[1])
text(0, dim(allPC.big.lat)[1]-10, "j", cex = 2)
legend("bottomleft", "Mean to hypothetical maximum", bty = "n", inset = c(0.1, 0.07))

# 11) Mean to Small, PCA allometry, dorsal view
par(mar = c(0, 0, 0, 0))  # sets the margins
plot(c(0, dim(allPC.big.dor)[2]), c(0, dim(allPC.big.dor)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(allPC.sm.dor, 0, 0, dim(allPC.big.dor)[2], dim(allPC.big.dor)[1])
text(0, dim(allPC.big.dor)[1]-10, "k", cex = 2)
legend("bottomleft", "Dorsal view of i", bty = "n", inset = c(0.2, 0.05))

# 12) Mean to Big, PCA allometry: dorsal view
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(allPC.big.dor)[2]), c(0, dim(allPC.big.dor)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(allPC.big.dor, 0, 0, dim(allPC.big.dor)[2], dim(allPC.big.dor)[1])
text(0, dim(allPC.big.dor)[1]-10, "l", cex = 2)
legend("bottomleft", "Dorsal view of j", bty = "n", inset = c(0.2, 0.05))

# Make gradient legend
legend_image <- as.raster(matrix(heat.colors(20), ncol = 1))
rasterImage(legend_image, 463, 165, 485, 230)
text(381, 223, "High variation", col = "dark grey", cex = 1)
text(431, 175, "Low", col = "dark grey", cex = 1)

dev.off()
```

## 3) Make multi-panel Figure 3: Heatmaps with Residuals
```{r}
# Set up muli-panel figure 
pdf("../Data/Results/Figure3_Heatmap_Residuals.pdf")
mat <- matrix(c(1:12), 4, 3, byrow = FALSE)  # 12 plots, 4 rows, 3 columns
layout(mat, widths = rep(1, dim(mat)[2]), heights = rep(0.5, dim(mat)[1])) 

## PC1 Residuals
# 1) Mean to Small, residuals PC1, lateral view
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resid.sm.lat)[2]), c(0, dim(resid.sm.lat)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(resid.sm.lat, 0, 0, dim(resid.sm.lat)[2], dim(resid.sm.lat)[1])
text(0, dim(resid.sm.lat)[1] - 10, "a", cex = 2)
legend("topright", inset = c(0.3, 0), legend = "PC1 shape residual", bty = "n", cex = 1.2)
legend("bottomleft", "Mean to hypothetical minimum", bty = "n", inset = c(0.1, 0.07))

# 2) Mean to Big, residuals PC1, lateral view
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resid.big.lat)[2]), c(0, dim(resid.big.lat)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(resid.big.lat, 0, 0, dim(resid.big.lat)[2], dim(resid.big.lat)[1])
text(0, dim(resid.big.lat)[1]-10, "b", cex = 2)
legend("bottomleft", "Mean to hypothetical maximum", bty = "n", inset = c(0.1, 0.07))

# 3) Mean to Small, residuals PC1, dorsal view
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resid.big.dor)[2]), c(0, dim(resid.big.dor)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(resid.sm.dor, 0, 0, dim(resid.big.dor)[2], dim(resid.big.dor)[1])
text(0, dim(resid.big.dor)[1]-10, "c", cex = 2)
legend("bottomleft", "Dorsal view of a", bty = "n", inset = c(0.2, 0.05))

# 4) Mean to Big, residuals PC1, dorsal view
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resid.big.dor)[2]), c(0, dim(resid.big.dor)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(resid.big.dor, 0, 0, dim(resid.big.dor)[2], dim(resid.big.dor)[1])
text(0, dim(resid.big.dor)[1]-10, "d", cex = 2)
legend("bottomleft", "Dorsal view of b", bty = "n", inset = c(0.2, 0.05))

## PC2 Residuals
# 5) Mean to Small, residuals PC2, lateral view
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resP2.sm.lat)[2]), c(0, dim(resP2.sm.lat)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(resP2.sm.lat, 0, 0, dim(resP2.sm.lat)[2], dim(resP2.sm.lat)[1])
text(0, dim(resP2.sm.lat)[1]-10, "e", cex = 2)
legend("topright", inset = c(0.3, 0), legend = "PC2 shape residual", bty = "n", cex = 1.2)
legend("bottomleft", "Mean to PC2 hypothetical minimum", bty = "n", inset = c(0.07, 0.07))

# 6) Mean to Big, residuals PC2, lateral view
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resP2.big.lat)[2]), c(0, dim(resP2.big.lat)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(resP2.big.lat, 0, 0, dim(resP2.big.lat)[2], dim(resP2.big.lat)[1])
text(0, dim(resP2.big.lat)[1]-10, "f", cex = 2)
legend("bottomleft", "Mean to PC2 hypothetical maximum", bty = "n", inset = c(0.07, 0.07))

# 7) Mean to Small, residuals PC2, dorsal view
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resP2.big.dor)[2]), c(0, dim(resP2.big.dor)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(resP2.sm.dor, 0, 0, dim(resP2.big.dor)[2], dim(resP2.big.dor)[1])
text(0, dim(resP2.big.dor)[1]-10, "g", cex = 2)
legend("bottomleft", "Dorsal view of e", bty = "n", inset = c(0.2, 0.05))

# 8) Mean to Big, residuals PC2, dorsal view
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resP2.big.dor)[2]), c(0, dim(resP2.big.dor)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(resP2.big.dor, 0, 0, dim(resP2.big.dor)[2], dim(resP2.big.dor)[1])
text(0, dim(resP2.big.dor)[1]-10, "h", cex = 2)
legend("bottomleft", "Dorsal view of f", bty = "n", inset = c(0.2, 0.05))

## PC2 Residuals -- without Notomys
# 9) Mean to Small, residuals PC2, lateral view - without Notomys
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resid.sm.lat.n2)[2]), c(0, dim(resid.sm.lat.n2)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(resid.sm.lat.n2, 0, 0, dim(resid.sm.lat.n2)[2], dim(resid.sm.lat.n2)[1])
text(0, dim(resid.sm.lat.n2)[1]-10, "i", cex = 2)
legend("topright", inset = c(0.12, 0), legend = expression(paste("PC2 residual (no ", italic("Notomys"), ")")), bty = "n", cex = 1.2)
legend("bottomleft", "Mean to hypothetical minimum", bty = "n", inset = c(0.1, 0.07))

# 10) Mean to Big, residuals PC2, lateral view - without Notomys
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resid.big.lat.n2)[2]), c(0, dim(resid.big.lat.n2)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(resid.big.lat.n2, 0, 0, dim(resid.big.lat.n2)[2], dim(resid.big.lat.n2)[1])
text(0, dim(resid.big.lat.n2)[1]-10, "j", cex = 2)
legend("bottomleft", "Mean to hypothetical maximum", bty = "n", inset = c(0.1, 0.07))

# 11) Mean to Small, residuals PC2, dorsal view - without Notomys
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resid.big.dor.n2)[2]), c(0, dim(resid.big.dor.n2)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(resid.sm.dor.n2, 0, 0, dim(resid.big.dor.n2)[2], dim(resid.big.dor.n2)[1])
text(0, dim(resid.big.dor.n2)[1]-10, "k", cex = 2)
legend("bottomleft", "Dorsal view of i", bty = "n", inset = c(0.2, 0.05))

# 12) Mean to Big, residuals PC2, dorsal view - without Notomys
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resid.big.dor.n2)[2]), c(0, dim(resid.big.dor.n2)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(resid.big.dor.n2, 0, 0, dim(resid.big.dor.n2)[2], dim(resid.big.dor.n2)[1])
text(0, dim(resid.big.dor.n2)[1]-10, "l", cex = 2)
legend("bottomleft", "Dorsal view of j", bty = "n", inset = c(0.2, 0.05))

# Make gradient legend
legend_image <- as.raster(matrix(heat.colors(20), ncol = 1))
rasterImage(legend_image, 443, 150, 465, 245)
text(365, 235, "High variation", col = "dark grey", cex = 1)
text(413, 165, "Low", col = "dark grey", cex = 1)

dev.off()
```