---
title: "05-plot_residual_PCAs"
author: "Ariel Marcy & Vera Weisbecker"
date: "8/15/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '../eco-rodents')
```

# Plot residual PCAs / Test for biome partitioning 
First, test for a phylogenetic signal difference in each of the 7 regions. 

Second, plot without-allometry PCA for all residuals.

Third, plot residuals by region -- a test for partitioning.

### Load packages, functions, and data
Tree, mean shape, and metadata for phylogenetic analyses come from the allometry rodents script 04. 
```{r message = FALSE}
library(geomorph)
library(dplyr)
source("../Functions/utilities.R")  # custom functions
load(file = "../Data/Processed/03-main-data.rda")
load(file = "../Data/Processed/03-tree-data.rda")
load(file = "../Data/Processed/03-graphing-vectors.rda")
load(file = "../Data/Processed/04-PCA.rda")
```

## 1) Extract residuals
Code thanks to Dr Vera Weisbecker
```{r}
# Use pgls of coords with size to get residuals
allom.GDF <- geomorph.data.frame(coords = mean.shapes, Csize = info.means$MeanCsize) 

dimnames(mean.shapes)[[3]] <- info.means$FullName

allometry.all <- procD.pgls(coords ~ Csize, aus.tree, data = allom.GDF)
summary(allometry.all)

# Turn the residuals into an appropriate array
allom.resid <- arrayspecs(allometry.all$pgls.residuals, dim(mean.shapes)[[1]], 3)

# Add the consensus shape (from the GPA) to the residuals if you want to understand the landmark variation in the residuals - needed for script 06
residuals <- allom.resid + array(mean.shapes, dim(allom.resid))
```

## 2) Plot but with only residuals no consensus added
```{r}
res.PCA <- plotTangentSpace(allom.resid, groups = alpha.col)
PlotPCA(res.PCA, 1, 2, alpha.col, alpha.pch, flip.axis2 = TRUE)
legend("bottomright", phylo.names, col = col.unique.spp, pch = pch.unique.spp, pt.bg = col.unique.spp, cex = .53, pt.cex = .7, ncol = 2)
```

## 3) Plot residuals with consensus (WITHOUT allometry) - IMPORTANT FIGURE
Color and legend by species
```{r}
res.PCA <- plotTangentSpace(residuals, groups = alpha.col, label = alpha.names)

# Make axis labels
PCs <- res.PCA$pc.summary$importance
PCx.per <- round(PCs[2, 1] * 100, digits = 1)  # % with 1 decimal
PCx.lab <- paste("PC1", " (", PCx.per, "%)", sep = "")
PCy.per <- round(PCs[2, 2] * 100, digits = 1)
PCy.lab <- paste("PC2", " (", PCy.per, "%)", sep = "")
        
# Pretty plot
PCA.plot <- plot(x = -res.PCA$pc.scores[, 1],  # Flip for consistency
                 y = -res.PCA$pc.scores[, 2],  
                 xlab = PCx.lab, 
                 ylab = PCy.lab,
                 asp = TRUE,
                 xlim = c(-0.12, 0.18),
                 ylim = c(-0.18, 0.12),
                 col = alpha.col, 
                 pch = alpha.pch, 
                 bg = alpha.col,
                 cex = 1.5,
                 cex.axis = 1.3, 
                 cex.lab = 1.3)
legend("bottomright", phylo.names, col = col.unique.spp, pch = pch.unique.spp, pt.bg = col.unique.spp, cex = .57, pt.cex = .7, ncol = 2)
```

## 4) Plot by guild and compare with mean allometry shapes 
Recolor plot above with points by diet and guild

NOTE CLEAN UP NEEDED HERE ONCE FINAL MULTI-PANEL FIGURE DECIDED ON
```{r}
# Define color pallete for diets of all specimens
col.diet.means <- PlotByGroup(info.means, "Feeding", col.diet.5)

# Points: Aquatic (circle), Arboreal (triangle), Terrestrial (square)
pch.gld.means <- PlotByGroup(info.means, "Guild", pch.gld.4)

# Make PCA of mean shapes with allometry
PCA.mean.allo <- plotTangentSpace(mean.shapes)

# Make axis labels
PCs.allo <- PCA.mean.allo$pc.summary$importance
PCx.per.allo <- round(PCs.allo[2, 1] * 100, digits = 1)  # % with 1 decimal
PCx.lab.allo <- paste("PC1", " (", PCx.per.allo, "%)", sep = "")
PCy.per.allo <- round(PCs.allo[2, 2] * 100, digits = 1)
PCy.lab.allo <- paste("PC2", " (", PCy.per.allo, "%)", sep = "")

# Allometry PCA
PCA.plot <- plot(x = -PCA.mean.allo$pc.scores[, 1],  # Flip for consistency
                 y = PCA.mean.allo$pc.scores[, 2],  
                 xlab = PCx.lab.allo, 
                 ylab = PCy.lab.allo,
                 asp = TRUE,
                 xlim = c(-0.1, 0.2),
                 col = alpha.col,
                 pch = alpha.pch,
                 bg = alpha.col,
                 #col = col.diet.means, 
                 #pch = pch.gld.means, 
                 #bg = col.diet.means,
                 main = "PCA of mean shapes with allometry",
                 cex = 1.5,
                 cex.axis = 1.3, 
                 cex.lab = 1.3)
legend("bottomright", legend = phylo.names, col = col.unique.spp, pch = pch.unique.spp, pt.bg = col.unique.spp, cex = 0.535, pt.cex = .7, ncol = 2)

legend("topright", legend = c("Carnivore ", "Frugivore ", "Granivore ", "Herbivore ", "Omnivore "), col = col.diet.5, border = NULL, pch = 15, cex = 0.9)
legend("bottomright", legend = c("Aquatic", "Arboreal", "Terrestrial", "Hopping"), col = "black", border = NULL, pch = c(21, 24, 22, 23), cex = 0.9)
```

### Residual PCA
```{r}
# Pretty plot
PCA.plot <- plot(x = -res.PCA$pc.scores[, 1],  # Flip for consistency
                 y = -res.PCA$pc.scores[, 2],  
                 xlab = PCx.lab, 
                 ylab = PCy.lab,
                 asp = TRUE,
                 xlim = c(-0.12, 0.18),
                 ylim = c(-0.18, 0.12),
                 col = alpha.col,
                 pch = alpha.pch,
                 bg = alpha.col,
                 #col = col.diet.means, 
                 #pch = pch.gld.means, 
                 #bg = col.diet.means,
                 main = "PCA of mean residual shapes (without allometry)",
                 cex = 1.5,
                 cex.axis = 1.3, 
                 cex.lab = 1.3)
legend("bottomright", legend = phylo.names, col = col.unique.spp, pch = pch.unique.spp, pt.bg = col.unique.spp, cex = 0.535, pt.cex = .7, ncol = 2)

#legend("topright", legend = c("Carnivore ", "Frugivore ", "Granivore ", "Herbivore ", "Omnivore "), col = col.diet.5, border = NULL, pch = 15, cex = 0.9)

#legend("bottomright", legend = c("Aquatic", "Arboreal", "Terrestrial", "Hopping"), col = "black", border = NULL, pch = c(21, 24, 22, 23), cex = 0.9)
```

### Save intermediate data
```{r}
save(res.PCA, allom.resid, residuals, file = "../Data/Processed/05-residual-data.rda")
```