---
title: "05-test-partitioning"
author: "Gabriele Sansalone & Ariel Marcy"
date: "8/15/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '../eco-rodents')
```

# Test biome partitioning 

First, testing for a phylogenetic signal difference in each of the 7 regional biomes. 

Second, code to run function sepregr.R by Dr Gabriele Sansalone

Tree, mean shape, and metadata for phylogenetic analyses come from the allometry rodents script 04. 

### Load packages, functions, and data
```{r message = FALSE}
library(geomorph)
library(dplyr)
library(geiger)
source("../Functions/utilities.R")  # custom functions
source("../Functions/sepregr.R")  # function by Dr Sansalone
load(file = "../Data/Processed/03-sm-PCA-with-color.rda")
load(file = "../Data/Processed/04-sm-data-with-traits.rda")
load(file = "../Data/Processed/allometry_rodents_04-phylo-data.rda")  # includes phylogenetic tree from Allometry Rodents paper
```

## Assess phylogenetic signal overall
Baseline comparison.
```{r}
# Test phylogenetic signal
physig.all <- physignal(mean.shapes.tree, aus.tree)
summary(physig.all)
```

## Assess phylogenetic signal in each region partition
Here, again, we subset the PCA scores by species which inhabit the 7 different ecological regions of Australia. We'll use a loop to test the phylogenetic signal in each region. 

Evolutionary tree comes from the Allometry Paper repo. 
```{r}
# Set up loop for particular regions in Australia
region.names <- c("Savannah", "Arid Zone", "NE Wet Forest", "NE Dry Forest", "Pilbara", "South West", "South East")
n.taxa <- c(rep("NA", length(region.names)))
K.region <- c(rep("NA", length(region.names)))
p.region <- c(rep("NA", length(region.names)))

# Loop through all 7 regions to plot morphospace of species found in each
for (i in 1:7) {
        # Subset info.traits and shape coords by region
        in.region <- FoundInRegion(info.phy, regions = i)
        info.region <- info.phy[in.region, ]
        shape.region <- shape.phy[, , in.region]
        dimnames(shape.region)[[3]] <- info.region$FullName
        
        # Subset tree by species in the region
        concord <- name.check(aus.tree, data.names = info.region$FullName)  # check concordance
        region.tree <- drop.tip(aus.tree, concord$tree_not_data)  # prune tree to regional sample
        n.taxa[i] <- length(region.tree$tip.label)
        
        # Re-order to match tree tip order
        tree.order <- match(region.tree$tip.label, info.region$FullName)
        info.region <- info.region[tree.order, ]  # metadata
        shape.region <- shape.region[, , tree.order]  # shape data
        
        # Test phylogenetic signal
        physig.region <- physignal(shape.region, region.tree)
        K.region[i] <- round(physig.region$phy.signal, 3)
        p.region[i] <- as.numeric(physig.region$pvalue)
}
```

### Make and export Table TK
Create and export Table TK. Includes Bonferroni p-value adjustment for running 7 biomes. 
```{r}
p.region <- round(p.adjust(p.region, method = "bonferroni"), 3)

table <- cbind(n.taxa, K.region, p.region)
colnames(table) <- c("N Taxa", "Phy. Signal", "p Value")
row.names(table) <- region.names

write.table(table, "../Data/Results/Table_RegionPhySig.csv", sep = ",", col.names = NA)
```



### Recycle Bin ###

### A look at ANOVA differences in centroid sizes
```{r}
test.data <- info.traits[, c("Taxa", "Genus", "Species", "CatNum", "Order", "Csize", "Feeding", "Guild")]  # relevant columns

# Take out aquatic carnivores, which skew data
#aquatic <- which(info.traits$Guild == "Aquatic")
#test.data <- test.data[-aquatic, ]  
#test.shape <- shape[, , -aquatic]


# Get summary statistics on each group
group_by(test.data$Csize, Guild) %>% 
        summarise(count = n(test.data$Csize), 
                  mean = mean(test.data$Csize),
                  sd = sd(test.data$Csize)) 

## Visualize data
install.packages("ggpubr")
library("ggpubr")

# Box Plots
ggboxplot(test.data, x = "Guild", y = "Csize", 
          color = "Guild", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
          order = c("Aquatic", "Arboreal", "Terrestrial"),
          ylab = "Csize", xlab = "Guild")

# Mean Plots
ggline(test.data, x = "Guild", y = "Csize", 
       add = c("mean_se", "jitter"), 
       order = c("Aquatic", "Arboreal", "Terrestrial"),
       ylab = "Csize", xlab = "Guild")

# Run ANOVA on just Csize and Guild
guild.aov <- aov(Csize ~ Guild, data = test.data)
summary(guild.aov)

# Multiple comparisons
TukeyHSD(guild.aov)
```

```{r}
# Run Procrustes ANOVA for guild
test.GDF <- geomorph.data.frame(coords = shape, size = test.data$Csize, guild = test.data$Guild)
procANOVA <- procD.lm(coords ~ size * guild, data = test.GDF)

# Run Procrustes ANOVA for guild
test.GDF <- geomorph.data.frame(coords = test.shape, size = test.data$Csize, feed = test.data$Feeding)
procANOVA <- procD.lm(coords ~ size * guild, data = test.GDF)
```



### Remove *Z. ped* from dataset
*Z. pedunculatus* is missing from the phylogenetic tree so needs to be removed from the shape dataset as well. Only removes 2 observations.
```{r}
is.Z.ped <- which(info.traits$Taxa == "Z_ped")
info.phy <- info.traits[-is.Z.ped, ]  # metadata
shape.phy <- shape[, , -is.Z.ped]  # shape coordinates
```
