---
title: "05-test-partitioning"
author: "Gabriele Sansalone & Ariel Marcy"
date: "8/15/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '../eco-rodents')
```

# Testing biome partitioning 

First, testing for a phylogenetic signal difference in each of the 7 regional biomes. 

Second, code to run function sepregr.R by Dr Gabriele Sansalone

### Load packages, functions, and data
```{r message = FALSE}
library(geomorph)
library(geiger)
source("../Functions/utilities.R")  # custom functions
source("../Functions/sepregr.R")  # function by Dr Sansalone
load(file = "../Data/Processed/03-sm-PCA-with-color.rda")
load(file = "../Data/Processed/04-sm-data-with-traits.rda")
load(file = "../Data/Processed/allometry_rodents_04-phylo-data.rda")  # includes phylogenetic tree from Allometry Rodents paper
```

### Remove *Z. ped* from dataset
*Z. pedunculatus* is missing from the phylogenetic tree so needs to be removed from the shape dataset as well. Only removes 2 observations.
```{r}
is.Z.ped <- which(info.traits$Taxa == "Z_ped")
info.phy <- info.traits[-is.Z.ped, ]  # metadata
shape.phy <- shape[, , -is.Z.ped]  # shape coordinates
```

## Assess phylogenetic signal in each region partition
Here, again, we subset the PCA scores by species which inhabit the 7 different ecological regions of Australia. We'll use a loop to test the phylogenetic signal in each region. 

Evolutionary tree comes from the Allometry Paper repo. 
```{r}
# Set up loop for particular regions in Australia
region.names <- c("Savannah", "Arid Zone", "NE Wet Forest", "NE Dry Forest", "Pilbara", "South West", "South East")

# Loop through all 7 regions to plot morphospace of species found in each
for (i in 1:7) {
        # Subset info.traits and shape coords by region
        in.region <- FoundInRegion(info.phy, regions = i)
        info.region <- info.phy[in.region, ]
        shape.region <- shape.phy[, , in.region]
        dimnames(shape.region)[[3]] <- info.region$FullName
        
        # Subset tree by species in the region
        concord <- name.check(aus.tree, data.names = info.region$FullName)  # check concordance
        region.tree <- drop.tip(aus.tree, concord$tree_not_data)  # prune tree to regional sample
        
        # Re-order to match tree tip order
        tree.order <- match(region.tree$tip.label, info.region$FullName)
        info.region <- info.region[tree.order, ]  # metadata
        shape.region <- shape.region[, , tree.order]  # shape data
        
        # Test phylogenetic signal
        print(region.names[i])
        physignal(shape.region, region.tree)
}
```
