---
title: "06-heatmap-both-datasets"
author: "Ariel Marcy"
date: "9/18/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '../eco-rodents')
```

# Heatmaps both with-allometry and without-allometry 3D coordinates
1) Make heatmaps for with-allometry.

2) Make heatmaps for without-allometry aka residuals aka size-less shape.

3) Make multi-panel figure for comparisons across heatmaps.

4) Directly compare by subtracting residual change from allometry change.

### Load packages, functions, and data
Tree, mean shape, and metadata for phylogenetic analyses come from the allometry rodents script 04. 
```{r message = FALSE}
library(geomorph)
library(data.table)
library(vegan)
library(geiger)  # for phylogenies
library(png)  # needed to work with screenshots
library(colorspace)  # plot colors

# Load libraries (and install if necessary) for landvR package
if(!require(devtools)) install.packages("devtools")
if(!require(dispRity)) install.packages("dispRity")
if(!require(landvR)) install_github("TGuillerme/landvR")
source("../Functions/utilities.R")  # custom functions
source("../Functions/plotGMPhyloMorphoSpace_plotmod.R")  # modified geomorph function
load(file = "../Data/Processed/03-main-data.rda")
load(file = "../Data/Processed/03-tree-data.rda")
load(file = "../Data/Processed/03-graphing-vectors.rda")
load(file = "../Data/Processed/04-PCA.rda")
load(file = "../Data/Processed/05-residual-data.rda")
```

## 1) Prepare heatmaps
This script is dominated by the code needed to generate, export, and clean the heatmap images which are tiled in Figures 2 and 3. 

### Make heatmaps involving the smallest and largest native species 
We specify native species because technically *Mus musculus* has smallest centroid size but since it is not native, it is less relevant than *P. delicatulus*, which is only slightly larger than it.
```{r}
# Find mean configuration - i.e. mean shape in dataset
mean.config <- (select.procrustes(mean.shapes, selector = mean))[[1]]

# Measure spherical coordinates differences from the mean for each specimen for each landmark
diff.mean <- coordinates.difference(coordinates = mean.shapes, reference = mean.config, type = "spherical")

# Get index for smallest and largest native species 
sm.index <- which(info.means$Species == "del")
big.index <- which(info.means$MeanCsize == max(info.means$MeanCsize))

# Designate mean shapes for individual species with indices
smallest.mean <- mean.shapes[, , sm.index]
biggest.mean <- mean.shapes[, , big.index]
```

### Export the heatmaps for individual allometric shape changes
Export images for plots
```{r}
# Lateral view of mean to smallest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(mean.config, smallest.mean, col = heat.colors, col.val = diff.mean$Pseudomys_delicatulus[, "radius"])
view3d(phi = 90)
rgl.snapshot("../Data/Processed/heatmap_IND_sm-mean_lat.png")

# Lateral view of mean to biggest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(mean.config, biggest.mean, col = heat.colors, col.val = diff.mean$Uromys_caudimaculatus[, "radius"])
view3d(phi = 90)
rgl.snapshot("../Data/Processed/heatmap_IND_mean-big_lat.png")

# Dorsal view of mean to smallest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(mean.config, smallest.mean, col = heat.colors, col.val = diff.mean$Pseudomys_delicatulus[, "radius"]) 
view3d(phi = 0)  # flip cranium
rgl.snapshot("../Data/Processed/heatmap_IND_sm-mean_dor.png")

# Dorsal view of mean to biggest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(mean.config, biggest.mean, col = heat.colors, col.val = diff.mean$Uromys_caudimaculatus[, "radius"])
view3d(phi = 0)  # flip cranium
rgl.snapshot("../Data/Processed/heatmap_IND_mean-big_dor.png")
```

## Make heatmaps involving smallest and largest fitted allometric shapes
First find mean specimen, then smallest and largest. Uses Dr Thomas Guillerme's new package, `landvR` to see how landmarks vary between these specimens.
```{r}
# Fit specimens
allom.fit <- procD.lm(mean.shapes~info.means$MeanCsize)
fit.shapes <- arrayspecs(allom.fit$fitted, 325, 3)

# Find mean configuration - i.e. mean shape in dataset
allom.mean <- (select.procrustes(fit.shapes, selector = mean))[[1]]

# Measure spherical coordinates differences from the mean for each specimen for each landmark
allom.diff.mean <- coordinates.difference(coordinates = fit.shapes, reference = allom.mean, type = "spherical")

# Find largest and smallest native species from mean centroid size
smallest.shape <- fit.shapes[, , sm.index]
biggest.shape <- fit.shapes[, , big.index]
```

### Export the variation for fitted allometry shape change
Export images for plots
```{r}
# Lateral view of smallest to mean
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(allom.mean, smallest.shape, col = heat.colors, col.val = allom.diff.mean$Pseudomys_delicatulus[, "radius"])
view3d(phi = 90)
rgl.snapshot("../Data/Processed/heatmap_ALL_sm-mean_lat.png")

# Lateral view of mean to biggest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(allom.mean, biggest.shape, col = heat.colors, col.val = allom.diff.mean$Uromys_caudimaculatus[, "radius"])
view3d(phi = 90)
rgl.snapshot("../Data/Processed/heatmap_ALL_mean-big_lat.png")

# Dorsal view of smallest to mean
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(allom.mean, smallest.shape, col = heat.colors, col.val = allom.diff.mean$Pseudomys_delicatulus[, "radius"]) 
view3d(phi = 0)  # flip cranium
rgl.snapshot("../Data/Processed/heatmap_ALL_sm-mean_dor.png")

# Dorsal view of mean to biggest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(allom.mean, biggest.shape, col = heat.colors, col.val = allom.diff.mean$Uromys_caudimaculatus[, "radius"])
view3d(phi = 0)  # flip cranium
rgl.snapshot("../Data/Processed/heatmap_ALL_mean-big_dor.png")
```

### Make heatmaps for PC1 with allometry
Using hypothetical shapes for min and max
```{r}
# Find mean configuration - i.e. mean shape in dataset
PC.allom.mean <- (select.procrustes(mean.shapes, selector = mean))[[1]]

# Measure spherical coordinates differences from the mean for each specimen for each landmark
PC.allom.diff.mean <- coordinates.difference(coordinates = mean.shapes, reference = PC.allom.mean, type = "spherical")

# Ordinate the data
twoD.shape <- two.d.array(mean.shapes)
ordination <- stats::prcomp(twoD.shape)

# Force the symmetric component of shape into class "gpagen" (required for variation.range)
gpagen.shape <- list()
gpagen.shape$coords <- mean.shapes
gpagen.shape$consensus <- allom.mean
class(gpagen.shape) <- "gpagen"

# Measure extremes of variation from mean on PC1
PC1.var <- variation.range(gpagen.shape, return.ID = FALSE, axis = 1, ordination = ordination, type = "spherical")

# Wrap specimens on the tangent space
wrap.PCA <- plotTangentSpace(mean.shapes, verbose = FALSE)

# Select extremes (the hypotheticals that make the warp-meshes)
allom.PC1.min.h <- wrap.PCA$pc.shapes[[2]]  # switched b/c axes are flipped
allom.PC1.max.h <- wrap.PCA$pc.shapes[[1]]
```

### Export the PC1 variation for with allometry shape change
Export images for plots
```{r}
# Lateral view of mean to smallest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(PC.allom.mean, allom.PC1.min.h, col = heat.colors, col.val = PC1.var[, "radius"])
view3d(phi = 90)
rgl.snapshot("../Data/Processed/heatmap_PCA_sm-mean_lat.png")

# Lateral view of mean to biggest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(PC.allom.mean, allom.PC1.max.h, col = heat.colors, col.val = PC1.var[, "radius"])
view3d(phi = 90)
rgl.snapshot("../Data/Processed/heatmap_PCA_mean-big_lat.png")

# Dorsal view of mean to smallest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(PC.allom.mean, allom.PC1.min.h, col = heat.colors, col.val = PC1.var[, "radius"]) 
view3d(phi = 0)  # flip cranium
rgl.snapshot("../Data/Processed/heatmap_PCA_sm-mean_dor.png")

# Dorsal view of mean to biggest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(PC.allom.mean, allom.PC1.max.h, col = heat.colors, col.val = PC1.var[, "radius"])
view3d(phi = 0)  # flip cranium
rgl.snapshot("../Data/Processed/heatmap_PCA_mean-big_dor.png")
```

### Make heatmaps for PC1 with residual coordinates
Steps as above.
```{r}
# Find mean configuration - i.e. mean shape in dataset
res.mean <- (select.procrustes(residuals, selector = mean))[[1]]

# Measure spherical coordinates differences from the mean for each specimen for each landmark
res.diff.mean <- coordinates.difference(coordinates = residuals, reference = res.mean, type = "spherical")

# Ordinate the data
twoD.shape.r <- two.d.array(residuals)
ordination.r <- stats::prcomp(twoD.shape.r)

# Force the symmetric component of shape into class "gpagen" (required for variation.range)
gpagen.shape <- list()
gpagen.shape$coords <- residuals
gpagen.shape$consensus <- res.mean
class(gpagen.shape) <- "gpagen"

# Measure extremes of variation from mean on PC1
PC1.var.r <- variation.range(gpagen.shape, return.ID = FALSE, axis = 1, ordination = ordination.r, type = "spherical")

# Wrap specimens on the tangent space
wrap.PCA.r <- plotTangentSpace(residuals, verbose = FALSE)

# Select extreme specimens (the ones that make the warp-meshes)
res.PC1.min.h <- wrap.PCA.r$pc.shapes[[2]]  # Flipped b/c axes are flipped
res.PC1.max.h <- wrap.PCA.r$pc.shapes[[1]]
```

### Export the variation for residual shape change
```{r}
# Lateral view of mean to smallest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean, res.PC1.min.h, col = heat.colors, col.val = PC1.var.r[, "radius"])
view3d(phi = 90)
rgl.snapshot("../Data/Processed/heatmap_RES_sm-mean_lat.png")

# Lateral view of mean to biggest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean, res.PC1.max.h, col = heat.colors, col.val = PC1.var.r[, "radius"])
view3d(phi = 90)
rgl.snapshot("../Data/Processed/heatmap_RES_mean-big_lat.png")

# Dorsal view of mean to smallest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean, res.PC1.min.h, col = heat.colors, col.val = PC1.var.r[, "radius"]) 
view3d(phi = 0)  # flip cranium
rgl.snapshot("../Data/Processed/heatmap_RES_sm-mean_dor.png")

# Dorsal view of mean to biggest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean, res.PC1.max.h, col = heat.colors, col.val = PC1.var.r[, "radius"])
view3d(phi = 0)  # flip cranium
rgl.snapshot("../Data/Processed/heatmap_RES_mean-big_dor.png")
```

### Make heatmaps for PC2 residual coordinates
Steps as above but for PC2; hypothetical shapes
```{r}
# Measure extremes of variation from mean on PC1
PC1.var.r2 <- variation.range(gpagen.shape, return.ID = FALSE, axis = 2, ordination = ordination.r, type = "spherical")

# Select extreme specimens (the ones that make the warp-meshes)
res.PC2.min.h <- wrap.PCA.r$pc.shapes[[4]]
res.PC2.max.h <- wrap.PCA.r$pc.shapes[[3]]
```

### Export the variation for residual shape change
```{r}
# Lateral view of mean to smallest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean, res.PC2.min.h, col = heat.colors, col.val = PC1.var.r2[, "radius"])
view3d(phi = 90)
rgl.snapshot("../Data/Processed/heatmap_RP2_sm-mean_lat.png")

# Lateral view of mean to biggest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean, res.PC2.max.h, col = heat.colors, col.val = PC1.var.r2[, "radius"])
view3d(phi = 90)
rgl.snapshot("../Data/Processed/heatmap_RP2_mean-big_lat.png")

# Dorsal view of mean to smallest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean, res.PC2.min.h, col = heat.colors, col.val = PC1.var.r2[, "radius"]) 
view3d(phi = 0)  # flip cranium
rgl.snapshot("../Data/Processed/heatmap_RP2_sm-mean_dor.png")

# Dorsal view of mean to biggest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean, res.PC2.max.h, col = heat.colors, col.val = PC1.var.r2[, "radius"])
view3d(phi = 0)  # flip cranium
rgl.snapshot("../Data/Processed/heatmap_RP2_mean-big_dor.png")
```

### Remove Notomys and find hypothetical shapes
```{r}
# Remove Notomys from dataset
index.not <- which(info.means$Genus == "Not")
residuals.no <- residuals[, , -index.not]
info.means.no <- info.means[-index.not, ]

# Find new mean
res.mean.no <- (select.procrustes(residuals.no, selector = mean))[[1]]

# Measure spherical coordinates differences from the mean for each specimen for each landmark
res.diff.mean.no <- coordinates.difference(coordinates = residuals.no, reference = res.mean.no, type = "spherical")

# Ordinate the data
twoD.shape.r.n <- two.d.array(residuals.no)
ordination.r.n <- stats::prcomp(twoD.shape.r.n)

# Force the symmetric component of shape into class "gpagen" (required for variation.range)
gpagen.shape.no <- list()
gpagen.shape.no$coords <- residuals.no
gpagen.shape.no$consensus <- res.mean.no
class(gpagen.shape.no) <- "gpagen"

# Measure extremes of variation from mean on PC1
PC1.var.r.n <- variation.range(gpagen.shape.no, return.ID = FALSE, axis = 1, ordination = ordination.r, type = "spherical")

# Wrap specimens on the tangent space
wrap.PCA.r.n <- plotTangentSpace(residuals.no, verbose = FALSE)

# Select extreme specimens (the ones that make the warp-meshes)
res.PC1.min.h.n <- wrap.PCA.r.n$pc.shapes[[1]] 
res.PC1.max.h.n <- wrap.PCA.r.n$pc.shapes[[2]]
```

### Export residual no Notomys heatmaps
```{r}
# Lateral view of mean to smallest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean, res.PC1.min.h.n, col = heat.colors, col.val = PC1.var.r.n[, "radius"])
view3d(phi = 90)
rgl.snapshot("../Data/Processed/heatmap_RES_sm-mean_lat_NOt.png")

# Lateral view of mean to biggest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean, res.PC1.max.h.n, col = heat.colors, col.val = PC1.var.r.n[, "radius"])
view3d(phi = 90)
rgl.snapshot("../Data/Processed/heatmap_RES_mean-big_lat_NOt.png")

# Dorsal view of mean to smallest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean, res.PC1.min.h.n, col = heat.colors, col.val = PC1.var.r.n[, "radius"]) 
view3d(phi = 0)  # flip cranium
rgl.snapshot("../Data/Processed/heatmap_RES_sm-mean_dor_NOt.png")

# Dorsal view of mean to biggest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean, res.PC1.max.h.n, col = heat.colors, col.val = PC1.var.r.n[, "radius"])
view3d(phi = 0)  # flip cranium
rgl.snapshot("../Data/Processed/heatmap_RES_mean-big_dor_NOt.png")
```

### Clean and load in all heatmaps
```{r}
# Remove whitespace from all heatmaps (requires ImageMagik installed on PC)
system("mogrify ../Data/Processed/heatmap_*.png -trim ../Data/Processed/*.png")

# Load in with allometry smallest to largest
indiv.sm.lat <- readPNG("../Data/Processed/heatmap_IND_sm-mean_lat.png")
indiv.big.lat <- readPNG("../Data/Processed/heatmap_IND_mean-big_lat.png")
indiv.sm.dor <- readPNG("../Data/Processed/heatmap_IND_sm-mean_dor.png")
indiv.big.dor <- readPNG("../Data/Processed/heatmap_IND_mean-big_dor.png")

# Load in with allometry smallest to largest
allom.sm.lat <- readPNG("../Data/Processed/heatmap_ALL_sm-mean_lat.png")
allom.big.lat <- readPNG("../Data/Processed/heatmap_ALL_mean-big_lat.png")
allom.sm.dor <- readPNG("../Data/Processed/heatmap_ALL_sm-mean_dor.png")
allom.big.dor <- readPNG("../Data/Processed/heatmap_ALL_mean-big_dor.png")

# Load in with allometry PC 1
allPC.sm.lat <- readPNG("../Data/Processed/heatmap_PCA_sm-mean_lat.png")
allPC.big.lat <- readPNG("../Data/Processed/heatmap_PCA_mean-big_lat.png")
allPC.sm.dor <- readPNG("../Data/Processed/heatmap_PCA_sm-mean_dor.png")
allPC.big.dor <- readPNG("../Data/Processed/heatmap_PCA_mean-big_dor.png")

# Load in without allometry residuals PC1
resid.sm.lat <- readPNG("../Data/Processed/heatmap_RES_sm-mean_lat.png")
resid.big.lat <- readPNG("../Data/Processed/heatmap_RES_mean-big_lat.png")
resid.sm.dor <- readPNG("../Data/Processed/heatmap_RES_sm-mean_dor.png")
resid.big.dor <- readPNG("../Data/Processed/heatmap_RES_mean-big_dor.png")

# Load in without allometry residuals PC2
resP2.sm.lat <- readPNG("../Data/Processed/heatmap_RP2_sm-mean_lat.png")
resP2.big.lat <- readPNG("../Data/Processed/heatmap_RP2_mean-big_lat.png")
resP2.sm.dor <- readPNG("../Data/Processed/heatmap_RP2_sm-mean_dor.png")
resP2.big.dor <- readPNG("../Data/Processed/heatmap_RP2_mean-big_dor.png")

# Load in without allometry residuals PC1 - no Notomys
resid.sm.lat.no <- readPNG("../Data/Processed/heatmap_RES_sm-mean_lat_NOt.png")
resid.big.lat.no <- readPNG("../Data/Processed/heatmap_RES_mean-big_lat_NOt.png")
resid.sm.dor.no <- readPNG("../Data/Processed/heatmap_RES_sm-mean_dor_NOt.png")
resid.big.dor.no <- readPNG("../Data/Processed/heatmap_RES_mean-big_dor_NOt.png")
```

## 2) Make multi-panel Figure 2, heatmaps with allometry
```{r}
# Set up muli-panel figure; lateral 
mat <- matrix(c(1:12), 4, 3, byrow = FALSE)  # 12 plots, 4 rows, 3 columns
layout(mat, widths = c(1, 1, 1), heights = c(0.65, 0.75, 0.5))  

## Individuals
# 1) Small to Mean, with allometry, lateral view
par(mar = c(0, 0, 0, 0))  # sets the margins
plot(c(0, dim(indiv.sm.lat)[2]), c(0, dim(indiv.sm.lat)[1]), type = "n", axes = F, xlab = "", ylab = "", asp = T) 
rasterImage(indiv.sm.lat, 0, 0, dim(indiv.sm.lat)[2], dim(indiv.sm.lat)[1])
text(0, dim(indiv.sm.lat)[1]-10, "A", cex = 2)
legend("topleft", inset = c(.01, 0), legend = "Smallest/Largest Individuals", bty = "n", cex = 1.2)

# 2) Mean to Big, with allometry, lateral view
par(mar = c(0, 0, -1, 0))
plot(c(0, dim(indiv.big.lat)[2]), c(0, dim(indiv.big.lat)[1]), type = "n", axes = F, xlab = "", ylab = "", asp = T) 
rasterImage(indiv.big.lat, 0, 0, dim(indiv.big.lat)[2], dim(indiv.big.lat)[1])
text(0, dim(indiv.big.lat)[1]-10, "B", cex = 2)

# 3) Small to Mean, with Allometry, dorsal view
par(mar = c(0, 0, 0, 0))  # sets the margins
plot(c(0, dim(indiv.big.dor)[2]), c(0, dim(indiv.big.dor)[1]), type = "n", axes = F, xlab = "", ylab = "", asp = T) 
rasterImage(indiv.sm.dor, 0, 0, dim(indiv.big.dor)[2], dim(indiv.big.dor)[1])
text(0, dim(indiv.big.dor)[1]-10, "C", cex = 2)

# 4) Mean to Big, with allometry, dorsal view
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(indiv.big.dor)[2]), c(0, dim(indiv.big.dor)[1]), type = "n", axes = F, xlab = "", ylab = "", asp = T) 
rasterImage(indiv.big.dor, 0, 0, dim(indiv.big.dor)[2], dim(indiv.big.dor)[1])
text(0, dim(indiv.big.dor)[1]-10, "D", cex = 2)

## Fitted allometry
# 5) Small to Mean, with allometry, lateral view
par(mar = c(0, 0, 0, 0))  # sets the margins
plot(c(0, dim(allom.sm.lat)[2]), c(0, dim(allom.sm.lat)[1]), type = "n", axes = F, xlab = "", ylab = "", asp = T) 
rasterImage(allom.sm.lat, 0, 0, dim(allom.sm.lat)[2], dim(allom.sm.lat)[1])
text(0, dim(allom.sm.lat)[1]-10, "E", cex = 2)
legend("topright", inset = c(0.22, 0), legend = "Fitted Allometry", bty = "n", cex = 1.2)

# 6) Mean to Big, with allometry, lateral view
par(mar = c(0, 0, -1, 0))
plot(c(0, dim(allom.big.lat)[2]), c(0, dim(allom.big.lat)[1]), type = "n", axes = F, xlab = "", ylab = "", asp = T) 
rasterImage(allom.big.lat, 0, 0, dim(allom.big.lat)[2], dim(allom.big.lat)[1])
text(0, dim(allom.big.lat)[1]-10, "F", cex = 2)

# 7) Small to Mean, with Allometry, dorsal view
par(mar = c(0, 0, 0, 0))  # sets the margins
plot(c(0, dim(allom.big.dor)[2]), c(0, dim(allom.big.dor)[1]), type = "n", axes = F, xlab = "", ylab = "", asp = T) 
rasterImage(allom.sm.dor, 0, 0, dim(allom.big.dor)[2], dim(allom.big.dor)[1])
text(0, dim(allom.big.dor)[1]-10, "G", cex = 2)

# 8) Mean to Big, with allometry, dorsal view
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(allom.big.dor)[2]), c(0, dim(allom.big.dor)[1]), type = "n", axes = F, xlab = "", ylab = "", asp = T) 
rasterImage(allom.big.dor, 0, 0, dim(allom.big.dor)[2], dim(allom.big.dor)[1])
text(0, dim(allom.big.dor)[1]-10, "H", cex = 2)

## PC1 
# 9) Small to Mean, PCA allometry, lateral view
par(mar = c(0, 0, 0, 0))  # sets the margins
plot(c(0, dim(allPC.sm.lat)[2]), c(0, dim(allPC.sm.lat)[1]), type = "n", axes = F, xlab = "", ylab = "", asp = T) 
rasterImage(allPC.sm.lat, 0, 0, dim(allPC.sm.lat)[2], dim(allPC.sm.lat)[1])
text(0, dim(allPC.sm.lat)[1]-10, "I", cex = 2)
legend("topright", inset = c(0.06, 0), legend = "Principal Component 1", bty = "n", cex = 1.2)

# 10) Mean to Big, PCA allometry, lateral view
par(mar = c(0, 0, -1, 0))
plot(c(0, dim(allPC.big.lat)[2]), c(0, dim(allPC.big.lat)[1]), type = "n", axes = F, xlab = "", ylab = "", asp = T) 
rasterImage(allPC.big.lat, 0, 0, dim(allPC.big.lat)[2], dim(allPC.big.lat)[1])
text(0, dim(allPC.big.lat)[1]-10, "J", cex = 2)

# 11) Small to Mean, PCA allometry, dorsal view
par(mar = c(0, 0, 0, 0))  # sets the margins
plot(c(0, dim(allPC.big.dor)[2]), c(0, dim(allPC.big.dor)[1]), type = "n", axes = F, xlab = "", ylab = "", asp = T) 
rasterImage(allPC.sm.dor, 0, 0, dim(allPC.big.dor)[2], dim(allPC.big.dor)[1])
text(0, dim(allPC.big.dor)[1]-10, "K", cex = 2)

# 12) Mean to Big, PCA allometry: dorsal view
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(allPC.big.dor)[2]), c(0, dim(allPC.big.dor)[1]), type = "n", axes = F, xlab = "", ylab = "", asp = T) 
rasterImage(allPC.big.dor, 0, 0, dim(allPC.big.dor)[2], dim(allPC.big.dor)[1])
text(0, dim(allPC.big.dor)[1]-10, "L", cex = 2)
```

## 3) Make multi-panel Figure 3: Heatmaps with Residuals
```{r}
# Set up muli-panel figure; lateral 
mat <- matrix(c(1:12), 4, 3, byrow = TRUE)  # 12 plots, 4 rows, 3 columns
layout(mat, widths = c(1, 1, 1), heights = c(0.65, 0.75, 0.5))  

## LATERAL VIEWS
# 1) Small to Mean, residuals PC1, lateral view
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resid.sm.lat)[2]), c(0, dim(resid.sm.lat)[1]), type = "n", axes = F, xlab = "", ylab = "", asp = T) 
rasterImage(resid.sm.lat, 0, 0, dim(resid.sm.lat)[2], dim(resid.sm.lat)[1])
text(0, dim(resid.sm.lat)[1]-10, "A", cex = 2)

# 2) Small to Mean, residuals PC1, lateral view - no Notomys
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resid.sm.lat.no)[2]), c(0, dim(resid.sm.lat.no)[1]), type = "n", axes = F, xlab = "", ylab = "", asp = T) 
rasterImage(resid.sm.lat.no, 0, 0, dim(resid.sm.lat.no)[2], dim(resid.sm.lat.no)[1])
text(0, dim(resid.sm.lat.no)[1]-10, "E", cex = 2)

# 3) Small to Mean, residuals PC2, lateral view
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resP2.sm.lat)[2]), c(0, dim(resP2.sm.lat)[1]), type = "n", axes = F, xlab = "", ylab = "", asp = T) 
rasterImage(resP2.sm.lat, 0, 0, dim(resP2.sm.lat)[2], dim(resP2.sm.lat)[1])
text(0, dim(resP2.sm.lat)[1]-10, "I", cex = 2)

# 4) Mean to Big, residuals PC1, lateral view
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resid.big.lat)[2]), c(0, dim(resid.big.lat)[1]), type = "n", axes = F, xlab = "", ylab = "", asp = T) 
rasterImage(resid.big.lat, 0, 0, dim(resid.big.lat)[2], dim(resid.big.lat)[1])
text(0, dim(resid.big.lat)[1]-10, "B", cex = 2)

# 5) Mean to Big, residuals PC1, lateral view - no Notomys
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resid.big.lat.no)[2]), c(0, dim(resid.big.lat.no)[1]), type = "n", axes = F, xlab = "", ylab = "", asp = T) 
rasterImage(resid.big.lat.no, 0, 0, dim(resid.big.lat.no)[2], dim(resid.big.lat.no)[1])
text(0, dim(resid.big.lat.no)[1]-10, "F", cex = 2)

# 6) Mean to Big, residuals PC2, lateral view
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resP2.big.lat)[2]), c(0, dim(resP2.big.lat)[1]), type = "n", axes = F, xlab = "", ylab = "", asp = T) 
rasterImage(resP2.big.lat, 0, 0, dim(resP2.big.lat)[2], dim(resP2.big.lat)[1])
text(0, dim(resP2.big.lat)[1]-10, "J", cex = 2)

## DORSAL VIEWS
# 7) Small to Mean, residuals PC1, dorsal view
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resid.big.dor)[2]), c(0, dim(resid.big.dor)[1]), type = "n", axes = F, xlab = "", ylab = "", asp = T) 
rasterImage(resid.sm.dor, 0, 0, dim(resid.big.dor)[2], dim(resid.big.dor)[1])
text(0, dim(resid.big.dor)[1]-10, "C", cex = 2)

# 8) Small to Mean, residuals PC1, dorsal view - no Notomys
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resid.big.dor.no)[2]), c(0, dim(resid.big.dor.no)[1]), type = "n", axes = F, xlab = "", ylab = "", asp = T) 
rasterImage(resid.sm.dor.no, 0, 0, dim(resid.big.dor.no)[2], dim(resid.big.dor.no)[1])
text(0, dim(resid.big.dor.no)[1]-10, "G", cex = 2)

# 9) Small to Mean, residuals PC2, dorsal view
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resP2.big.dor)[2]), c(0, dim(resP2.big.dor)[1]), type = "n", axes = F, xlab = "", ylab = "", asp = T) 
rasterImage(resP2.sm.dor, 0, 0, dim(resP2.big.dor)[2], dim(resP2.big.dor)[1])
text(0, dim(resP2.big.dor)[1]-10, "K", cex = 2)

# 10) Mean to Big, residuals PC1, dorsal view
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resid.big.dor)[2]), c(0, dim(resid.big.dor)[1]), type = "n", axes = F, xlab = "", ylab = "", asp = T) 
rasterImage(resid.big.dor, 0, 0, dim(resid.big.dor)[2], dim(resid.big.dor)[1])
text(0, dim(resid.big.dor)[1]-10, "D", cex = 2)

# 11) Mean to Big, residuals PC1, dorsal view - no Notomys
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resid.big.dor.no)[2]), c(0, dim(resid.big.dor.no)[1]), type = "n", axes = F, xlab = "", ylab = "", asp = T) 
rasterImage(resid.big.dor.no, 0, 0, dim(resid.big.dor.no)[2], dim(resid.big.dor.no)[1])
text(0, dim(resid.big.dor.no)[1]-10, "H", cex = 2)

# 12) Mean to Big, residuals PC2, dorsal view
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resP2.big.dor)[2]), c(0, dim(resP2.big.dor)[1]), type = "n", axes = F, xlab = "", ylab = "", asp = T) 
rasterImage(resP2.big.dor, 0, 0, dim(resP2.big.dor)[2], dim(resP2.big.dor)[1])
text(0, dim(resP2.big.dor)[1]-10, "L", cex = 2)
```