---
title: "06-heatmap-both-datasets"
author: "Ariel Marcy"
date: "9/18/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '../eco-rodents')
```

# Heatmaps both with-allometry and without-allometry 3D coordinates
1) Make heatmaps for with-allometry.

2) Make heatmaps for without-allometry aka residuals aka size-less shape.

3) Make multi-panel figure for comparisons across heatmaps.

4) Directly compare by subtracting residual change from allometry change.

### Load packages, functions, and data
Tree, mean shape, and metadata for phylogenetic analyses come from the allometry rodents script 04. 
```{r message = FALSE}
library(geomorph)
library(data.table)
library(vegan)
library(geiger)  # for phylogenies
library(png)  # needed to work with screenshots
library(colorspace)  # plot colors

# Load libraries (and install if necessary) for landvR package
if(!require(devtools)) install.packages("devtools")
if(!require(dispRity)) install.packages("dispRity")
if(!require(landvR)) install_github("TGuillerme/landvR")
source("../Functions/utilities.R")  # custom functions
source("../Functions/plotGMPhyloMorphoSpace_plotmod.R")  # modified geomorph function
load(file = "../Data/Processed/03-main-data.rda")
load(file = "../Data/Processed/03-tree-data.rda")
load(file = "../Data/Processed/03-graphing-vectors.rda")
load(file = "../Data/Processed/05-residual-data.rda")
```

## 1) Heatmaps for fitted allometry coordinates
First find mean speciemen, then smallest and largest. Uses Dr Thomas Guillerme's new package, `landvR` to see how landmarks vary between these specimens.
```{r}
# Fit specimens
allom.fit <- procD.lm(mean.shapes~info.means$MeanCsize)
fit.shapes <- arrayspecs(allom.fit$fitted, 325, 3)

# Find mean configuration - i.e. mean shape in dataset
allom.mean <- (select.procrustes(fit.shapes, selector = mean))[[1]]

# Measure spherical coordinates differences from the mean for each specimen for each landmark
allom.diff.mean <- coordinates.difference(coordinates = fit.shapes, reference = allom.mean, type = "spherical")

# Find largest and smallest native species from mean centroid size
big.index <- which(info.means$MeanCsize == max(info.means$MeanCsize))
biggest.shape <- fit.shapes[, , big.index]
sm.index <- which(info.means$Species == "del")
smallest.shape <- fit.shapes[, , sm.index]
```

### Export the variation for with allometry shape change
Export images for plots
```{r}
# Lateral view of smallest to mean
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(allom.mean, smallest.shape, col = heat.colors, col.val = allom.diff.mean$Pseudomys_delicatulus[, "radius"])
view3d(phi = 90)
rgl.snapshot("../Data/Processed/heatmap_ALL_sm-mean_lat.png")

# Lateral view of mean to biggest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(allom.mean, biggest.shape, col = heat.colors, col.val = allom.diff.mean$Uromys_caudimaculatus[, "radius"])
view3d(phi = 90)
rgl.snapshot("../Data/Processed/heatmap_ALL_mean-big_lat.png")

# Dorsal view of smallest to mean
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(allom.mean, smallest.shape, col = heat.colors, col.val = allom.diff.mean$Pseudomys_delicatulus[, "radius"]) 
view3d(phi = 0)  # flip cranium
rgl.snapshot("../Data/Processed/heatmap_ALL_sm-mean_dor.png")

# Dorsal view of mean to biggest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(allom.mean, biggest.shape, col = heat.colors, col.val = allom.diff.mean$Uromys_caudimaculatus[, "radius"])
view3d(phi = 0)  # flip cranium
rgl.snapshot("../Data/Processed/heatmap_ALL_mean-big_dor.png")
```

## 2) Heatmaps for PC1 with allometry


## 3) Heatmaps for PC1 with residual coordinates
Steps as above.
```{r}
# Find mean configuration - i.e. mean shape in dataset
res.mean <- (select.procrustes(residuals, selector = mean))[[1]]

# Measure spherical coordinates differences from the mean for each specimen for each landmark
res.diff.mean <- coordinates.difference(coordinates = residuals, reference = res.mean, type = "spherical")

# Find smallest and largest residual shapes
res.sm.shape <- residuals[, , sm.index]
res.big.shape <- residuals[, , big.index]
```

### Export the variation for residual shape change
Export images for plots.
```{r}
# Lateral view of mean to smallest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean, res.sm.shape, col = heat.colors, col.val = res.diff.mean$Pseudomys_delicatulus[, "radius"])
view3d(phi = 90)
rgl.snapshot("../Data/Processed/heatmap_RES_sm-mean_lat.png")

# Lateral view of mean to biggest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean, res.big.shape, col = heat.colors, col.val = res.diff.mean$Uromys_caudimaculatus[, "radius"])
view3d(phi = 90)
rgl.snapshot("../Data/Processed/heatmap_RES_mean-big_lat.png")

# Dorsal view of mean to smallest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean, res.sm.shape, col = heat.colors, col.val = res.diff.mean$Pseudomys_delicatulus[, "radius"]) 
view3d(phi = 0)  # flip cranium
rgl.snapshot("../Data/Processed/heatmap_RES_sm-mean_dor.png")

# Dorsal view of mean to biggest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean, res.big.shape, col = heat.colors, col.val = res.diff.mean$Uromys_caudimaculatus[, "radius"])
view3d(phi = 0)  # flip cranium
rgl.snapshot("../Data/Processed/heatmap_RES_mean-big_dor.png")
```

### 4) Make multipanel figure comparing shape change between smallest-to-mean and mean-to-largest, with allometry vs without (residuals)
First, clean up and load in all the images just created
```{r}
# Remove whitespace from all heatmaps (requires ImageMagik installed on PC)
system("mogrify ../Data/Processed/heatmap_*.png -trim ../Data/Processed/*.png")

# Load in with allometry 
allom.sm.lat <- readPNG("../Data/Processed/heatmap_ALL_sm-mean_lat.png")
allom.big.lat <- readPNG("../Data/Processed/heatmap_ALL_mean-big_lat.png")
allom.sm.dor <- readPNG("../Data/Processed/heatmap_ALL_sm-mean_dor.png")
allom.big.dor <- readPNG("../Data/Processed/heatmap_ALL_mean-big_dor.png")

# Load in without allometry residuals
resid.sm.lat <- readPNG("../Data/Processed/heatmap_RES_sm-mean_lat.png")
resid.big.lat <- readPNG("../Data/Processed/heatmap_RES_mean-big_lat.png")
resid.sm.dor <- readPNG("../Data/Processed/heatmap_RES_sm-mean_dor.png")
resid.big.dor <- readPNG("../Data/Processed/heatmap_RES_mean-big_dor.png")
```

## Make multi-panel Figure TK
```{r}
# Set up muli-panel figure; lateral 
mat <- matrix(c(1, 2, 3, 4, 5, 6, 7, 8), 4, 2, byrow = TRUE)  # 4 rows, 2 columns
layout(mat, widths = c(1, 1, 1), heights = c(0.65, 0.75, 0.5))  

## LATERAL VIEWS
# 1) Small to Mean, with Allometry, lateral view
par(mar = c(0, 0, 0, 0))  # sets the margins
plot(c(0, dim(allom.sm.lat)[2]), c(0, dim(allom.sm.lat)[1]), type = "n", axes = F, xlab = "", ylab = "", asp = T) 
rasterImage(allom.sm.lat, 0, 0, dim(allom.sm.lat)[2], dim(allom.sm.lat)[1])
text(0, dim(allom.sm.lat)[1]-10, "A", cex = 2)

# 2) Small to Mean, without Allometry, lateral view
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resid.sm.lat)[2]), c(0, dim(resid.sm.lat)[1]), type = "n", axes = F, xlab = "", ylab = "", asp = T) 
rasterImage(resid.sm.lat, 0, 0, dim(resid.sm.lat)[2], dim(resid.sm.lat)[1])
text(0, dim(resid.sm.lat)[1]-10, "B", cex = 2)

# 3) Mean to Big, with Allometry, lateral view
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(allom.big.lat)[2]), c(0, dim(allom.big.lat)[1]), type = "n", axes = F, xlab = "", ylab = "", asp = T) 
rasterImage(allom.big.lat, 0, 0, dim(allom.big.lat)[2], dim(allom.big.lat)[1])
text(0, dim(allom.big.lat)[1]-10, "C", cex = 2)

# 4) Mean to Big, without Allometry, lateral view
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resid.big.lat)[2]), c(0, dim(resid.big.lat)[1]), type = "n", axes = F, xlab = "", ylab = "", asp = T) 
rasterImage(resid.big.lat, 0, 0, dim(resid.big.lat)[2], dim(resid.big.lat)[1])
text(0, dim(resid.big.lat)[1]-10, "D", cex = 2)

## DORSAL VIEWS
# 5) Small to Mean, with Allometry, dorsal view
par(mar = c(0, 0, 0, 0))  # sets the margins
plot(c(0, dim(allom.big.dor)[2]), c(0, dim(allom.big.dor)[1]), type = "n", axes = F, xlab = "", ylab = "", asp = T) 
rasterImage(allom.sm.dor, 0, 0, dim(allom.big.dor)[2], dim(allom.big.dor)[1])
text(0, dim(allom.big.dor)[1]-10, "E", cex = 2)

# 6) Small to Mean, without Allometry, dorsal view
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resid.big.dor)[2]), c(0, dim(resid.big.dor)[1]), type = "n", axes = F, xlab = "", ylab = "", asp = T) 
rasterImage(resid.sm.dor, 0, 0, dim(resid.big.dor)[2], dim(resid.big.dor)[1])
text(0, dim(resid.big.dor)[1]-10, "F", cex = 2)

# 7) Mean to Big, with Allometry, dorsal view
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(allom.big.dor)[2]), c(0, dim(allom.big.dor)[1]), type = "n", axes = F, xlab = "", ylab = "", asp = T) 
rasterImage(allom.big.dor, 0, 0, dim(allom.big.dor)[2], dim(allom.big.dor)[1])
text(0, dim(allom.big.dor)[1]-10, "G", cex = 2)

# 8) Mean to Big, without Allometry: dorsal view
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resid.big.dor)[2]), c(0, dim(resid.big.dor)[1]), type = "n", axes = F, xlab = "", ylab = "", asp = T) 
rasterImage(resid.big.dor, 0, 0, dim(resid.big.dor)[2], dim(resid.big.dor)[1])
text(0, dim(resid.big.dor)[1]-10, "H", cex = 2)
```

## Heatmap species of interest
Export images for plots
```{r}
notomys <- mean.shapes[, , which(info.means$Taxa == "N_fus")]

# Lateral view of smallest to mean
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(allom.mean, notomys, col = heat.colors, col.val = allom.diff.mean$Notomys_fuscus[, "radius"])
view3d(phi = 90)
rgl.snapshot("../Data/Processed/heatmap_ALL_not-mean_lat.png")

# Dorsal view of smallest to mean
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(allom.mean, notomys, col = heat.colors, col.val = allom.diff.mean$Notomys_fuscus[, "radius"]) 
view3d(phi = 0)  # flip cranium
rgl.snapshot("../Data/Processed/heatmap_ALL_no-mean_dor.png")

```

```{r}
# Find mean configuration - i.e. mean shape in dataset
allom.mean <- (select.procrustes(mean.shapes, selector = mean))[[1]]

# Measure spherical coordinates differences from the mean for each specimen for each landmark
allom.diff.mean <- coordinates.difference(coordinates = mean.shapes, reference = allom.mean, type = "spherical")
