---
title: "07-plot-phylomorph-dist"
author: "Ariel Marcy"
date: "1/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '../eco-rodents')
```

# Make phylo-morphological distance plots
Assess limits of morphological distance (aka divergence) as it relates to phylogenetic distance. Calculated for full shape dataset (with allometry) and for shape residual dataset (allometry-free).

Creates bipanel Figure 5.

### Load packages, functions, and data from previous steps
```{r message = FALSE}
library(geomorph)
library(stringr)
library(picante)  # needed for distance matrix
source("../Functions/utilities.R")  # custom functions
load(file = "../Data/Processed/03-main-data.rda")
load(file = "../Data/Processed/03-tree-data.rda")
load(file = "../Data/Processed/03-graphing-vectors.rda")
load(file = "../Data/Processed/04-residual-data.rda")
```

## Analyze phylogenetic and morphological distance matrices
First, generate distances matrices that can be comparable.
```{r}
# Phylo distance matrix
phylo.matrix <- cophenetic(aus.tree) / 2  # divide by two to convert distance to millions of years since Last Common Ancestor

# Morphological distance matrix
mean.gpa <- gpagen(mean.shapes, ProcD = TRUE)  # use mean shapes of species
#morph.matrix <- as.matrix(mean.gpa$procD)  # ProcD = Procrustes distance
#mean.gpa$procD isn't created on some computers because of memory allocation issues. As per the instructions in geomorph, instead creating the distance matrix using the dist function. Note that this requires taking the coordinates into a 2D array, not mentioned in instructions. 
morph.matrix <- dist(two.d.array(mean.gpa$coords), method="euclidean")

morph.matrix <- as.matrix(morph.matrix)

# Make names shorter
long.names <- colnames(phylo.matrix)
short.names <- NULL 
for(i in 1:length(long.names)) {
        sep <- strsplit(long.names[i], "_")
        new.name <- paste(str_sub(sep[[1]][1], 1, 3), str_sub(sep[[1]][2], 1, 3), sep = "_")
        short.names <- c(short.names, new.name)
}

# Replace with short names
row.names(phylo.matrix) <- short.names  # phylo matrix
colnames(phylo.matrix) <- short.names
row.names(morph.matrix) <- short.names  # morph matrix
colnames(morph.matrix) <- short.names
```

### Turn distance matrices into dataframes and then combine
Write function that creates the same, comparable table structure taking in the distance matrices from above, and the name for the distance in the matrix.
```{r}
# Function takes matrix from above and user-provided label for dist. measured
make.table <- function(matrix, d.name) {
        new.table <- as.data.frame(as.table(matrix))  # matrix --> dataframe
        new.table[, 4] <- NA  # initialize pair name column
        colnames(new.table) <- c("spp1", "spp2", d.name, "pair")
        
        # Remove same-same comparisons (where distance = 0)
        new.table <- new.table[which(new.table[, d.name] > 0), ]
        
        # Remove permutations by alphabetizing the species names for each pair
        for(i in 1:dim(new.table)[1]) {
                pair.names <- c(as.character(new.table[i, "spp1"]), as.character(new.table[i, "spp2"]))
                pair.index <- str_order(c(as.character(new.table[i, "spp1"]), as.character(new.table[i, "spp2"])))
                pair <- paste(pair.names[pair.index[1]], pair.names[pair.index[2]], sep = "-")  # pair names in alpha order
                new.table[i, "pair"] <- pair
        }
        
        new.table <- new.table[!duplicated(new.table[, "pair"]), ]  # only unique pair combinations
        return(new.table)
}

# Run on both matrices, then merge by pair
phylo.table <- make.table(phylo.matrix, "phylo.d")
morph.table <- make.table(morph.matrix, "morph.d")
master.table <- merge(phylo.table, morph.table[, 3:4], by = "pair")
```

## Repeat Phylo/Morph distance plot using shape residuals
```{r}
# Morphological distance matrix for residuals
mean.gpa.res <- gpagen(residuals, ProcD = TRUE)

#again here separately creating the distance matrix
#morph.matrix.res <- as.matrix(mean.gpa.res$procD)

morph.matrix.res <- dist(two.d.array(mean.gpa.res$coords), method="euclidean")

morph.matrix.res <- as.matrix(morph.matrix.res)


# Shorten row and column names
row.names(morph.matrix.res) <- short.names
colnames(morph.matrix.res) <- short.names

# Make comparable table to phylo distance using function defined above
morph.table.r <- make.table(morph.matrix.res, "morph.d.r")
master.table.r <- merge(phylo.table, morph.table.r[, 3:4], by = "pair")
```

## Plots for phylo-morph distance pairs
First, set up colors
```{r}
# Create color key
just.genera <- NULL
for(i in 1:length(short.names)){
        sep <- strsplit(short.names[i], "_")
        just.genera <- c(just.genera, sep[[1]][1])
}
gen.col <- col.phylo[match(just.genera, names(col.phylo))]
names(gen.col)[5] <- "Mas"
gen.col[5] <- gen.col[16]  # Fix Mastacomys color

# Points coded for each individual specimen
alpha.col.spp1 <- gen.col[2:37]  # subset because C pen isn't in this col
alpha.col.spp2 <- gen.col

# Save in table for ease in plotting later on
master.table$col.spp1 <- PlotByGroup(master.table, "spp1", alpha.col.spp1)
master.table$col.spp2 <- PlotByGroup(master.table, "spp2", alpha.col.spp2)
```

### Export Figure 5
```{r}
genera.phylo.full <- c("Mus", "Rattus", "Pogonomys", "Xeromys", "Hydromys", "Notomys", "Pseudomys", "Zyzomys", "Leggadina", "Uromys", "Melomys", "Leporillus", "Mesembriomys", "Conilurus")

# Set up multi-panel figure for export
setEPS()
postscript("../Data/Results/Figure5_PhyloMorph_Dist.eps")
mat <- matrix(c(1, 2), 2, 1, byrow = TRUE)  # 2 rows, 1 column
layout(mat, widths = rep.int(1, ncol(mat)), heights = rep.int(1, nrow(mat))) 

# A) Plot with allometry
par(mar = c(1, 4, 3, 1))
plot(master.table$phylo.d, master.table$morph.d, col = master.table$col.spp1, bg = master.table$col.spp2, pch = 21, xaxt = 'n', ylim = c(0.01, 0.205), ylab = "Morph. distance")
axis(1, labels = FALSE)
text(0.25, 0.197, "a", cex = 1.5)
legend("topleft", legend = "full shape", inset = c(0.004, 0.01), bty = "n")

# B) Plot with residuals
par(mar = c(4.25, 4, 0, 1))
plot(master.table.r$phylo.d, master.table.r$morph.d.r, col = master.table$col.spp1, bg = master.table$col.spp2, pch = 21, ylim = c(0.01, 0.205), xlab = "", ylab = "Morph. distance")
title(xlab = "Phylogenetic distance (~million years since common ancestor)", line = 2.3)
text(0.25, 0.197, "b", cex = 1.5)
legend("topleft", legend = "shape residual", inset = c(0.004, 0.01), bty = "n")
legend(4, 0.21, legend = genera.phylo.full, text.font = 3, col = col.phylo, pch = 16, cex = 0.65, pt.cex = 0.7, ncol = 5, x.intersp = 0.8, text.width = 1.2, bty = "n", inset = c(0.12, 0))

dev.off()
```