---
title: "07-test-LM-region-differences"
author: "Ariel Marcy"
date: "10/21/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '../eco-rodents')
```

# Test variation within morphospace
What is causing high levels of allometry? Is it functional integration of the gnawing apparatus (mostly in the top cranium-rostrum area), for example? 

1) Tests modularity adjusting for phylogenetic signal using CR coefficients

2) Tests modularity with Mantel tests of PC scores distance matrices from each module

3) Tests for global integration

Makes Figure 4, Table 1, and Supplementary Figure TK

### Load packages, functions, and data from previous steps
```{r message = FALSE}
library(png)  # needed to work with screenshots
library(wesanderson)  # plot colors
library(colorspace)  # darken colors
library(vegan)  # for Mantel tests
library(geomorph)
library(data.table)
source("../Functions/utilities.R")  # custom functions
load(file = "../Data/Processed/03-main-data.rda")
load(file = "../Data/Processed/03-tree-data.rda")
load(file = "../Data/Processed/03-graphing-vectors.rda")
load(file = "../Data/Processed/05-residual-data.rda")

# Can load in time consuming modularity test after first run:
load(file = "../Data/Processed/07-modularity.rda")
```

## Choose the modules
The five modules will need to be designated from the LM numbers. The first time, this was done with `define.modules()` but each module was then saved in a .csv file so this step can be repeated much more quickly.

### Load in the partition definitions
```{r}
lm.modules <- read.csv("../Data/Processed/Landmark_modules.csv", sep = ",", header = TRUE)
```

### Visualize the landmark types
```{r}
# Deconstruct function from Thomas's code in Landmark-test
ref <- mshape(mean.shapes)  
lm.levels <- unique(lm.modules[, 5])
lm.colors <- c("blue", "gray", "black")  # semi, patch, fixed

# Subset the landmarks according to the partitions
lm.parts <- list()
for(i in 1:length(lm.levels)){
        lm.parts[[i]] <- which(lm.modules[, 5] == lm.levels[[i]])
}

# 3D plot of the partitions
open3d()
for (i in 1:length(lm.levels)){
        spheres3d(ref[lm.parts[[i]], 1], ref[lm.parts[[i]], 2], ref[lm.parts[[i]], 3], col = lm.colors[i], lit = TRUE, radius = 0.001, asp = FALSE)
}
```

### Visualize the modules from Goswami 2006/7
6 modules hypothesized for mammals, our dataset supports 5 (missing zygomatic)
```{r}
# Deconstruct function from Thomas's code in Landmark-test
ref <- mshape(mean.shapes)
mod.levels<- sort(unique(lm.modules[, 4]))  # 5 (of 6) modules from Goswami

# Define nice colors for modules
mod.colors <- c(wes_palette("GrandBudapest1")[2:4], wes_palette("GrandBudapest2")[1:2])

# Subset the landmarks according to the modules
mod.parts <- list()
for(i in 1:length(mod.levels)){
  mod.parts[[i]] <- which(lm.modules[, 4] == mod.levels[[i]])
}
```

### Make image of 3D plot of the modules for Fig. 4A
```{r}
open3d(windowRect = c(0, 0, 1200, 1200))
for(i in 1:length(mod.levels)) {
        spheres3d(ref[mod.parts[[i]], 1], ref[mod.parts[[i]], 2], ref[mod.parts[[i]], 3], col = mod.colors[i], lit = TRUE, radius = 0.0015, asp = FALSE)
        #rgl.texts(ref[mod.parts[[i]], 1], ref[mod.parts[[i]], 2], ref[mod.parts[[i]], 3], mod.parts[[i]], offset = 100, color = c("black"))
}
view3d(phi = 90)  # Lateral view, dorsal side up
rgl.snapshot("../Data/Processed/modules.png")

# Remove whitespace from new image (requires ImageMagik installed on PC)
system("mogrify ../Data/Processed/modules.png -trim ../Data/Processed/*.png")

# Load png back in
module.plot <- readPNG("../Data/Processed/modules.png")
```

## Test modularity with 5 modules for full shape dataset
This test uses covariance ratio (CR) coefficients (Adams 2016) to quantify the degree of modularity between hypothesized modules/partitions. The observed CR is compared to the distribution of randomized CR's, found by randomizing the landmarks into partitions of identical sample sizes. Accounts for random changes due to evolution via Brownian motion (Felice and Adams 2014).
```{r}
# Test with the phylogentic tree (and mean shapes)
mod.test.5.phy <- phylo.modularity(mean.shapes, partition.gp = lm.modules[, 4], aus.tree)

# Make pairwise comparison table
mod.result.phy <- as.matrix(mod.test.5.phy$CR.mat)  # results
colnames(mod.result.phy) <- unique(lm.modules[, 4])  # names from 1st instance
rownames(mod.result.phy) <- unique(lm.modules[, 4])  # ...in lm.module[, 4]
alpha.key <- match(unique(lm.modules[, 4]), mod.levels)  # alphabetize table
mod.result.phy <- mod.result.phy[order(alpha.key), order(alpha.key)]  # alphabetize table
```

## Test modularity with 5 modules for shape residual dataset -- with phylogeny
As above but with residuals
```{r}
# Test with the phylogentic tree (and mean shapes)
mod.test.5.res <- phylo.modularity(residuals, partition.gp = lm.modules[, 4], aus.tree)

# Make pairwise comparison table
mod.result.res <- as.matrix(mod.test.5.res$CR.mat)  # results
colnames(mod.result.res) <- unique(lm.modules[, 4])  # names from 1st instance
rownames(mod.result.res) <- unique(lm.modules[, 4])  # ...in lm.module[, 4]
mod.result.res <- mod.result.res[order(alpha.key), order(alpha.key)]

# Save time consuming tests
save(mod.test.5.phy, mod.test.5.res, mod.result.phy, mod.result.res, file = "../Data/Processed/07-modularity.rda")
```

## Make and export Figure 4
```{r}
# Set up for export and for multiple panels
pdf("../Data/Results/Figure4_Modularity.pdf")
mat <- matrix(c(1, 1, 2, 3), 2, 2, byrow = TRUE)  # 3 plots, 2 row, 2 columns
layout(mat, widths = rep.int(1, ncol(mat)), heights = rep.int(1, nrow(mat)))  

# 1) Plot of modules
par(mar = c(1, 1, 1, 1))  # sets the margins
plot(c(0, dim(module.plot)[2]), c(0, dim(module.plot)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE)  # module.plot = png made above
rasterImage(module.plot, 0, 0, dim(module.plot)[2], dim(module.plot)[1])
# Labels for module plot
legend("topleft", "A", cex = 1.5, bty = "n")
legend("topleft", "module definitions", cex = 0.9, bty = "n", inset = c(0.04, 0.015))
legend("topleft", "(observed condition)", cex = 0.9, bty = "n", inset = c(0.038, 0.065))
text(80, -3, "Basicranium", cex = 0.9, bty = "n", col = darken(mod.colors[1]))
text(500, 28, "Molar", cex = 0.9, bty = "n", col = mod.colors[2])
text(560, 349, "Orbital", cex = 0.9, bty = "n", col = darken(mod.colors[3]))
text(835, 270, "Rostrum", cex = 0.9, bty = "n", col = darken(mod.colors[4], amount = 0.3))
text(-30, 220, "Vault", cex = 0.9, bty = "n", col = darken(mod.colors[5], amount = 0.35))

# 2) Plot of observed and resampled CR coefficients (w/ phylo correction)
par(mar = c(3.5, 3.5, 1.5, 0.5))
hist(mod.test.5.phy$random.CR,
     main = NULL,
     xlab = "",
     xlim = c(0.65, 1.1),
     ylab = "",
     cex.axis = 0.9,
     col = "black")
# Labels for histogram
title(xlab = "Covariance ratio (CR) coefficient", line = 2.4)
title(ylab = "Frequency (out of 1000)", line = 2.35)
arrows(x0 = mod.test.5.phy$CR, y0 = 175, y1 = 8, length = 0.15, lwd = 1.5, col = "darkred")
text(mod.test.5.phy$CR - 0.03, 200, "observed CR", cex = 0.9, col = "darkred")
legend("topleft", "B", cex = 1.5, bty = "n", inset = -0.05)
legend("topleft", "full shape", cex = 0.9, bty = "n", inset = c(0.033, 0.02))

# 3) Plot of CR coefficients (w/ phylo correction) for shape residual dataset
par(mar = c(3.5, 3.5, 1.5, 0.5))
hist(mod.test.5.res$random.CR,
     main = NULL,
     xlab = "",
     xlim = c(0.65, 1.1),
     ylab = "",
     cex.axis = 0.9,
     col = "black")
# Labels for histogram
title(xlab = "Covariance ratio (CR) coefficient", line = 2.4)
title(ylab = "Frequency (out of 1000)", line = 2.35)
arrows(x0 = mod.test.5.res$CR - 0.01, y0 = 105, y1 = 8, length = 0.15, lwd = 1.5, col = "darkred")
text(mod.test.5.res$CR - 0.01, 120, "observed CR", cex = 0.9, col = "darkred")
legend("topleft", "C", cex = 1.5, bty = "n", inset = -0.05)
legend("topleft", "residual shape", cex = 0.9, bty = "n", inset = c(0.033, 0.02))

dev.off()
```

## Mantel Test to see if PCs of modules are equivalent
Uses method similar to workflow from Sherratt et al. TK
```{r}
# Define (admittably brute force) function to run Mantel test on 5 modules named as in Figure 4A. Takes in a shape dataset and landmark definitions of modules made earlier in the script.
MantelModularity <- function(shapes, lm.modules) { 
  # Separate modules - using 5-module framework 
  mod.1 <- shapes[which(lm.modules[, 4] == "basicran"), ,]
  mod.2 <- shapes[which(lm.modules[, 4] == "molar"), ,]
  mod.3 <- shapes[which(lm.modules[, 4] == "orbital"), ,]
  mod.4 <- shapes[which(lm.modules[, 4] == "rostrum"), ,]
  mod.5 <- shapes[which(lm.modules[, 4] == "vault"), ,]
  
  # Run PCAs
  PCA.1 <- plotTangentSpace(mod.1)
  PCA.2 <- plotTangentSpace(mod.2)
  PCA.3 <- plotTangentSpace(mod.3)
  PCA.4 <- plotTangentSpace(mod.4)
  PCA.5 <- plotTangentSpace(mod.5)

  # Make distance matrices
  d.PCA.1 <- dist(PCA.1$pc.scores)
  d.PCA.2 <- dist(PCA.2$pc.scores)
  d.PCA.3 <- dist(PCA.3$pc.scores)
  d.PCA.4 <- dist(PCA.4$pc.scores)
  d.PCA.5 <- dist(PCA.5$pc.scores)
  
  # Run Mantel tests
  m.1v2 <- mantel(d.PCA.1, d.PCA.2, method = "pearson", permutations = 1000) 
  m.1v3 <- mantel(d.PCA.1, d.PCA.3, method = "pearson", permutations = 1000)
  m.1v4 <- mantel(d.PCA.1, d.PCA.4, method = "pearson", permutations = 1000)
  m.1v5 <- mantel(d.PCA.1, d.PCA.5, method = "pearson", permutations = 1000)

  m.2v3 <- mantel(d.PCA.2, d.PCA.3, method = "pearson", permutations = 1000) 
  m.2v4 <- mantel(d.PCA.2, d.PCA.4, method = "pearson", permutations = 1000)
  m.2v5 <- mantel(d.PCA.2, d.PCA.5, method = "pearson", permutations = 1000)

  m.3v4 <- mantel(d.PCA.3, d.PCA.4, method = "pearson", permutations = 1000)
  m.3v5 <- mantel(d.PCA.3, d.PCA.5, method = "pearson", permutations = 1000)

  m.4v5 <- mantel(d.PCA.4, d.PCA.5, method = "pearson", permutations = 1000)
  
  # Bonferroni correction
  p.vals.bon <- round(p.adjust(c(m.1v2$signif, m.1v3$signif, m.1v4$signif, m.1v5$signif, m.2v3$signif, m.2v4$signif, m.2v5$signif, m.3v4$signif, m.3v5$signif, m.4v5$signif), method = "bonferroni"), 3)
  
  # Make results table columns
  col1 <- c("-", p.vals.bon[1], p.vals.bon[2], p.vals.bon[3], p.vals.bon[4])
  col2 <- c(round(m.1v2$statistic, 3), "-", p.vals.bon[5], p.vals.bon[6], p.vals.bon[7])
  col3 <- c(round(m.1v3$statistic, 3), round(m.2v3$statistic, 3), "-", p.vals.bon[8], p.vals.bon[9])
  col4 <- c(round(m.1v4$statistic, 3), round(m.2v4$statistic, 3), round(m.3v4$statistic, 3), "-", p.vals.bon[10])
  col5 <- c(round(m.1v5$statistic, 3), round(m.2v5$statistic, 3), round(m.3v5$statistic, 3), round(m.4v5$statistic, 3), "-")

  # Combine columns into table
  modules <- sort(unique(as.factor(lm.modules[, 4])))
  mantel.results <- data.table(Modules = modules, basicranium = col1, molar = col2, orbital = col3, rostrum = col4, vault = col5)
  
  return(mantel.results)
  }

# Run on the full and shape residual datasets respectively
mantel.full <- MantelModularity(mean.shapes, lm.modules)
mantel.res <- MantelModularity(residuals, lm.modules)

# Add landmark numbers; col #4 gives the Goswami module definitions
n.num <- c(sum(lm.modules[, 4] == "basicran"), sum(lm.modules[, 4] == "molar"), sum(lm.modules[, 4] == "orbital"), sum(lm.modules[, 4] == "rostrum"), sum(lm.modules[, 4] == "vault"))
mantel.full <- cbind(mantel.full, n.num)  # add to tables for export
mantel.res <- cbind(mantel.res, n.num)

# Export Table 1A and 1B
write.table(mantel.full, "../Data/Results/Table1A_MantelTest_Full.csv", sep = ",", col.names = NA)
write.table(mantel.res, "../Data/Results/Table1B_MantelTest_Res.csv", sep = ",", col.names = NA)
```

### Whole integration test - Supplementary Figure TK
Test whether rodent skull is integrated as a whole
```{r}
# Set up for export
pdf("../Data/Results/SFigure_GlobalIntegration.pdf")
globalIntegration(mean.shapes, ShowPlot = TRUE)
text(12, -16, "slope of -1", col = "red")
dev.off()
```

```{r}
globalIntegration(residuals, ShowPlot = TRUE)


  mod.1 <- mean.shapes[which(lm.modules[, 4] == "basicran"), ,]
  mod.2 <- mean.shapes[which(lm.modules[, 4] == "molar"), ,]
  mod.3 <- mean.shapes[which(lm.modules[, 4] == "orbital"), ,]
  mod.4 <- mean.shapes[which(lm.modules[, 4] == "rostrum"), ,]
  mod.5 <- mean.shapes[which(lm.modules[, 4] == "vault"), ,]

  
  physignal(mod.1, aus.tree)
  physignal(mod.2, aus.tree)
  physignal(mod.3, aus.tree)
  physignal(mod.4, aus.tree)
  physignal(mod.5, aus.tree)
```