---
title: "06-test-modularity"
author: "Ariel Marcy"
date: "10/21/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = '../eco-rodents')
```

# Test variation within morphospace
What is causing high levels of allometry? Is it functional integration of the gnawing apparatus (mostly in the top cranium-rostrum area), for example? 

1) Tests modularity adjusting for phylogenetic signal using CR coefficients

2) Tests modularity with Mantel tests of PC scores distance matrices from each module

3) Tests for global integration

Makes Figure 4, Table 1, and Supplementary Figure 3

### Load packages, functions, and data from previous steps
```{r message = FALSE}
library(png)         # CRAN v0.1-8  
library(wesanderson) # CRAN v0.3.6
library(colorspace)  # CRAN v2.1-0  
library(vegan)       # CRAN v2.6-4  
library(geomorph)    # CRAN v4.0.6
library(data.table)  # CRAN v1.14.8
library(magick)      # CRAN v2.8.0

# This sets the wd to local
library(rstudioapi) # Safely Access the RStudio API
setwd(dirname(getActiveDocumentContext()$path))

source("../Functions/utilities.R")  # custom functions
load(file = "../Data/Processed/03-main-data.rda")
load(file = "../Data/Processed/03-tree-data.rda")
load(file = "../Data/Processed/03-graphing-vectors.rda")
load(file = "../Data/Processed/allometry_residuals.rda")

# Can load in time consuming modularity tests after first run:
#load(file = "../Data/Processed/06-modularity.rda")
```

## Choose the modules
The five modules will need to be designated from the LM numbers. The first time, this was done with `define.modules()` but each module was then saved in a .csv file so this step can be repeated much more quickly.

### Load in the partition definitions and numbers
```{r}
lm.modules <- read.csv("../Data/Processed/Landmark_modules.csv", sep = ",", header = TRUE)

# Add landmark numbers; col #4 gives the Goswami module definitions
 n.num <- c(sum(lm.modules[, 4] == "basicran"), sum(lm.modules[, 4] == "molar"), sum(lm.modules[, 4] == "orbital"), sum(lm.modules[, 4] == "rostrum"), sum(lm.modules[, 4] == "vault"))
```

### Visualize the landmark types
```{r}
# Deconstruct function from Thomas's code in Landmark-test
ref <- mshape(mean.shapes)  
lm.levels <- unique(lm.modules[, 5])
lm.colors <- c("blue", "gray", "black")  # semi, patch, fixed

# Subset the landmarks according to the partitions
lm.parts <- list()
for(i in 1:length(lm.levels)){
        lm.parts[[i]] <- which(lm.modules[, 5] == lm.levels[[i]])
}

# 3D plot of the partitions
open3d()
for (i in 1:length(lm.levels)){
        spheres3d(ref[lm.parts[[i]], 1], ref[lm.parts[[i]], 2], ref[lm.parts[[i]], 3], col = lm.colors[i], lit = TRUE, radius = 0.001, asp = FALSE)
}
```

### Visualize the modules from Goswami 2006/7
6 modules hypothesized for mammals, our dataset supports 5 (missing zygomatic)
```{r}
# Deconstruct function from Thomas's code in Landmark-test
ref <- mshape(mean.shapes)
mod.levels<- sort(unique(lm.modules[, 4]))  # 5 (of 6) modules from Goswami

# Define nice colors for modules
mod.colors <- c(wes_palette("GrandBudapest1")[2:4], wes_palette("GrandBudapest2")[1:2])

# Subset the landmarks according to the modules
mod.parts <- list()
for(i in 1:length(mod.levels)){
  mod.parts[[i]] <- which(lm.modules[, 4] == mod.levels[[i]])
}
```

### Make image of 3D plot of the modules for Fig. 4A
```{r}
open3d(windowRect = c(0, 0, 2400, 2400))
for(i in 1:length(mod.levels)) {
        spheres3d(ref[mod.parts[[i]], 1], ref[mod.parts[[i]], 2], ref[mod.parts[[i]], 3], col = mod.colors[i], lit = TRUE, radius = 0.0015, asp = FALSE)
        #rgl.texts(ref[mod.parts[[i]], 1], ref[mod.parts[[i]], 2], ref[mod.parts[[i]], 3], mod.parts[[i]], offset = 100, color = c("black"))
}
#view3d(phi = 0)
#rgl.snapshot("../Data/Processed/module_ventral.png")

view3d(phi = 90)  # Lateral view, dorsal side up
rgl.snapshot("../Data/Processed/modules.png")

# Remove whitespace from new image (requires ImageMagik installed on PC)
system("mogrify ../Data/Processed/modules.png -trim ../Data/Processed/*.png")
```

## Test INTEGRATION with 5 modules for full and residual shape dataset - PLS

```{r}

basicran <- mean.shapes[which(lm.modules[, 4] == "basicran"), ,]
molar <- mean.shapes[which(lm.modules[, 4] == "molar"), ,]
orbital <- mean.shapes[which(lm.modules[, 4] == "orbital"), ,]
rostrum <- mean.shapes[which(lm.modules[, 4] == "rostrum"), ,]
vault <- mean.shapes[which(lm.modules[, 4] == "vault"), ,]

module_list <- list( basicran=basicran, molar=molar, orbital=orbital, rostrum=rostrum, vault=vault)

load(file = "../Data/Processed/06-integration.rda")


# Test with the phylogentic tree (and mean shapes)
# int.test.5.phy <- phylo.integration(mean.shapes, partition.gp = lm.modules[, 4], phy=aus.tree)
# int.test.5.res <- phylo.integration(residuals, partition.gp = lm.modules[, 4], phy=aus.tree)
# # Save time consuming tests
# save(int.test.5.phy, int.test.5.res, file = "../Data/Processed/06-integration.rda")

R_pls_matrix <- matrix( nrow=length(mean_modules), ncol=length(mean_modules))
rownames(R_pls_matrix) <- names(mean_modules)
colnames(R_pls_matrix) <- names(mean_modules)

R_pls_matrix[lower.tri(R_pls_matrix)] <- round((int.test.5.phy$pairwise.P.values), 3)
#for some reason, the upper.tri does not order the p values correctly. So below is an interim step that fils the upper triangle with appropriate p values.
R_pls_matrix <- t(R_pls_matrix)

R_pls_matrix[lower.tri(R_pls_matrix)] <- round(int.test.5.phy$r.pls.mat, 2)

# now for RESIDUALS

R_pls_matrix_res <- matrix( nrow=length(mean_modules), ncol=length(mean_modules))
rownames(R_pls_matrix_res) <- names(mean_modules)
colnames(R_pls_matrix_res) <- names(mean_modules)

R_pls_matrix_res[lower.tri(R_pls_matrix_res)] <- round((int.test.5.res$pairwise.P.values), 3)
#for some reason, the upper.tri does not order the p values correctly. So below is an interim step that fils the upper triangle with appropriate p values.
R_pls_matrix_res <- t(R_pls_matrix_res)

R_pls_matrix_res[lower.tri(R_pls_matrix_res)] <- round(int.test.5.res$r.pls.mat, 2)

t(R_pls_matrix_res)

Integration_results_supplementary <- cbind(R_pls_matrix, c(rep(NA,5)),R_pls_matrix_res)


#Or maybe it is a bit clearer to take out probability values and instead go for just highlighting them. 


Integration_results <- R_pls_matrix

#There's a lot of transposing going on here, but I don't know how to improve
Integration_results[upper.tri(Integration_results)] <- t(R_pls_matrix_res)[upper.tri(t(R_pls_matrix_res))]
  
diag(Integration_results) <-  unlist(lapply(module_list, nrow))
  
  
 


write.csv(Integration_results, file="../Data/Results/Integration_pls_results.csv")
write.csv(Integration_results_supplementary, file="../Data/Results/Supp_Full_Integration_pls_results.csv")

```

#test INTEGRATION between pairs of modules; this repeats the full-shape integration analysis, which contains the pairwise values, but is needed to get the analysis objects to also compare integration strengths

```{r}


#setting up the combinations to be analysed

Combinations <- c( names(module_list), names(module_list) )
Combinations <- unique(Combinations)
Combinations <- combn( Combinations , 2 )


#Make data frame
Combination_matrix <- matrix(nrow=ncol(Combinations), ncol=nrow(Combinations)) 
colnames(Combination_matrix) <- c("Rpls", "p")
rownames(Combination_matrix) <- paste(Combinations[1,], Combinations[2,], sep=" x ")
Combination_matrix <- as.data.frame(Combination_matrix)

#an object to receive all comparisons

integration_analyses <- vector("list", 10)



for (i in 1:ncol(Combinations)){
  
temp <- phylo.integration(module_list[[which(names(module_list)== Combinations[1,i])]],
                          module_list[[which(names(module_list)== Combinations[2,i])]],
                          phy=aus.tree)

integration_analyses[[i]]<- temp

Combination_matrix$Rpls[i] <- temp$r.pls
Combination_matrix$p [i] <- temp$P.value

}

#give integration analyses objects the appropriate names

names(integration_analyses) <- paste(Combinations[1,], Combinations[2,], sep=" x ")



#Now fix up table
R_pls_matrix <- matrix( nrow=length(module_list), ncol=length(module_list))


R_pls_matrix[lower.tri(R_pls_matrix)] <- round(Combination_matrix$p, 3)
#necessary because I can't figure out how to fill the upper triangle properly, so transposing lower triangle instead
R_pls_matrix <- t(R_pls_matrix)
R_pls_matrix[lower.tri(R_pls_matrix)] <- round(Combination_matrix$Rpls, 2)

rownames(R_pls_matrix) <- names(module_list)
colnames(R_pls_matrix) <- names(module_list)


#note that the R-pls values should be the same as the overall integration output
#R_pls_matrix
#Integration_results; R_pls_matrix

#Create all combinations of modules

Combinations_compare <- c( names(module_list), names(module_list) )
Combinations_compare <- unique( Combinations_compare )
Combinations_compare <- combn( Combinations_compare , 2 )
#Combinations_compare <- paste(Combinations_compare[1,], Combinations_compare[2,])



#make empty list to receive integration analyses objects and name appropriately
phylo_integration_list <- vector("list",ncol(Combinations_compare))
names(phylo_integration_list) <- c( paste(Combinations_compare[1,], Combinations_compare[2,]))


for (i in 1:ncol(Combinations_compare)){
  
phylo_integration_list [[i]] <- phylo.integration(module_list[[which(names(module_list)==Combinations_compare[2,i])]],
                          module_list[[which(names(module_list)==Combinations_compare[1,i])]],
                          phy=aus.tree, iter=1000)

}

#After all this, we now just run a simple compare.pls line to compare integration strengths!

pls_comparison <- compare.pls(phylo_integration_list)


pls_comparison_matrix_P <- matrix(round(pls_comparison$pairwise.P,3), 
                                nrow=ncol(Combinations_compare), 
                                ncol=ncol(Combinations_compare), 
                                dimnames=list(rownames(pls_comparison$pairwise.P), colnames(pls_comparison$pairwise.P)))

pls_comparison_matrix_Z <- matrix(round(pls_comparison$pairwise.z, 2), 
                                         nrow=ncol(Combinations_compare), 
                                         ncol=ncol(Combinations_compare),
                                         dimnames=list(rownames(pls_comparison$pairwise.P), colnames(pls_comparison$pairwise.P)))


#Here we just add the p-values as the upper triangle of the effect size table
pls_comparison_matrix_Z[upper.tri(pls_comparison_matrix_Z)]<- pls_comparison_matrix_P[upper.tri(pls_comparison_matrix_P, diag=FALSE)]

write.csv (pls_comparison_matrix_Z,file="../Data/Results/integration_strength_comparisons.csv")


#manual checks
# rostrum_molar <- phylo.integration( rostrum, molar, phy = aus.tree)
# orbital_vault <- phylo.integration(orbital,vault, phy = aus.tree)
# vault_basicran <- phylo.integration( vault, basicran, phy = aus.tree)
# 
# testcompare <- compare.pls( rostrum_molar, orbital_vault, vault_basicran)
# testcompare$pairwise.z
# testcompare$pairwise.P
# 
# #effect sizes and p values differ depending on the order in which modules are input?
# test4 <- phylo.integration(module_list$molar, module_list$rostrum, phy = aus.tree)
# test5 <- phylo.integration(module_list$rostrum, module_list$molar, phy = aus.tree)
# 
# compare.pls(test4,test5)
# 
# #serious differences with rostrum and orbital
# 
# test5 <- phylo.integration(module_list$rostrum, module_list$orbital, phy = aus.tree)
# test6 <- phylo.integration(module_list$orbital, module_list$rostrum, phy = aus.tree)
# 
# compare.pls(test5,test6)
# 

```


#Test INTEGRATION strengths between full and residual datasets

```{r}

full_vs_res_int_strengths <- compare.pls( int.test.5.phy,int.test.5.res)

```


#test RESIDUALS INTEGRATION between pairs of modules; 

```{r}

basicran_res <- residuals[which(lm.modules[, 4] == "basicran"), ,]
molar_res <- residuals[which(lm.modules[, 4] == "molar"), ,]
orbital_res <- residuals[which(lm.modules[, 4] == "orbital"), ,]
rostrum_res <- residuals[which(lm.modules[, 4] == "rostrum"), ,]
vault_res <- residuals[which(lm.modules[, 4] == "vault"), ,]

module_list_res <- list( basicran_res=basicran_res, molar_res=molar_res, orbital_res=orbital_res, rostrum_res=rostrum_res, vault_res=vault_res)


#C#setting up the combinations to be analysed


Combinations <- c( names(module_list_res), names(module_list_res) )
Combinations <- unique( Combinations )
Combinations <- combn( Combinations , 2 )


#Make data frame
Combination_matrix <- matrix(nrow=ncol(Combinations), ncol=nrow(Combinations)) 
colnames(Combination_matrix) <- c("Rpls", "p")
rownames(Combination_matrix) <- paste(Combinations[1,], Combinations[2,], sep=" x ")
Combination_matrix <- as.data.frame(Combination_matrix)

#an object to receive all comparisons

integration_analyses_res <- vector("list", 10)



for (i in 1:ncol(Combinations)){
  
temp <- phylo.integration(module_list_res[[which(names(module_list_res)== Combinations[1,i])]],
                          module_list_res[[which(names(module_list_res)== Combinations[2,i])]],
                          phy=aus.tree)

integration_analyses_res[[i]]<- temp

Combination_matrix$Rpls[i] <- temp$r.pls
Combination_matrix$p [i] <- temp$P.value

}


#give integration analyses objects the appropriate names

names(integration_analyses_res) <- paste(Combinations[1,], Combinations[2,], sep=" x ")

#Now fix up table
R_pls_matrix_res <- matrix( nrow=length(module_list), ncol=length(module_list))

R_pls_matrix_res[lower.tri(R_pls_matrix_res)] <- round(Combination_matrix$p, 3)
#again, somehow the order is wrong with the upper tri and I don't know how to fix it
R_pls_matrix_res <- t(R_pls_matrix_res)
R_pls_matrix_res[lower.tri(R_pls_matrix_res)] <- round(Combination_matrix$Rpls, 2)

rownames(R_pls_matrix_res) <- names(module_list_res)
colnames(R_pls_matrix_res) <- names(module_list_res)


#note that the R-pls values should be the same as the overall integration output for residuals
Integration_results; R_pls_matrix_res



#manual tests to make sure the matrix reflects reality - it does

phylo.integration(basicran_res, molar_res, phy=aus.tree)
phylo.integration(rostrum_res, molar_res, phy=aus.tree)
phylo.integration(molar_res, rostrum_res, phy=aus.tree)

#Now create a matrix that asks if the levels of integration are significantly different using compare.pls

#Create all combinations of modules

Combinations_compare <- c( names(module_list_res), names(module_list_res) )
Combinations_compare <- unique( Combinations_compare )
Combinations_compare <- combn( Combinations_compare , 2 )



#make empty list to receive integration analyses objects and name appropriately
phylo_integration_list <- vector("list",ncol(Combinations_compare))
names(phylo_integration_list) <- c( paste(Combinations_compare[1,], Combinations_compare[2,]))


for (i in 1:ncol(Combinations_compare)){
  
phylo_integration_list [[i]] <- phylo.integration(module_list_res[[which(names(module_list_res)==Combinations_compare[2,i])]],
                          module_list_res[[which(names(module_list_res)==Combinations_compare[1,i])]],
                          phy=aus.tree, iter=1000)

}

pls_comparison_res <- compare.pls(phylo_integration_list)

#then pull out results for just a matrix to present
pls_comparison_matrix_P_res <- matrix(round(pls_comparison_res$pairwise.P,3), 
                                nrow=ncol(Combinations_compare), 
                                ncol=ncol(Combinations_compare), 
                                dimnames=list(rownames(pls_comparison_res$pairwise.P), colnames(pls_comparison_res$pairwise.P)))

pls_comparison_matrix_Z_res <- matrix(round(pls_comparison$pairwise.z, 3), 
                                         nrow=ncol(Combinations_compare), 
                                         ncol=ncol(Combinations_compare),
                                         dimnames=list(rownames(pls_comparison_res$pairwise.P), colnames(pls_comparison_res$pairwise.P)))


#Here we just add the p-values as the upper triangle of the effect size table
pls_comparison_matrix_Z_res[upper.tri(pls_comparison_matrix_Z_res)]<- pls_comparison_matrix_P_res[upper.tri(pls_comparison_matrix_P_res, diag=FALSE)]

write.csv (pls_comparison_matrix_Z,file="../Data/Results/integration_strength_comparisons_residuals.csv")

```



#test strength of INTEGRATION between full and residual partitions

```{r}

#quick check that all is well

names(integration_analyses) == gsub("_res", "", names(integration_analyses_res))

int_strength_full_vs_res <- vector("list", length(integration_analyses))
names(int_strength_full_vs_res)<- names(integration_analyses)


for (i in 1:length(integration_analyses)){
  
  int_strength_full_vs_res [[i]] <- compare.pls(integration_analyses[[i]], integration_analyses_res[[i]])
  
  
}

#Tabulate results - effect sizes, pairwise effect size differences, p values

integration_strength_comparison_table_full_vs_res <- matrix (ncol = 4, nrow = length(integration_analyses), dimnames = list( names(integration_analyses), c("Z full", "Z residual", "Z difference", "p")))

for(i in 1:length(integration_analyses)){
  
  #So pulling out all the values requires unlisting the pls results read-out for each comparison pair and turning it into a vector. You can find the relevant numbers by just going through the readout of one comparison pair.
  integration_strength_comparison_table_full_vs_res[i,] <- as.vector(unlist(int_strength_full_vs_res[[i]])[c(1,2,6,10)])
  
  
}

#some post-processing rounding

integration_strength_comparison_table_full_vs_res [,1:3] <- round (integration_strength_comparison_table_full_vs_res [,1:3], 2)
integration_strength_comparison_table_full_vs_res [,4]  <- round (integration_strength_comparison_table_full_vs_res[,4], 3)

write.csv(integration_strength_comparison_table_full_vs_res, file = "../Data/Results/integration_strength_comparison_table_full_vs_res.csv")

#and because it is a lot of coding to get to this point, and a lot of potential for human error, here are  manual checks. 

# rostrum_molar <- phylo.integration( molar, rostrum, phy = aus.tree)
# orbital_vault <- phylo.integration(orbital,vault, phy = aus.tree)
# vault_basicran <- phylo.integration( basicran, vault, phy = aus.tree)
# 
# rostrum_molar_res <- phylo.integration( molar_res, rostrum_res, phy = aus.tree)
# orbital_vault_res <- phylo.integration(orbital_res,vault_res, phy = aus.tree)
# vault_basicran_res <- phylo.integration( basicran_res, vault_res, phy = aus.tree)
# 
# compare.pls (rostrum_molar, rostrum_molar_res) #OK
# compare.pls (orbital_vault, orbital_vault_res) #OK
# compare.pls (vault_basicran, vault_basicran_res)#OK

```


## Test MODULARITY with 5 modules for full shape dataset
This test uses covariance ratio (CR) coefficients (Adams 2016) to quantify the degree of modularity between hypothesized modules/partitions. The observed CR is compared to the distribution of randomized CR's, found by randomizing the landmarks into partitions of identical sample sizes. Accounts for random changes due to evolution via Brownian motion (Felice and Adams 2014). Modularity investigations only ask if significant modular structure is present given a specific hypothesis of modularity. Thus, unlike integraiton analyses, the CR values themselves are not tested statistically.
```{r}
#save time by pre-loaded file
load(file = "../Data/Processed/06-modularity.rda")

# Test with the phylogentic tree (and mean shapes)
# mod.test.5.phy <- phylo.modularity(mean.shapes, partition.gp = lm.modules[, 4], phy=aus.tree)
# mod.test.5.res <- phylo.modularity(residuals, partition.gp = lm.modules[, 4], phy=aus.tree)
# Save time consuming tests
# save(mod.test.5.phy, mod.test.5.res,  file = "../Data/Processed/06-modularity.rda")


# Make pairwise comparison table
mod.result.phy <- as.matrix(mod.test.5.phy$CR.mat)  # results
colnames(mod.result.phy) <- unique(lm.modules[, 4])  # names from 1st instance
rownames(mod.result.phy) <- unique(lm.modules[, 4])  # ...in lm.module[, 4]
alpha.key <- match(unique(lm.modules[, 4]), mod.levels)  # alphabetize table
mod.result.phy <- mod.result.phy[order(alpha.key), order(alpha.key)]  # alphabetize table


# # Make pairwise comparison table for RESIDUALS
mod.result.res <- as.matrix(mod.test.5.res$CR.mat)  # results

colnames(mod.result.res) <- unique(lm.modules[, 4])  # names from 1st instance
rownames(mod.result.res) <- unique(lm.modules[, 4])  # ...in lm.module[, 4]
mod.result.res <- mod.result.res[order(alpha.key), order(alpha.key)]

#phylo.modularity provides a pairwise table. Here, upper is residual data and lower is full data

Modularity_matrix <- matrix(NA, nrow = nrow(mod.result.phy), ncol = nrow(mod.result.phy))

Modularity_matrix[upper.tri(Modularity_matrix)]<- mod.result.res[upper.tri(mod.result.res, diag=FALSE)]

Modularity_matrix[lower.tri(Modularity_matrix)]<- mod.result.phy[lower.tri(mod.result.phy, diag=FALSE)]

rownames(Modularity_matrix)<- rownames(mod.result.phy)
colnames(Modularity_matrix)<- colnames(mod.result.phy)

diag(Modularity_matrix) <- n.num

Modularity_matrix <- round(Modularity_matrix, 2)

write.csv(Modularity_matrix, file="../Data/Results/modularity_pairwise.csv")

```


## Make and export Figure 4
```{r}
# Load png back in
module.plot <- readPNG("../Data/Processed/modules.png")

# Set up for export and for multiple panels
pdf("../Data/Results/Figure4_Modularity.pdf")
mat <- matrix(c(1, 1, 2, 3,4,5), 3, 2, byrow = TRUE)  # 3 plots, 2 row, 2 columns
layout(mat, widths = rep.int(1, ncol(mat)), heights = rep.int(0.5, nrow(mat))) 

# 1) Plot of modules
par(mar = c(1, 1, 1, 1))  # sets the margins
plot(c(0, dim(module.plot)[2]), c(0, dim(module.plot)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE)  # module.plot = png made above
rasterImage(module.plot, 0, 0, dim(module.plot)[2], dim(module.plot)[1])
# Labels for module plot
legend("topleft", "a", cex = 1.5, bty = "n")
legend("topleft", "module definitions (observed condition)", cex = 0.9, bty = "n", inset = c(0.05, 0.015))
text(1200, -3, "Basicranium", bty = "n", col = darken(mod.colors[1]))
text(600, 150, "Molar", bty = "n", col = mod.colors[2])
text(600, 400, "Orbital", bty = "n", col = darken(mod.colors[3]))
text(160, 200, "Rostrum", bty = "n", col = darken(mod.colors[4], amount = 0.3))
text(1550,400, "Vault", bty = "n", col = darken(mod.colors[5], amount = 0.35))




############Modularity

#1 plot of observed and resampled CR coefficients (for full dataset/w phylocorrection)

par(mar = c(3.5, 3.5, 1.5, 0.5))
plot(density(mod.test.5.phy$random.CR),
     main = "",
     xlab = "",
     xlim = c(0, 1),
     ylim = c(0, 100),
     ylab = "",
     cex.axis = 0.9,
     col = "black")

title(xlab = "CR coefficient (low CR = high modularity)", line = 2.4)
title(ylab = "Density", line = 2.35)
arrows(x0 = mod.test.5.phy$CR, y0 = 30, y1 = 0, length = 0.15, lwd = 1.5, col = "darkred")
text(mod.test.5.phy$CR , 35, paste( "observed: ", round(mod.test.5.phy$CR, 2)), cex = 0.9, col = "darkred")
legend("topleft", "b", cex = 1.5, bty = "n", inset = -0.05)
legend("topleft", "full shape", cex = 0.9, bty = "n", inset = c(0.06, 0))

# 2) Plot of pls coefficients (w/ phylo correction) for shape residual dataset
par(mar = c(3.5, 3.5, 1.5, 0.5))
plot(density(mod.test.5.res$random.CR),
     main = "",
     xlab = "",
     xlim = c(0, 1),
     ylim = c(0, 100),
     ylab = "",
     cex.axis = 0.9,
     col = "black")
# Labels for histogram
title(xlab = "CR coefficient (low CR = high modularity)", line = 2.4)

arrows(x0 = mod.test.5.res$CR, y0 = 30, y1 = 0, length = 0.15, lwd = 1.5, col = "darkred")
text(mod.test.5.res$CR , 35, paste("observed: ", round(mod.test.5.res$CR, 2)), cex = 0.9, col = "darkred")
legend("topleft", "c", cex = 1.5, bty = "n", inset = -0.05)
legend("topleft", "shape residual", cex = 0.9, bty = "n", inset = c(0.07, 0))

#Integration

# 3) Plot of observed and resampled pls coefficients (w/ phylo correction)
par(mar = c(3.5, 3.5, 1.5, 0.5))
plot(density(int.test.5.phy$random.r),
     main = "",
     xlab = "",
     xlim = c(0, 1),
     ylim = c(0, 5),
     ylab = "",
     cex.axis = 0.9,
     col = "black")
title(xlab = "PLS coefficient (low PLS = low integration))", line = 2.4)
title(ylab = "Density", line = 2.35)
arrows(x0 = int.test.5.phy$r.pls, y0 = 1.5, y1 = 0, length = 0.15, lwd = 1.5, col = "darkred")
text(int.test.5.phy$r.pls , 2.1, paste( "observed: ", round(int.test.5.phy$r.pls, 2)), cex = 0.9, col = "darkred")
legend("topleft", "d", cex = 1.5, bty = "n", inset = -0.05)
legend("topleft", "full shape", cex = 0.9, bty = "n", inset = c(0.06, 0))

# 4) Plot of pls coefficients (w/ phylo correction) for shape residual dataset
par(mar = c(3.5, 3.5, 1.5, 0.5))
plot(density(int.test.5.res$random.r),
     main = "",
     xlab = "",
     xlim = c(0, 1),
     ylim = c(0, 5),
     ylab = "",
     cex.axis = 0.9,
     col = "black")
# Labels for histogram
title(xlab = "PLS coefficient (low PLS = low integration)", line = 2.4)

arrows(x0 = int.test.5.res$r.pls, y0 = 1.5, y1 = 0, length = 0.15, lwd = 1.5, col = "darkred")
text(int.test.5.res$r.pls , 2.1, paste("observed: ", round(int.test.5.res$r.pls, 2)), cex = 0.9, col = "darkred")
legend("topleft", "e", cex = 1.5, bty = "n", inset = -0.05)
legend("topleft", "shape residual", cex = 0.9, bty = "n", inset = c(0.07, 0))


dev.off()
```

## Mantel Test to see if PCs of modules are equivalent
Uses method similar to workflow from Heatherington and Sherratt 2015 DOI: https://doi.org/10.1111/pala.12159
```{r}
# Define (admittably brute force) function to run Mantel test on 5 modules named as in Figure 4A. Takes in a shape dataset and landmark definitions of modules made earlier in the script.
MantelModularity <- function(shapes, lm.modules) { 
  # Separate modules - using 5-module framework 
  mod.1 <- shapes[which(lm.modules[, 4] == "basicran"), ,]
  mod.2 <- shapes[which(lm.modules[, 4] == "molar"), ,]
  mod.3 <- shapes[which(lm.modules[, 4] == "orbital"), ,]
  mod.4 <- shapes[which(lm.modules[, 4] == "rostrum"), ,]
  mod.5 <- shapes[which(lm.modules[, 4] == "vault"), ,]
  
  # Run PCAs
  PCA.1 <- gm.prcomp(mod.1)
  PCA.2 <- gm.prcomp(mod.2)
  PCA.3 <- gm.prcomp(mod.3)
  PCA.4 <- gm.prcomp(mod.4)
  PCA.5 <- gm.prcomp(mod.5)

  # Make distance matrices
  d.PCA.1 <- dist(PCA.1$x)
  d.PCA.2 <- dist(PCA.2$x)
  d.PCA.3 <- dist(PCA.3$x)
  d.PCA.4 <- dist(PCA.4$x)
  d.PCA.5 <- dist(PCA.5$x)
  
  # Run Mantel tests
  m.1v2 <- mantel(d.PCA.1, d.PCA.2, method = "pearson", permutations = 1000) 
  m.1v3 <- mantel(d.PCA.1, d.PCA.3, method = "pearson", permutations = 1000)
  m.1v4 <- mantel(d.PCA.1, d.PCA.4, method = "pearson", permutations = 1000)
  m.1v5 <- mantel(d.PCA.1, d.PCA.5, method = "pearson", permutations = 1000)

  m.2v3 <- mantel(d.PCA.2, d.PCA.3, method = "pearson", permutations = 1000) 
  m.2v4 <- mantel(d.PCA.2, d.PCA.4, method = "pearson", permutations = 1000)
  m.2v5 <- mantel(d.PCA.2, d.PCA.5, method = "pearson", permutations = 1000)

  m.3v4 <- mantel(d.PCA.3, d.PCA.4, method = "pearson", permutations = 1000)
  m.3v5 <- mantel(d.PCA.3, d.PCA.5, method = "pearson", permutations = 1000)

  m.4v5 <- mantel(d.PCA.4, d.PCA.5, method = "pearson", permutations = 1000)
  
  # Bonferroni correction
  p.vals.bon <- round(p.adjust(c(m.1v2$signif, m.1v3$signif, m.1v4$signif, m.1v5$signif, m.2v3$signif, m.2v4$signif, m.2v5$signif, m.3v4$signif, m.3v5$signif, m.4v5$signif), method = "bonferroni"), 3)
  
  # Make results table columns
  col1 <- c("-", p.vals.bon[1], p.vals.bon[2], p.vals.bon[3], p.vals.bon[4])
  col2 <- c(round(m.1v2$statistic, 3), "-", p.vals.bon[5], p.vals.bon[6], p.vals.bon[7])
  col3 <- c(round(m.1v3$statistic, 3), round(m.2v3$statistic, 3), "-", p.vals.bon[8], p.vals.bon[9])
  col4 <- c(round(m.1v4$statistic, 3), round(m.2v4$statistic, 3), round(m.3v4$statistic, 3), "-", p.vals.bon[10])
  col5 <- c(round(m.1v5$statistic, 3), round(m.2v5$statistic, 3), round(m.3v5$statistic, 3), round(m.4v5$statistic, 3), "-")

  # Combine columns into table
  modules <- sort(unique(as.factor(lm.modules[, 4])))
  mantel.results <- data.table(Modules = modules, basicranium = col1, molar = col2, orbital = col3, rostrum = col4, vault = col5)
  
  return(mantel.results)
  }

# Run on the full and shape residual datasets respectively. This outputs a weird data table format that then needs to be turned into a matrix for upper/lower triangle extraction
mantel.full <- MantelModularity(mean.shapes, lm.modules)
mantel.res <- MantelModularity(residuals, lm.modules)

#this screams at you but just because of the NAs. the values are fine. Annoyingly, it also cancels the running of the whole chunk!
mantel.full <- apply(as.matrix.noquote(mantel.full[,2:6]),2,as.numeric)

mantel.res <- apply(as.matrix.noquote(mantel.res[,2:6]),2,as.numeric)

# Add landmark numbers; col #4 gives the Goswami module definitions
 n.num <- c(sum(lm.modules[, 4] == "basicran"), sum(lm.modules[, 4] == "molar"), sum(lm.modules[, 4] == "orbital"), sum(lm.modules[, 4] == "rostrum"), sum(lm.modules[, 4] == "vault"))
# mantel.full <- cbind(mantel.full, n.num)  # add to tables for export
# mantel.res <- cbind(mantel.res, n.num)


```

Turn the above into a table . This is a really nifty way of combining upper and lower triangles
```{r}

Mantel_matrix <- matrix(NA, nrow = length(mod.levels), ncol = length(mod.levels))
#         
Mantel_matrix[upper.tri(Mantel_matrix)]<- mantel.res[upper.tri(mantel.full, diag=FALSE)]
# Here we have to flip the second matrix by transposing (t(mantel.res)) because the values are originally on the upper diagonal 
Mantel_matrix[lower.tri(Mantel_matrix)]<- t(mantel.full)[(lower.tri(t(mantel.res), diag=FALSE))]

diag(Mantel_matrix)<- n.num
                     
 rownames(Mantel_matrix)<- mod.levels
 colnames(Mantel_matrix)<- mod.levels

write.csv(Mantel_matrix, file="../Data/Results/Mantel_results.csv")

```



### Run global integration tests and make Supplementary Figure 3
Test whether rodent skull is integrated as a whole
```{r}
pdf("../Data/Results/SFigure3_GItests.pdf")
mat <- matrix(c(1:2), 1, 2, byrow = TRUE)  # 2 plots, 1 rows, 2 columns
layout(mat, widths = rep(1, dim(mat)[2]), heights = rep(0.5, dim(mat)[1]))   

# A) Full shape test
par(mar = c(4.5, 4.5, 2, 1))
globalIntegration(mean.shapes, ShowPlot = TRUE)
legend("bottomleft", "A", cex = 1.5, bty = 'n', inset = c(-0.15, 0))
legend("bottomleft", "full shape", cex = 1, bty = 'n', inset = c(0, 0.012))

# B) Residual test
par(mar = c(4.5, 4.5, 2, 1))
globalIntegration(residuals, ShowPlot = TRUE)
legend("bottomleft", "B", cex = 1.5, bty = 'n', inset = c(-0.15, 0))
legend("bottomleft", "shape residual", cex = 1, bty = 'n', inset = c(0, 0.012))

dev.off()
```
