---
title: "07-test-LM-region-differences"
author: "Ariel Marcy"
date: "10/21/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '../eco-rodents')
```

# Test variation within morphospace
What is causing high levels of allometry? Is it functional integration of the gnawing apparatus (mostly in the top cranium-rostrum area)? 

If so, we would expect to see the top cranium-rostrum to be the least variable across species compared to the basicranium-top cranium and basicranium-rostrum comparisons. 

### Load packages, functions, and data from previous steps
```{r message = FALSE}
library(png)  # needed to work with screenshots
library(colorspace)  # plot colors
# Load libraries (and install if necessary) for landvR package
if(!require(devtools)) install.packages("devtools")
if(!require(dispRity)) install.packages("dispRity")
if(!require(landvR)) install_github("TGuillerme/landvR")
library(data.table)
library(vegan)
library(geomorph)
source("../Functions/utilities.R")  # custom functions
load(file = "../Data/Processed/03-sm-PCA-with-color.rda")
```

## Choose the partitions
The three partitions will need to be designated from the LM numbers. The first time, this was done with `define.modules()` but each module was then saved in a .csv file so this step can be repeated much more quickly. 

### Choose the partitions for the first time
```{r}
modules <- define.modules(shape[, , 1], nmodules = 3)

```


### Load in the partitions for all subsequent times
```{r}


```

