---
title: "08-visualize-phylo-distances"
author: "Ariel Marcy"
date: "1/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '../eco-rodents')
```

# Test biome pairs


### Load packages, functions, and data from previous steps
```{r message = FALSE}
library(geomorph)
library(stringr)
library(picante)  # needed for distance matrix
source("../Functions/utilities.R")  # custom functions
load(file = "../Data/Processed/03-main-data.rda")
load(file = "../Data/Processed/03-tree-data.rda")
load(file = "../Data/Processed/03-graphing-vectors.rda")
load(file = "../Data/Processed/05-residual-data.rda")
```

## Analyze phylogenetic and morphological distance matrices
To analyze phylogenetic and morphological distance in the context of biomes, we first need to generate distances matrices that can be comparable.
```{r}
# Phylo distance matrix
phylo.matrix <- cophenetic(aus.tree) / 2  # divide by two to convert distance to millions of years since Last Common Ancestor

# Morphological distance matrix
mean.gpa <- gpagen(mean.shapes, ProcD = TRUE)  # use mean shapes of species
morph.matrix <- as.matrix(mean.gpa$procD)

# Make names shorter
long.names <- colnames(phylo.matrix)
short.names <- NULL 
for(i in 1:length(long.names)) {
        sep <- strsplit(long.names[i], "_")
        new.name <- paste(str_sub(sep[[1]][1], 1, 3), str_sub(sep[[1]][2], 1, 3), sep = "_")
        short.names <- c(short.names, new.name)
}

# Replace with short names
row.names(phylo.matrix) <- short.names  # phylo matrix
colnames(phylo.matrix) <- short.names
row.names(morph.matrix) <- short.names  # morph matrix
colnames(morph.matrix) <- short.names
```

### Turn distance matrices into dataframes and then combine
```{r}
# Write function that creates the same, comparable table structure taking in the distance matrices from above, and the name for the distance in the matrix.
make.table <- function(matrix, d.name) {
        new.table <- as.data.frame(as.table(matrix))
        new.table[, 4] <- NA  # initialize pair name column
        colnames(new.table) <- c("spp1", "spp2", d.name, "pair")
        
        # Remove same-same comparisons (where distance = 0)
        new.table <- new.table[which(new.table[, d.name] > 0), ]
        
        # Remove permutations by alphabetizing the species names for each pair
        for(i in 1:dim(new.table)[1]) {
                pair.names <- c(as.character(new.table[i, "spp1"]), as.character(new.table[i, "spp2"]))
                pair.index <- str_order(c(as.character(new.table[i, "spp1"]), as.character(new.table[i, "spp2"])))
                pair <- paste(pair.names[pair.index[1]], pair.names[pair.index[2]], sep = "-")  # pair names in alpha order
                new.table[i, "pair"] <- pair
        }
        
        new.table <- new.table[!duplicated(new.table[, "pair"]), ]  # only unique pair combinations
        return(new.table)
}

# Run on both matrices & merge by pair
phylo.table <- make.table(phylo.matrix, "phylo.d")
morph.table <- make.table(morph.matrix, "morph.d")
master.table <- merge(phylo.table, morph.table[, 3:4], by = "pair")
```

## Plot phylo-morph distance pairs (Figure TK)
```{r}
# Create color key
just.genera <- NULL
for(i in 1:length(short.names)){
        sep <- strsplit(short.names[i], "_")
        just.genera <- c(just.genera, sep[[1]][1])
}
gen.col <- col.phylo[match(just.genera, names(col.phylo))]
names(gen.col)[5] <- "Mas"
gen.col[5] <- gen.col[16]  # Fix Mastacomys color

# Points coded for each individual specimen
alpha.col.spp1 <- gen.col[2:37]  # subset because C pen isn't in this col
alpha.col.spp2 <- gen.col

# Save in table for ease in plotting later on
master.table$col.spp1 <- PlotByGroup(master.table, "spp1", alpha.col.spp1)
master.table$col.spp2 <- PlotByGroup(master.table, "spp2", alpha.col.spp2)

# Plot
plot(master.table$phylo.d, master.table$morph.d, col = master.table$col.spp1, bg = master.table$col.spp2, pch = 21, xlab = "Phylogenetic distance (~million years since common ancestor)", ylab = "Morphological distance")
```

## Repeat Phylo/Morph distance plot using residuals
```{r}
# Morphological distance matrix for residuals
mean.gpa.res <- gpagen(Allom.all.plus.consensus, ProcD = TRUE)
morph.matrix.res <- as.matrix(mean.gpa.res$procD)

# Shorten row and column names
row.names(morph.matrix.res) <- short.names
colnames(morph.matrix.res) <- short.names

# Make comparable table to phylo distance using function defined above
morph.table.r <- make.table(morph.matrix.res, "morph.d.r")
master.table.r <- merge(phylo.table, morph.table.r[, 3:4], by = "pair")

# Plot
plot(master.table.r$phylo.d, master.table.r$morph.d.r, col = master.table$col.spp1, bg = master.table$col.spp2, pch = 21, xlab = "Phylogenetic distance (~million years since common ancestor)", ylab = "Morphological distance RESIDUALS")
```

## Plot phylo/morph distance tables by species in biome paired with others in SAME BIOME
```{r}
# Set up loop for multi-part figure broken down by 7 regions in Australia
region.names <- c("Savannah", "Arid Zone", "NE Wet Forest", "NE Dry Forest", "Pilbara", "South West", "South East")

info.r <- info[, c("Genus", "Species", "Savannah", "AridZone", "NEWetForest", "NEDryForest", "Pilbara", "SW", "SE")]  # Prune to only relevant metadata cols
info.r$Taxa <- paste(info.r$Genus, info.r$Species, sep = "_")
info.r <- unique(info.r)  # remove extra spp entries

par(mfrow=c(4,2), mar = c(3,3,3,3))

# For loop to cycle through each biome and plot phylo/morph distance 
for (i in 1:length(region.names)) {
        # Prune spp1 column to only species in each biome
        in.region <- FoundInRegion(info.r, regions = i)
        info.region <- unique(info.r[in.region, c("Genus", "Species")])
        info.region$spp1 <- paste(info.region$Genus, info.region$Species, sep = "_")
        spp1.region <- merge(master.table, info.region, by = "spp1")

        # Repeat with spp2 column (leaving only same-biome pairs)
        info.region$spp2 <- paste(info.region$Genus, info.region$Species, sep = "_")
        same.region.pairs <- merge(spp1.region, info.region, by = "spp2")
        
        # Plot
        plot(same.region.pairs$phylo.d, 
             same.region.pairs$morph.d, 
             col = same.region.pairs$col.spp2, 
             bg = same.region.pairs$col.spp1, 
             pch = 21,
             main = region.names[i],
             xlab = "Phylogenetic distance (~million years since LCA)",  ylab = "Morphological distance")
}
```

## Plot phylo/morph distance tables by species in biome paired with others in DIFFERENT BIOME
```{r}
# Figure set up
par(mfrow=c(4,2), mar = c(3,3,3,3))  

# For loop to cycle through each region and plot phylo/morph distance 
diff.reg.pairs <- NULL

for (i in 1:length(region.names)) {
        # Prune spp1 column to only species in each region
        in.region <- FoundInRegion(info.r, regions = i)
        info.region <- info.r[in.region, ]
        info.region$spp1 <- paste(info.region$Genus, info.region$Species, sep = "_")
        spp1.region <- merge(master.table, info.region, by = "spp1")

       # For each species in region, ID spp that never overlap and keep only those pairs (different region pairs)
        info.in.region <- info.r[in.region, ]  # all region presence/absence
        diff.spp <- info.r[!in.region, ]
        diff.reg.pairs <- NULL
        diff.all.pairs <- NULL
        for(s in 1:dim(info.in.region)[1]){  # For each species in a region...
                # Make vector of regions it can be found
                region.vector <- info.in.region[s, ]
                exist.in <- which(region.vector == 1)
                
                # Find DIFF pairs that don't exist any of the same regions 
                for(r in 1:length(exist.in)){  # ...ID different pairs
                        presence <- diff.spp[, exist.in[r]]
                        diff.spp <- diff.spp[presence < 1, ]
                        
                        # Alphabetize pair names
                        for(p in 1:dim(diff.spp)[1]){
                                pair.names <- c(diff.spp[p, "Taxa"], region.vector[, "Taxa"])
                                pair.index <- str_order(c(diff.spp[p, "Taxa"], region.vector[, "Taxa"]))
                                pair <- paste(pair.names[pair.index[1]], pair.names[pair.index[2]], sep = "-")  # pair names in alpha order
                        }
                        
                        # Find pair data for region's plot
                        diff.spp.pairs <- master.table[which(master.table$pair == pair), ]
                        diff.all.pairs <- rbind(diff.all.pairs, diff.spp.pairs)
                }
                diff.reg.pairs <- rbind(diff.reg.pairs, diff.all.pairs)
                
                }
        
        # Plot
        plot(diff.all.pairs$phylo.d, 
             diff.all.pairs$morph.d, 
             col = diff.all.pairs$col.spp2, 
             bg = diff.all.pairs$col.spp1, 
             pch = 21,
             main = paste(region.names[i], "same-different"),
             xlab = "Phylogenetic distance (~million years since LCA)",  ylab = "Morphological distance")
        # text(diff.all.pairs$phylo.d, 
             # diff.all.pairs$morph.d + 0.003, 
             # labels = diff.all.pairs$pair, cex = 0.7)
}
```
