---
title: "App04-test-fitted-allometry"
author: "Ariel Marcy"
date: "3/5/2020"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '../eco-rodents')
```

# Test for differences in the 3 methods to get fitted allometric shapes from procD.lm()
1) Simply using the `$fitted` object found in a procD.lm object.

2) Using the mysterious `$GM$fitted` object not described in the help file.

3) Using the lengthy workflow suggested by the function's documentation involving plot objects and regression scores.

This script produces Supplementary Figure TK and is relevant to Figure 2's second column, which is currently using the method from #1 (script 05).  

# Two conclusions after running this:
1) `$fitted` and `$GM$fitted` are exactly the same (`$fitted` just needs `arrayspecs()` to be in the right coordinates format for *geomorph*). Either is appropriate to use for allometric fitted shape heatmaps - see Figure 2 for this ms. 

2) Surprisingly, the workflow recommended in the documentation (#3) seems to produce the residual shape variation along PC1! You can compare column 3 in the figure made by this script with column 1 in Figure 3, the residual heatmap figure in this ms. 

### Load packages, functions, and data
```{r message = FALSE}
library(geomorph)
library(data.table)
library(geiger)  # for phylogenies
library(png)  # needed to work with screenshots
library(colorspace)  # plot colors
# Load libraries (and install if necessary) for landvR package
if(!require(devtools)) install.packages("devtools")
if(!require(dispRity)) install.packages("dispRity")
if(!require(landvR)) install_github("TGuillerme/landvR")
# Load previous data
load(file = "../Data/Processed/03-main-data.rda")
load(file = "../Data/Processed/03-tree-data.rda")
load(file = "../Data/Processed/03-graphing-vectors.rda")
```

## 1) Heatmaps using procD.lm's $fitted
First find mean specimen, then smallest and largest fitted allometric shape. Uses Dr Thomas Guillerme's new package, `landvR` to see how landmarks vary between these specimens.

Exact same code as script 05; can skip if script 05 has been run.
```{r}
# Fit specimens
allom.fit <- procD.lm(mean.shapes ~ log(info.means$MeanCsize))
fit.shapes <- arrayspecs(allom.fit$fitted, 325, 3)

# Find mean configuration - i.e. mean shape in dataset
allom.mean <- (select.procrustes(fit.shapes, selector = mean))[[1]]

# Measure spherical coordinates differences from the mean for each specimen for each landmark
allom.diff.mean <- coordinates.difference(coordinates = fit.shapes, reference = allom.mean, type = "spherical")

# Find largest and smallest native species from mean centroid size
sm.index <- which(info.means$Species == "del")
big.index <- which(info.means$MeanCsize == max(info.means$MeanCsize))
smallest.shape <- fit.shapes[, , sm.index]
biggest.shape <- fit.shapes[, , big.index]
```

### Export the heatmaps for fitted allometry shape change
Export images for plots
```{r}
# Lateral view of smallest to mean
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(allom.mean, smallest.shape, col = heat.colors, col.val = allom.diff.mean$Pseudomys_delicatulus[, "radius"])
view3d(phi = 90)
rgl.snapshot("../Data/Processed/heatmap_ALL_sm-mean_lat.png")

# Lateral view of mean to biggest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(allom.mean, biggest.shape, col = heat.colors, col.val = allom.diff.mean$Uromys_caudimaculatus[, "radius"])
view3d(phi = 90)
rgl.snapshot("../Data/Processed/heatmap_ALL_mean-big_lat.png")

# Dorsal view of smallest to mean
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(allom.mean, smallest.shape, col = heat.colors, col.val = allom.diff.mean$Pseudomys_delicatulus[, "radius"]) 
view3d(phi = 0)  # flip cranium
rgl.snapshot("../Data/Processed/heatmap_ALL_sm-mean_dor.png")

# Dorsal view of mean to biggest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(allom.mean, biggest.shape, col = heat.colors, col.val = allom.diff.mean$Uromys_caudimaculatus[, "radius"])
view3d(phi = 0)  # flip cranium
rgl.snapshot("../Data/Processed/heatmap_ALL_mean-big_dor.png")
```

## 2) Heatmaps using procD.lm's $GM$fitted
```{r}
# Fitted specimens from object GM
fit.shapes.GM <- allom.fit$GM$fitted

# Find mean configuration - i.e. mean shape in dataset
allom.mean.GM <- (select.procrustes(fit.shapes.GM, selector = mean))[[1]]

# Measure spherical coordinates differences from the mean for each specimen for each landmark
allom.diff.mean.GM <- coordinates.difference(coordinates = fit.shapes.GM, reference = allom.mean.GM, type = "spherical")

# Store largest and smallest native species from mean centroid size
smallest.shape.GM <- fit.shapes.GM[, , sm.index]
biggest.shape.GM <- fit.shapes.GM[, , big.index]
```

### Export the heatmaps for $GM$fitted
Export images for plots
```{r}
# Lateral view of smallest to mean
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(allom.mean.GM, smallest.shape.GM, col = heat.colors, col.val = allom.diff.mean.GM$Pseudomys_delicatulus[, "radius"])
view3d(phi = 90)
rgl.snapshot("../Data/Processed/heatmap_AGM_sm-mean_lat.png")

# Lateral view of mean to biggest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(allom.mean.GM, biggest.shape.GM, col = heat.colors, col.val = allom.diff.mean.GM$Uromys_caudimaculatus[, "radius"])
view3d(phi = 90)
rgl.snapshot("../Data/Processed/heatmap_AGM_mean-big_lat.png")

# Dorsal view of smallest to mean
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(allom.mean.GM, smallest.shape.GM, col = heat.colors, col.val = allom.diff.mean.GM$Pseudomys_delicatulus[, "radius"]) 
view3d(phi = 0)  # flip cranium
rgl.snapshot("../Data/Processed/heatmap_AGM_sm-mean_dor.png")

# Dorsal view of mean to biggest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(allom.mean.GM, biggest.shape.GM, col = heat.colors, col.val = allom.diff.mean.GM$Uromys_caudimaculatus[, "radius"])
view3d(phi = 0)  # flip cranium
rgl.snapshot("../Data/Processed/heatmap_AGM_mean-big_dor.png")
```

## 3) Suggested workaround from procD.lm's documentation
```{r}
# Save procD.lm object as regression plot
allom.plot <- plot(allom.fit, type = "regression", predictor = info.means$MeanCsize, reg.type = "RegScore") 

# Extract predicted shapes for smallest and largest
preds <- shape.predictor(allom.fit$GM$fitted, x = allom.plot$RegScore,
                         predmin = min(allom.plot$RegScore),
                         predmax = max(allom.plot$RegScore) 
                         )

# Measure spherical coordinates differences from the mean for each specimen for each landmark - assumes mean shape of the $GM$fitted dataset is the appropriate consensus shape given its usage in the shape.predictor function above.
allom.diff.mean.Reg <- coordinates.difference(coordinates = preds, reference = allom.mean.GM, type = "spherical")
```

### Export the heatmaps for suggested workaround
Export images for plots
```{r}
# Lateral view of smallest to mean
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(allom.mean.GM, preds$predmin, col = heat.colors, col.val = allom.diff.mean.Reg$predmin[, "radius"])
view3d(phi = 90)
rgl.snapshot("../Data/Processed/heatmap_REG_sm-mean_lat.png")

# Lateral view of mean to biggest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(allom.mean.GM, preds$predmax, col = heat.colors, col.val = allom.diff.mean.Reg$predmax[, "radius"])
view3d(phi = 90)
rgl.snapshot("../Data/Processed/heatmap_REG_mean-big_lat.png")

# Dorsal view of smallest to mean
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(allom.mean.GM, preds$predmax, col = heat.colors, col.val = allom.diff.mean.Reg$predmin[, "radius"]) 
view3d(phi = 0)  # flip cranium
rgl.snapshot("../Data/Processed/heatmap_REG_sm-mean_dor.png")

# Dorsal view of mean to biggest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(allom.mean.GM, preds$predmax, col = heat.colors, col.val = allom.diff.mean.Reg$predmax[, "radius"])
view3d(phi = 0)  # flip cranium
rgl.snapshot("../Data/Processed/heatmap_REG_mean-big_dor.png")
```

### Clean and load in all heatmaps
```{r}
# Remove whitespace from all heatmaps (requires ImageMagik installed on PC)
system("mogrify ../Data/Processed/heatmap_*.png -trim ../Data/Processed/*.png")

# 1) Load in $fitted allometry
allom.sm.lat <- readPNG("../Data/Processed/heatmap_ALL_sm-mean_lat.png")
allom.big.lat <- readPNG("../Data/Processed/heatmap_ALL_mean-big_lat.png")
allom.sm.dor <- readPNG("../Data/Processed/heatmap_ALL_sm-mean_dor.png")
allom.big.dor <- readPNG("../Data/Processed/heatmap_ALL_mean-big_dor.png")

# 2) Load in $GM$fitted
GMfit.sm.lat <- readPNG("../Data/Processed/heatmap_AGM_sm-mean_lat.png")
GMfit.big.lat <- readPNG("../Data/Processed/heatmap_AGM_mean-big_lat.png")
GMfit.sm.dor <- readPNG("../Data/Processed/heatmap_AGM_sm-mean_dor.png")
GMfit.big.dor <- readPNG("../Data/Processed/heatmap_AGM_mean-big_dor.png")

# 3) Suggested workaround
regsc.sm.lat <- readPNG("../Data/Processed/heatmap_RES_sm-mean_lat.png")
regsc.big.lat <- readPNG("../Data/Processed/heatmap_RES_mean-big_lat.png")
regsc.sm.dor <- readPNG("../Data/Processed/heatmap_RES_sm-mean_dor.png")
regsc.big.dor <- readPNG("../Data/Processed/heatmap_RES_mean-big_dor.png")
```

## 4) Make multi-panel Supplementary Figure TK for comparison
```{r}
# Set up export and muli-panel figure 
pdf("../Data/Results/SFigure_Allometry_Compare.pdf")
mat <- matrix(c(1:12), 4, 3, byrow = FALSE)  # 12 plots, 4 rows, 3 columns
layout(mat, widths = rep(1, dim(mat)[2]), heights = rep(0.5, dim(mat)[1]))  

## Fitted allometry
# 1) Mean to Small
par(mar = c(0, 0, 0, 0))  # sets the margins
plot(c(0, dim(allom.sm.lat)[2]), c(0, dim(allom.sm.lat)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(allom.sm.lat, 0, 0, dim(allom.sm.lat)[2], dim(allom.sm.lat)[1])
text(0, dim(allom.sm.lat)[1]-10, "A", cex = 2)
legend("topright", inset = c(0.05, 0), legend = "procD.lm $fitted", bty = "n", cex = 1.2)
#legend("bottomleft", "Mean to P. delicatulus, fitted shape", bty = "n", inset = c(0.1, 0.05))

# 2) Mean to Big
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(allom.big.lat)[2]), c(0, dim(allom.big.lat)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(allom.big.lat, 0, 0, dim(allom.big.lat)[2], dim(allom.big.lat)[1])
text(0, dim(allom.big.lat)[1]-10, "B", cex = 2)
#legend("bottomleft", "Mean to U. caudimaculatus, fitted shape", bty = "n", inset = c(0.1, 0.12))

# 3) Mean to Small
par(mar = c(0, 0, 0, 0))  # sets the margins
plot(c(0, dim(allom.big.dor)[2]), c(0, dim(allom.big.dor)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(allom.sm.dor, 0, 0, dim(allom.big.dor)[2], dim(allom.big.dor)[1])
text(0, dim(allom.big.dor)[1]-10, "C", cex = 2)
#legend("bottomleft", "Dorsal view of E", bty = "n", inset = c(0.1, -0.11))

# 4) Mean to Big
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(allom.big.dor)[2]), c(0, dim(allom.big.dor)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(allom.big.dor, 0, 0, dim(allom.big.dor)[2], dim(allom.big.dor)[1])
text(0, dim(allom.big.dor)[1]-10, "D", cex = 2)
#legend("bottomleft", "Dorsal view of F", bty = "n", inset = c(0.1, 0.18))

## $GM$fitted 
# 5) Mean to Small
par(mar = c(0, 0, 0, 0))  # sets the margins
plot(c(0, dim(GMfit.sm.lat)[2]), c(0, dim(GMfit.sm.lat)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(GMfit.sm.lat, 0, 0, dim(GMfit.sm.lat)[2], dim(GMfit.sm.lat)[1])
text(0, dim(GMfit.sm.lat)[1]-10, "E", cex = 2)
legend("topright", inset = c(0.06, 0), legend = "$GM$fitted", bty = "n", cex = 1.2)
#legend("bottomleft", "Mean to hypothetical minimum", bty = "n", inset = c(0.1, 0.05))

# 6) Mean to Big
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(GMfit.big.lat)[2]), c(0, dim(GMfit.big.lat)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(GMfit.big.lat, 0, 0, dim(GMfit.big.lat)[2], dim(GMfit.big.lat)[1])
text(0, dim(GMfit.big.lat)[1]-10, "F", cex = 2)
#legend("bottomleft", "Mean to hypothetical maximum", bty = "n", inset = c(0.1, 0.12))

# 7) Mean to Small
par(mar = c(0, 0, 0, 0))  # sets the margins
plot(c(0, dim(GMfit.big.dor)[2]), c(0, dim(GMfit.big.dor)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(GMfit.sm.dor, 0, 0, dim(GMfit.big.dor)[2], dim(GMfit.big.dor)[1])
text(0, dim(GMfit.big.dor)[1]-10, "G", cex = 2)
#legend("bottomleft", "Dorsal view of I", bty = "n", inset = c(0.1, -0.11))

# 8) Mean to Big
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(GMfit.big.dor)[2]), c(0, dim(GMfit.big.dor)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(GMfit.big.dor, 0, 0, dim(GMfit.big.dor)[2], dim(GMfit.big.dor)[1])
text(0, dim(GMfit.big.dor)[1]-10, "H", cex = 2)
#legend("bottomleft", "Dorsal view of J", bty = "n", inset = c(0.1, 0.18))

## Suggested workaround
# 9) Mean to Small
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(regsc.sm.lat)[2]), c(0, dim(regsc.sm.lat)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(regsc.sm.lat, 0, 0, dim(regsc.sm.lat)[2], dim(regsc.sm.lat)[1])
text(0, dim(regsc.sm.lat)[1]-10, "I", cex = 2)
legend("topright", inset = c(0.06, 0), legend = "regscore workflow", bty = "n", cex = 1.2)
#legend("bottomleft", "Mean to hypothetical minimum", bty = "n", inset = c(0.1, 0.05))

# 10) Mean to Big
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(regsc.big.lat)[2]), c(0, dim(regsc.big.lat)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(regsc.big.lat, 0, 0, dim(regsc.big.lat)[2], dim(regsc.big.lat)[1])
text(0, dim(regsc.big.lat)[1]-10, "J", cex = 2)
#legend("bottomleft", "Mean to hypothetical maximum", bty = "n", inset = c(0.1, 0.12))

# 11) Mean to Small
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(regsc.big.dor)[2]), c(0, dim(regsc.big.dor)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(regsc.sm.dor, 0, 0, dim(regsc.big.dor)[2], dim(regsc.big.dor)[1])
text(0, dim(regsc.big.dor)[1]-10, "K", cex = 2)
#legend("bottomleft", "Dorsal view of A", bty = "n", inset = c(0.1, -0.11))

# 12) Mean to Big
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(regsc.big.dor)[2]), c(0, dim(regsc.big.dor)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(regsc.big.dor, 0, 0, dim(regsc.big.dor)[2], dim(regsc.big.dor)[1])
text(0, dim(regsc.big.dor)[1]-10, "L", cex = 2)
#legend("bottomleft", "Dorsal view of B", bty = "n", inset = c(0.1, 0.18))

# Make gradient legend
legend_image <- as.raster(matrix(heat.colors(20), ncol = 1))
rasterImage(legend_image, 483, 150, 507, 215)
text(407, 203, "High variation", col = "dark grey", cex = 1)
text(453, 162, "Low", col = "dark grey", cex = 1)

dev.off()
```