---
title: "App04-residuals-without"
author: "Ariel Marcy"
date: "3/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '../eco-rodents')
```

# Supplementary heatmaps residual shape without Notomys and carnivores
1) Make heatmaps with shape dataset including allometry.
        A) Smallest to mean to largest
        B) Smallest projected shape to mean to largest projected shape
        C) PC1 minimum to mean to PC maximum

Produces multipanel Figure 2.

2) Make heatmaps for residuals aka size-less shape dataset.
        A) PC1 min to mean to PC1 max
        B) PC1 min on subset without Notomys to mean to PC1 max of this subset
        C) PC2 min to mean to PC2 max

Produces multipanel Figure 3.

### Load packages, functions, and data
Tree, mean shape, and metadata for phylogenetic analyses come from the allometry rodents script 04. 
```{r message = FALSE}
library(geomorph)
library(data.table)
library(vegan)
library(geiger)  # for phylogenies
library(png)  # needed to work with screenshots
library(colorspace)  # plot colors

# Load libraries (and install if necessary) for landvR package
if(!require(devtools)) install.packages("devtools")
if(!require(dispRity)) install.packages("dispRity")
if(!require(landvR)) install_github("TGuillerme/landvR")
source("../Functions/utilities.R")  # custom functions
source("../Functions/plotGMPhyloMorphoSpace_plotmod.R")  # modified geomorph function
load(file = "../Data/Processed/03-main-data.rda")
load(file = "../Data/Processed/03-tree-data.rda")
load(file = "../Data/Processed/03-graphing-vectors.rda")
load(file = "../Data/Processed/04-PCA.rda")
load(file = "../Data/Processed/05-residual-data.rda")
```

## 1) Prepare heatmaps
This script is dominated by the code needed to generate, export, and clean the heatmap images which are tiled in supplementary figure TK.

### Make heatmaps for PC1 with residual coordinates
Steps as above.
```{r}
# Find mean configuration - i.e. mean shape in dataset
res.mean <- (select.procrustes(residuals, selector = mean))[[1]]

# Measure spherical coordinates differences from the mean for each specimen for each landmark
res.diff.mean <- coordinates.difference(coordinates = residuals, reference = res.mean, type = "spherical")

# Ordinate the data
twoD.shape.r <- two.d.array(residuals)
ordination.r <- stats::prcomp(twoD.shape.r)

# Force the symmetric component of shape into class "gpagen" (required for variation.range)
gpagen.shape <- list()
gpagen.shape$coords <- residuals
gpagen.shape$consensus <- res.mean
class(gpagen.shape) <- "gpagen"

# Measure extremes of variation from mean on PC1
PC1.var.r <- variation.range(gpagen.shape, return.ID = FALSE, axis = 1, ordination = ordination.r, type = "spherical")

# Wrap specimens on the tangent space
wrap.PCA.r <- plotTangentSpace(residuals, verbose = FALSE)

# Select extreme specimens (the ones that make the warp-meshes)
res.PC1.min.h <- wrap.PCA.r$pc.shapes[[2]]  # Flipped b/c axes are flipped
res.PC1.max.h <- wrap.PCA.r$pc.shapes[[1]]
```

### Export the variation for residual shape change
```{r}
# Lateral view of mean to smallest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean, res.PC1.min.h, col = heat.colors, col.val = PC1.var.r[, "radius"])
view3d(phi = 90)
rgl.snapshot("../Data/Processed/heatmap_RES_sm-mean_lat.png")

# Lateral view of mean to biggest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean, res.PC1.max.h, col = heat.colors, col.val = PC1.var.r[, "radius"])
view3d(phi = 90)
rgl.snapshot("../Data/Processed/heatmap_RES_mean-big_lat.png")

# Dorsal view of mean to smallest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean, res.PC1.min.h, col = heat.colors, col.val = PC1.var.r[, "radius"]) 
view3d(phi = 0)  # flip cranium
rgl.snapshot("../Data/Processed/heatmap_RES_sm-mean_dor.png")

# Dorsal view of mean to biggest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean, res.PC1.max.h, col = heat.colors, col.val = PC1.var.r[, "radius"])
view3d(phi = 0)  # flip cranium
rgl.snapshot("../Data/Processed/heatmap_RES_mean-big_dor.png")
```

### Make heatmaps for PC2 residual coordinates
Steps as above but for PC2; hypothetical shapes
```{r}
# Measure extremes of variation from mean on PC1
PC1.var.r2 <- variation.range(gpagen.shape, return.ID = FALSE, axis = 2, ordination = ordination.r, type = "spherical")

# Select extreme specimens (the ones that make the warp-meshes)
res.PC2.min.h <- wrap.PCA.r$pc.shapes[[4]]
res.PC2.max.h <- wrap.PCA.r$pc.shapes[[3]]
```

### Export the variation for residual shape change
```{r}
# Lateral view of mean to smallest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean, res.PC2.min.h, col = heat.colors, col.val = PC1.var.r2[, "radius"])
view3d(phi = 90)
rgl.snapshot("../Data/Processed/heatmap_RP2_sm-mean_lat.png")

# Lateral view of mean to biggest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean, res.PC2.max.h, col = heat.colors, col.val = PC1.var.r2[, "radius"])
view3d(phi = 90)
rgl.snapshot("../Data/Processed/heatmap_RP2_mean-big_lat.png")

# Dorsal view of mean to smallest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean, res.PC2.min.h, col = heat.colors, col.val = PC1.var.r2[, "radius"]) 
view3d(phi = 0)  # flip cranium
rgl.snapshot("../Data/Processed/heatmap_RP2_sm-mean_dor.png")

# Dorsal view of mean to biggest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean, res.PC2.max.h, col = heat.colors, col.val = PC1.var.r2[, "radius"])
view3d(phi = 0)  # flip cranium
rgl.snapshot("../Data/Processed/heatmap_RP2_mean-big_dor.png")
```

### Remove Notomys and find hypothetical shapes
```{r}
# Remove Notomys from dataset
index.not <- which(info.means$Genus == "Not")
residuals.no <- residuals[, , -index.not]
info.means.no <- info.means[-index.not, ]

# Find new mean
res.mean.no <- (select.procrustes(residuals.no, selector = mean))[[1]]

# Measure spherical coordinates differences from the mean for each specimen for each landmark
res.diff.mean.no <- coordinates.difference(coordinates = residuals.no, reference = res.mean.no, type = "spherical")

# Ordinate the data
twoD.shape.r.n <- two.d.array(residuals.no)
ordination.r.n <- stats::prcomp(twoD.shape.r.n)

# Force the symmetric component of shape into class "gpagen" (required for variation.range)
gpagen.shape.no <- list()
gpagen.shape.no$coords <- residuals.no
gpagen.shape.no$consensus <- res.mean.no
class(gpagen.shape.no) <- "gpagen"

# Measure extremes of variation from mean on PC1
PC1.var.r.n <- variation.range(gpagen.shape.no, return.ID = FALSE, axis = 1, ordination = ordination.r, type = "spherical")

# Wrap specimens on the tangent space
wrap.PCA.r.n <- plotTangentSpace(residuals.no, verbose = FALSE)

# Select extreme specimens (the ones that make the warp-meshes)
res.PC1.min.h.n <- wrap.PCA.r.n$pc.shapes[[1]] 
res.PC1.max.h.n <- wrap.PCA.r.n$pc.shapes[[2]]
```

### Export residual no Notomys heatmaps
```{r}
# Lateral view of mean to smallest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean, res.PC1.min.h.n, col = heat.colors, col.val = PC1.var.r.n[, "radius"])
view3d(phi = 90)
rgl.snapshot("../Data/Processed/heatmap_RES_sm-mean_lat_NOt.png")

# Lateral view of mean to biggest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean, res.PC1.max.h.n, col = heat.colors, col.val = PC1.var.r.n[, "radius"])
view3d(phi = 90)
rgl.snapshot("../Data/Processed/heatmap_RES_mean-big_lat_NOt.png")

# Dorsal view of mean to smallest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean, res.PC1.min.h.n, col = heat.colors, col.val = PC1.var.r.n[, "radius"]) 
view3d(phi = 0)  # flip cranium
rgl.snapshot("../Data/Processed/heatmap_RES_sm-mean_dor_NOt.png")

# Dorsal view of mean to biggest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean, res.PC1.max.h.n, col = heat.colors, col.val = PC1.var.r.n[, "radius"])
view3d(phi = 0)  # flip cranium
rgl.snapshot("../Data/Processed/heatmap_RES_mean-big_dor_NOt.png")

# Remove whitespace from all heatmaps (requires ImageMagik installed on PC)
system("mogrify ../Data/Processed/heatmap_*.png -trim ../Data/Processed/*.png")
```

### Load in all heatmaps
```{r}
# Load in with allometry smallest to largest
indiv.sm.lat <- readPNG("../Data/Processed/heatmap_IND_sm-mean_lat.png")
indiv.big.lat <- readPNG("../Data/Processed/heatmap_IND_mean-big_lat.png")
indiv.sm.dor <- readPNG("../Data/Processed/heatmap_IND_sm-mean_dor.png")
indiv.big.dor <- readPNG("../Data/Processed/heatmap_IND_mean-big_dor.png")

# Load in with allometry smallest to largest
allom.sm.lat <- readPNG("../Data/Processed/heatmap_ALL_sm-mean_lat.png")
allom.big.lat <- readPNG("../Data/Processed/heatmap_ALL_mean-big_lat.png")
allom.sm.dor <- readPNG("../Data/Processed/heatmap_ALL_sm-mean_dor.png")
allom.big.dor <- readPNG("../Data/Processed/heatmap_ALL_mean-big_dor.png")

# Load in with allometry PC 1
allPC.sm.lat <- readPNG("../Data/Processed/heatmap_PCA_sm-mean_lat.png")
allPC.big.lat <- readPNG("../Data/Processed/heatmap_PCA_mean-big_lat.png")
allPC.sm.dor <- readPNG("../Data/Processed/heatmap_PCA_sm-mean_dor.png")
allPC.big.dor <- readPNG("../Data/Processed/heatmap_PCA_mean-big_dor.png")

# Load in without allometry residuals PC1
resid.sm.lat <- readPNG("../Data/Processed/heatmap_RES_sm-mean_lat.png")
resid.big.lat <- readPNG("../Data/Processed/heatmap_RES_mean-big_lat.png")
resid.sm.dor <- readPNG("../Data/Processed/heatmap_RES_sm-mean_dor.png")
resid.big.dor <- readPNG("../Data/Processed/heatmap_RES_mean-big_dor.png")

# Load in without allometry residuals PC2
resP2.sm.lat <- readPNG("../Data/Processed/heatmap_RP2_sm-mean_lat.png")
resP2.big.lat <- readPNG("../Data/Processed/heatmap_RP2_mean-big_lat.png")
resP2.sm.dor <- readPNG("../Data/Processed/heatmap_RP2_sm-mean_dor.png")
resP2.big.dor <- readPNG("../Data/Processed/heatmap_RP2_mean-big_dor.png")

# Load in without allometry residuals PC1 - no Notomys
resid.sm.lat.no <- readPNG("../Data/Processed/heatmap_RES_sm-mean_lat_NOt.png")
resid.big.lat.no <- readPNG("../Data/Processed/heatmap_RES_mean-big_lat_NOt.png")
resid.sm.dor.no <- readPNG("../Data/Processed/heatmap_RES_sm-mean_dor_NOt.png")
resid.big.dor.no <- readPNG("../Data/Processed/heatmap_RES_mean-big_dor_NOt.png")
```

## 3) Make multi-panel Figure 3: Heatmaps with Residuals
```{r}
# Set up muli-panel figure; lateral 
pdf("../Data/Results/Figure3_Heatmap_Residuals.pdf")
mat <- matrix(c(1:12), 4, 3, byrow = FALSE)  # 12 plots, 4 rows, 3 columns
layout(mat, widths = rep(1, dim(mat)[2]), heights = rep(0.5, dim(mat)[1])) 

## PC1 Residuals
# 1) Mean to Small, residuals PC1, lateral view
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resid.sm.lat)[2]), c(0, dim(resid.sm.lat)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(resid.sm.lat, 0, 0, dim(resid.sm.lat)[2], dim(resid.sm.lat)[1])
text(0, dim(resid.sm.lat)[1]-10, "A", cex = 2)
legend("topright", inset = c(0.3, 0), legend = "PC1 Residuals", bty = "n", cex = 1.2)
legend("bottomleft", "Mean to hypothetical minimum", bty = "n", inset = c(0.1, 0.07))

# 2) Mean to Big, residuals PC1, lateral view
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resid.big.lat)[2]), c(0, dim(resid.big.lat)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(resid.big.lat, 0, 0, dim(resid.big.lat)[2], dim(resid.big.lat)[1])
text(0, dim(resid.big.lat)[1]-10, "B", cex = 2)
legend("bottomleft", "Mean to hypothetical maximum", bty = "n", inset = c(0.1, 0.07))

# 3) Mean to Small, residuals PC1, dorsal view
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resid.big.dor)[2]), c(0, dim(resid.big.dor)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(resid.sm.dor, 0, 0, dim(resid.big.dor)[2], dim(resid.big.dor)[1])
text(0, dim(resid.big.dor)[1]-10, "C", cex = 2)
legend("bottomleft", "Dorsal view of A", bty = "n", inset = c(0.2, 0.05))

# 4) Mean to Big, residuals PC1, dorsal view
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resid.big.dor)[2]), c(0, dim(resid.big.dor)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(resid.big.dor, 0, 0, dim(resid.big.dor)[2], dim(resid.big.dor)[1])
text(0, dim(resid.big.dor)[1]-10, "D", cex = 2)
legend("bottomleft", "Dorsal view of B", bty = "n", inset = c(0.2, 0.05))

## PC 1 Residuals -- without Notomys
# 5) Mean to Small, residuals PC1, lateral view - no Notomys
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resid.sm.lat.no)[2]), c(0, dim(resid.sm.lat.no)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(resid.sm.lat.no, 0, 0, dim(resid.sm.lat.no)[2], dim(resid.sm.lat.no)[1])
text(0, dim(resid.sm.lat.no)[1]-10, "E", cex = 2)
legend("topright", inset = c(0.1, 0), legend = expression(paste("PC1 Residuals (no ", italic("Notomys"), ")")), bty = "n", cex = 1.2)
legend("bottomleft", "Mean to new hypothetical minimum", bty = "n", inset = c(0.07, 0.07))

# 6) Mean to Big, residuals PC1, lateral view - no Notomys
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resid.big.lat.no)[2]), c(0, dim(resid.big.lat.no)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(resid.big.lat.no, 0, 0, dim(resid.big.lat.no)[2], dim(resid.big.lat.no)[1])
text(0, dim(resid.big.lat.no)[1]-10, "F", cex = 2)
legend("bottomleft", "Mean to new hypothetical maximum", bty = "n", inset = c(0.07, 0.07))

# 7) Mean to Small, residuals PC1, dorsal view - no Notomys
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resid.big.dor.no)[2]), c(0, dim(resid.big.dor.no)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(resid.sm.dor.no, 0, 0, dim(resid.big.dor.no)[2], dim(resid.big.dor.no)[1])
text(0, dim(resid.big.dor.no)[1]-10, "G", cex = 2)
legend("bottomleft", "Dorsal view of E", bty = "n", inset = c(0.2, 0.05))

# 8) Mean to Big, residuals PC1, dorsal view - no Notomys
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resid.big.dor.no)[2]), c(0, dim(resid.big.dor.no)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(resid.big.dor.no, 0, 0, dim(resid.big.dor.no)[2], dim(resid.big.dor.no)[1])
text(0, dim(resid.big.dor.no)[1]-10, "H", cex = 2)
legend("bottomleft", "Dorsal view of F", bty = "n", inset = c(0.2, 0.05))

## PC2 (whole residual dataset)
# 9) Mean to Small, residuals PC2, lateral view
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resP2.sm.lat)[2]), c(0, dim(resP2.sm.lat)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(resP2.sm.lat, 0, 0, dim(resP2.sm.lat)[2], dim(resP2.sm.lat)[1])
text(0, dim(resP2.sm.lat)[1]-10, "I", cex = 2)
legend("topright", inset = c(0.17, 0), legend = "PC2 Residuals (all spp.)", bty = "n", cex = 1.2)
legend("bottomleft", "Mean to PC2 hypothetical minimum", bty = "n", inset = c(0.07, 0.07))

# 10) Mean to Big, residuals PC2, lateral view
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resP2.big.lat)[2]), c(0, dim(resP2.big.lat)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(resP2.big.lat, 0, 0, dim(resP2.big.lat)[2], dim(resP2.big.lat)[1])
text(0, dim(resP2.big.lat)[1]-10, "J", cex = 2)
legend("bottomleft", "Mean to PC2 hypothetical maximum", bty = "n", inset = c(0.07, 0.07))

# 11) Mean to Small, residuals PC2, dorsal view
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resP2.big.dor)[2]), c(0, dim(resP2.big.dor)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(resP2.sm.dor, 0, 0, dim(resP2.big.dor)[2], dim(resP2.big.dor)[1])
text(0, dim(resP2.big.dor)[1]-10, "K", cex = 2)
legend("bottomleft", "Dorsal view of I", bty = "n", inset = c(0.2, 0.05))

# 12) Mean to Big, residuals PC2, dorsal view
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resP2.big.dor)[2]), c(0, dim(resP2.big.dor)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(resP2.big.dor, 0, 0, dim(resP2.big.dor)[2], dim(resP2.big.dor)[1])
text(0, dim(resP2.big.dor)[1]-10, "L", cex = 2)
legend("bottomleft", "Dorsal view of J", bty = "n", inset = c(0.2, 0.05))

dev.off()
```