---
title: "App05-supplementary-heatmaps"
author: "Ariel Marcy"
date: "3/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '../eco-rodents')
```

# Supplementary heatmaps residual shape without Notomys and carnivores
1) Make heatmaps for residuals without Notomys PC2.

2) Make heatmaps for residuals without Notomys and without carnivores PC1 & PC2

Produces multipanel Supplementary Figure 1.

### Load packages, functions, and data
```{r message = FALSE}
library(geomorph)
library(data.table)
library(vegan)
library(geiger)  # for phylogenies
library(png)  # needed to work with screenshots
library(colorspace)  # plot colors
# Load libraries (and install if necessary) for landvR package
if(!require(devtools)) install.packages("devtools")
if(!require(dispRity)) install.packages("dispRity")
if(!require(landvR)) install_github("TGuillerme/landvR")
source("../Functions/utilities.R")  # custom functions
source("../Functions/plotGMPhyloMorphoSpace_plotmod.R")  # modified geomorph function
load(file = "../Data/Processed/03-main-data.rda")
load(file = "../Data/Processed/03-tree-data.rda")
load(file = "../Data/Processed/03-graphing-vectors.rda")
load(file = "../Data/Processed/04-PCA.rda")
load(file = "../Data/Processed/05-residual-data.rda")
```

## 1) Prepare heatmaps
This script is dominated by the code needed to generate, export, and clean the heatmap images which are tiled in Supplementary Figure 1.

### Remove Notomys and find hypothetical shapes for PC2
```{r}
# Remove Notomys from dataset
index.not <- which(info.means$Genus == "Not")
residuals.no <- residuals[, , -index.not]
info.means.no <- info.means[-index.not, ]

# Find new mean
res.mean.no <- (select.procrustes(residuals.no, selector = mean))[[1]]

# Measure spherical coordinates differences from the mean for each specimen for each landmark
res.diff.mean.no <- coordinates.difference(coordinates = residuals.no, reference = res.mean.no, type = "spherical")

# Ordinate the data
twoD.shape.r.n <- two.d.array(residuals.no)
ordination.r.n <- stats::prcomp(twoD.shape.r.n)

# Force the symmetric component of shape into class "gpagen" (required for variation.range)
gpagen.shape.no <- list()
gpagen.shape.no$coords <- residuals.no
gpagen.shape.no$consensus <- res.mean.no
class(gpagen.shape.no) <- "gpagen"

# Measure extremes of variation from mean on PC1
PC1.var.r.n2 <- variation.range(gpagen.shape.no, return.ID = FALSE, axis = 2, ordination = ordination.r.n, type = "spherical")

# Wrap specimens on the tangent space
wrap.PCA.r.n2 <- plotTangentSpace(residuals.no, verbose = FALSE)

# Select extreme specimens (the ones that make the warp-meshes)
res.PC1.min.h.n2 <- wrap.PCA.r.n2$pc.shapes[[3]] 
res.PC1.max.h.n2 <- wrap.PCA.r.n2$pc.shapes[[4]]
```

### Export residual no Notomys heatmaps
```{r}
# Lateral view of mean to smallest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean.no, res.PC1.min.h.n2, col = heat.colors, col.val = PC1.var.r.n2[, "radius"])
view3d(phi = 90)
rgl.snapshot("../Data/Processed/Heatmaps/NO2_mean-sm_lat.png")

# Lateral view of mean to biggest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean.no, res.PC1.max.h.n2, col = heat.colors, col.val = PC1.var.r.n2[, "radius"])
view3d(phi = 90)
rgl.snapshot("../Data/Processed/Heatmaps/NO2_mean-bg_lat.png")

# Dorsal view of mean to smallest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean.no, res.PC1.min.h.n2, col = heat.colors, col.val = PC1.var.r.n2[, "radius"]) 
view3d(phi = 0)  # flip cranium
rgl.snapshot("../Data/Processed/Heatmaps/NO2_mean-sm_dor.png")

# Dorsal view of mean to biggest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean.no, res.PC1.max.h.n2, col = heat.colors, col.val = PC1.var.r.n2[, "radius"])
view3d(phi = 0)  # flip cranium
rgl.snapshot("../Data/Processed/Heatmaps/NO2_mean-bg_dor.png")
```

### Remove Notomys and carnivores, then find hypothetical shapes for PC1
```{r}
# Remove Notomys and carnivores from dataset
index.NC <- c(which(info.means$Genus == "Not"), which(info.means$Genus == "Hyd", which(info.means$Genus == "Xer")))
residuals.NC <- residuals[, , -index.NC]
info.means.no <- info.means[-index.NC, ]

# Find new mean
res.mean.NC <- (select.procrustes(residuals.NC, selector = mean))[[1]]

# Measure spherical coordinates differences from the mean for each specimen for each landmark
res.diff.mean.NC <- coordinates.difference(coordinates = residuals.NC, reference = res.mean.NC, type = "spherical")

# Ordinate the data
twoD.shape.r.NC <- two.d.array(residuals.NC)
ordination.r.NC <- stats::prcomp(twoD.shape.r.NC)

# Force the symmetric component of shape into class "gpagen" (required for variation.range)
gpagen.shape.NC <- list()
gpagen.shape.NC$coords <- residuals.NC
gpagen.shape.NC$consensus <- res.mean.NC
class(gpagen.shape.NC) <- "gpagen"

# Measure extremes of variation from mean on PC1
PC1.var.NC1 <- variation.range(gpagen.shape.NC, return.ID = FALSE, axis = 1, ordination = ordination.r.NC, type = "spherical")

# Wrap specimens on the tangent space
wrap.PCA.r.NC <- plotTangentSpace(residuals.NC, verbose = FALSE)

# Select extreme specimens (the ones that make the warp-meshes)
res.PC1.min.NC <- wrap.PCA.r.NC$pc.shapes[[1]] 
res.PC1.max.NC <- wrap.PCA.r.NC$pc.shapes[[2]]
```

### Export residual PC1 no Notomys nor carnivores heatmaps
```{r}
# Lateral view of mean to smallest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean.NC, res.PC1.min.NC, col = heat.colors, col.val = PC1.var.NC1[, "radius"])
view3d(phi = 90)
rgl.snapshot("../Data/Processed/Heatmaps/NC1_mean-sm_lat.png")

# Lateral view of mean to biggest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean.NC, res.PC1.max.NC, col = heat.colors, col.val = PC1.var.NC1[, "radius"])
view3d(phi = 90)
rgl.snapshot("../Data/Processed/Heatmaps/NC1_mean-bg_lat.png")

# Dorsal view of mean to smallest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean.NC, res.PC1.min.NC, col = heat.colors, col.val = PC1.var.NC1[, "radius"]) 
view3d(phi = 0)  # flip cranium
rgl.snapshot("../Data/Processed/Heatmaps/NC1_mean-sm_dor.png")

# Dorsal view of mean to biggest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean.NC, res.PC1.max.NC, col = heat.colors, col.val = PC1.var.NC1[, "radius"])
view3d(phi = 0)  # flip cranium
rgl.snapshot("../Data/Processed/Heatmaps/NC1_mean-bg_dor.png")
```

### Make heatmaps for PC2 residual coordinates (no notomys or carnivores)
```{r}
# Measure extremes of variation from mean on PC1
PC1.var.r.NC2 <- variation.range(gpagen.shape.NC, return.ID = FALSE, axis = 2, ordination = ordination.r.NC, type = "spherical")

# Select extreme specimens (the ones that make the warp-meshes)
res.PC2.min.NC2 <- wrap.PCA.r.NC$pc.shapes[[4]]
res.PC2.max.NC2 <- wrap.PCA.r.NC$pc.shapes[[3]]
```

### Export residual PC1 no Notomys nor carnivores heatmaps
```{r}
# Lateral view of mean to smallest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean.NC, res.PC2.min.NC2, col = heat.colors, col.val = PC1.var.r.NC2[, "radius"])
view3d(phi = 90)
rgl.snapshot("../Data/Processed/Heatmaps/NC2_mean-sm_lat.png")

# Lateral view of mean to biggest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean.NC, res.PC2.max.NC2, col = heat.colors, col.val = PC1.var.r.NC2[, "radius"])
view3d(phi = 90)
rgl.snapshot("../Data/Processed/Heatmaps/NC2_mean-bg_lat.png")

# Dorsal view of mean to smallest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean.NC, res.PC2.min.NC2, col = heat.colors, col.val = PC1.var.r.NC2[, "radius"]) 
view3d(phi = 0)  # flip cranium
rgl.snapshot("../Data/Processed/Heatmaps/NC2_mean-sm_dor.png")

# Dorsal view of mean to biggest
open3d(windowRect = c(0, 0, 600, 600))
procrustes.var.plot(res.mean.NC, res.PC2.max.NC2, col = heat.colors, col.val = PC1.var.r.NC2[, "radius"])
view3d(phi = 0)  # flip cranium
rgl.snapshot("../Data/Processed/Heatmaps/NC2_mean-bg_dor.png")
```

### Load in all heatmaps
```{r}
# Remove whitespace from all heatmaps (requires ImageMagik installed on PC)
system("mogrify ../Data/Processed/Heatmaps/*.png -trim ../Data/Processed/Heatmaps/*.png")

# Load in without allometry residuals PC2 - no Notomys
resid.sm.lat.n2 <- readPNG("../Data/Processed/Heatmaps/NO2_mean-sm_lat.png")
resid.big.lat.n2 <- readPNG("../Data/Processed/Heatmaps/NO2_mean-bg_lat.png")
resid.sm.dor.n2 <- readPNG("../Data/Processed/Heatmaps/NO2_mean-sm_dor.png")
resid.big.dor.n2 <- readPNG("../Data/Processed/Heatmaps/NO2_mean-bg_dor.png")

# Load in without notomys or carnivores residuals PC1
resNC1.sm.lat <- readPNG("../Data/Processed/Heatmaps/NC1_mean-sm_lat.png")
resNC1.big.lat <- readPNG("../Data/Processed/Heatmaps/NC1_mean-bg_lat.png")
resNC1.sm.dor <- readPNG("../Data/Processed/Heatmaps/NC1_mean-sm_dor.png")
resNC1.big.dor <- readPNG("../Data/Processed/Heatmaps/NC1_mean-bg_dor.png")

# Load in without notomys or carnivores residuals PC2
resNC2.sm.lat <- readPNG("../Data/Processed/Heatmaps/NC2_mean-sm_lat.png")
resNC2.big.lat <- readPNG("../Data/Processed/Heatmaps/NC2_mean-bg_lat.png")
resNC2.sm.dor <- readPNG("../Data/Processed/Heatmaps/NC2_mean-sm_dor.png")
resNC2.big.dor <- readPNG("../Data/Processed/Heatmaps/NC2_mean-bg_dor.png")
```

## 2) Make multi-panel Supplementary Figure 1: Heatmaps with Residuals
```{r}
# Set up muli-panel figure
pdf("../Data/Results/SFigure1_NOt_Heatmap_Residuals.pdf")
mat <- matrix(c(1:12), 4, 3, byrow = FALSE)  # 12 plots, 4 rows, 3 columns
layout(mat, widths = rep(1, dim(mat)[2]), heights = rep(0.5, dim(mat)[1])) 

## PC1 Residuals -- without Notomys
# 1) Mean to Small, residuals PC2, lateral view - without Notomys
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resid.sm.lat.n2)[2]), c(0, dim(resid.sm.lat.n2)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(resid.sm.lat.n2, 0, 0, dim(resid.sm.lat.n2)[2], dim(resid.sm.lat.n2)[1])
text(0, dim(resid.sm.lat.n2)[1]-10, "A", cex = 2)
legend("topright", inset = c(0.17, 0), legend = expression(paste("PC2 Residuals (no ", italic("Notomys"), ")")), bty = "n", cex = 1.1)
legend("bottomleft", "Mean to hypothetical minimum", bty = "n", inset = c(0.1, 0.07))

# 2) Mean to Big, residuals PC2, lateral view - without Notomys
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resid.big.lat.n2)[2]), c(0, dim(resid.big.lat.n2)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(resid.big.lat.n2, 0, 0, dim(resid.big.lat.n2)[2], dim(resid.big.lat.n2)[1])
text(0, dim(resid.big.lat.n2)[1]-10, "B", cex = 2)
legend("bottomleft", "Mean to hypothetical maximum", bty = "n", inset = c(0.1, 0.07))

# 3) Mean to Small, residuals PC2, dorsal view - without Notomys
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resid.big.dor.n2)[2]), c(0, dim(resid.big.dor.n2)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(resid.sm.dor.n2, 0, 0, dim(resid.big.dor.n2)[2], dim(resid.big.dor.n2)[1])
text(0, dim(resid.big.dor.n2)[1]-10, "C", cex = 2)
legend("bottomleft", "Dorsal view of A", bty = "n", inset = c(0.2, 0.05))

# 4) Mean to Big, residuals PC2, dorsal view - without Notomys
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resid.big.dor.n2)[2]), c(0, dim(resid.big.dor.n2)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(resid.big.dor.n2, 0, 0, dim(resid.big.dor.n2)[2], dim(resid.big.dor.n2)[1])
text(0, dim(resid.big.dor.n2)[1]-10, "D", cex = 2)
legend("bottomleft", "Dorsal view of B", bty = "n", inset = c(0.2, 0.05))

## PC 1 Residuals -- without Notomys or carnivores
# 5) Mean to Small, residuals PC1, lateral view - no Notomys or carnivores
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resNC1.sm.lat)[2]), c(0, dim(resNC1.sm.lat)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(resNC1.sm.lat, 0, 0, dim(resNC1.sm.lat)[2], dim(resNC1.sm.lat)[1])
text(0, dim(resNC1.sm.lat)[1]-10, "E", cex = 2)
legend("topright", inset = c(0, 0), legend = expression(paste("PC1 Residuals (no ", italic("Notomys"), " or carnivores)")), bty = "n", cex = 1.05)
legend("bottomleft", "Mean to new hypothetical minimum", bty = "n", inset = c(0.07, 0.07))

# 6) Mean to Big, residuals PC1, lateral view - no Notomys or carnivores
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resNC1.big.lat)[2]), c(0, dim(resNC1.big.lat)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(resNC1.big.lat, 0, 0, dim(resNC1.big.lat)[2], dim(resNC1.big.lat)[1])
text(0, dim(resNC1.big.lat)[1]-10, "F", cex = 2)
legend("bottomleft", "Mean to new hypothetical maximum", bty = "n", inset = c(0.07, 0.07))

# 7) Mean to Small, residuals PC1, dorsal view - no Notomys or carnivores
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resNC1.sm.dor)[2]), c(0, dim(resNC1.sm.dor)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(resNC1.sm.dor, 0, 0, dim(resNC1.sm.dor)[2], dim(resNC1.sm.dor)[1])
text(0, dim(resNC1.sm.dor)[1]-10, "G", cex = 2)
legend("bottomleft", "Dorsal view of E", bty = "n", inset = c(0.2, 0.05))

# 8) Mean to Big, residuals PC1, dorsal view - no Notomys or carnivores
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resNC1.big.dor)[2]), c(0, dim(resNC1.big.dor)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(resNC1.big.dor, 0, 0, dim(resNC1.big.dor)[2], dim(resNC1.big.dor)[1])
text(0, dim(resNC1.big.dor)[1]-10, "H", cex = 2)
legend("bottomleft", "Dorsal view of F", bty = "n", inset = c(0.2, 0.05))

## PC2 Residuals -- no Notomys or carnivores
# 9) Mean to Small, residuals PC2, lateral view - no Notomys or carnivores
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resNC2.sm.lat)[2]), c(0, dim(resNC2.sm.lat)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(resNC2.sm.lat, 0, 0, dim(resNC2.sm.lat)[2], dim(resNC2.sm.lat)[1])
text(0, dim(resNC2.sm.lat)[1]-10, "I", cex = 2)
legend("topright", inset = c(0, 0), legend = expression(paste("PC2 Residuals (no ", italic("Notomys"), " or carnivores)")), bty = "n", cex = 1.05)
legend("bottomleft", "Mean to PC2 hypothetical minimum", bty = "n", inset = c(0.07, 0.07))

# 10) Mean to Big, residuals PC2, lateral view - no Notomys or carnivores
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resNC2.big.lat)[2]), c(0, dim(resNC2.big.lat)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(resNC2.big.lat, 0, 0, dim(resNC2.big.lat)[2], dim(resNC2.big.lat)[1])
text(0, dim(resNC2.big.lat)[1]-10, "J", cex = 2)
legend("bottomleft", "Mean to PC2 hypothetical maximum", bty = "n", inset = c(0.07, 0.07))

# 11) Mean to Small, residuals PC2, dorsal view - no Notomys or carnivores
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resNC2.sm.dor)[2]), c(0, dim(resNC2.sm.dor)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(resNC2.sm.dor, 0, 0, dim(resNC2.sm.dor)[2], dim(resNC2.sm.dor)[1])
text(0, dim(resNC2.sm.dor)[1]-10, "K", cex = 2)
legend("bottomleft", "Dorsal view of I", bty = "n", inset = c(0.2, 0.05))

# 12) Mean to Big, residuals PC2, dorsal view - no Notomys or carnivores
par(mar = c(0, 0, 0, 0))
plot(c(0, dim(resNC2.big.dor)[2]), c(0, dim(resNC2.big.dor)[1]), type = "n", axes = FALSE, xlab = "", ylab = "", asp = TRUE) 
rasterImage(resNC2.big.dor, 0, 0, dim(resNC2.big.dor)[2], dim(resNC2.big.dor)[1])
text(0, dim(resNC2.big.dor)[1]-10, "L", cex = 2)
legend("bottomleft", "Dorsal view of J", bty = "n", inset = c(0.2, 0.05))

# Make gradient legend
legend_image <- as.raster(matrix(heat.colors(20), ncol = 1))
rasterImage(legend_image, 443, 150, 465, 215)
text(365, 208, "High variation", col = "dark grey", cex = 1)
text(413, 162, "Low", col = "dark grey", cex = 1)

dev.off()
```