---
title: "App01-compare-sliding-methods"
author: "Ariel Marcy"
date: "10/25/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '../aus-rodent-skulls')
```

# Appendix 1: Checking methodology of sliding in Viewbox and _geomorph_
While landmarking, all points were slid in Viewbox according to our template and their proprietary software. Next, we slid the landmarks again using _geomorph_'s functions (script 02). 

Here we check whether sliding twice impacted the landmark variation patterns by looking at the variation on a dataset that was not slid in _geomorph_ - i.e. all points were treated as fixed landmarks. 

### Load packages and functions
```{r message = FALSE}
# Load libraries (and install if necessary) for landvR package
if(!require(devtools)) install.packages("devtools")
if(!require(dispRity)) install.packages("dispRity")
if(!require(landvR)) install_github("TGuillerme/landvR")

library(stringr)
library(data.table)
library(geomorph)
source("../Functions/utilities.R")  # custom functions
load(file = "../Data/Processed/01-data-for-appendix.rda")  # pre-gpagen() shape data from script 01, to be re-slid in this script
load(file = "../Data/Processed/02-data-for-analyses.rda")  # main data set
```

## Run GPA treating all points as fixed landmarks
Here, we conserve the ratios set up in Viewbox. 
```{r}
sm.Y.unslid <- gpagen(S, Proj = TRUE, ProcD = TRUE)
```

## Run bilateral symmetry on unslid dataset
```{r}
sm.land.pairs <- FindPairs(sm.pt.names)
sm.biY.unslid <- bilat.symmetry(sm.Y.unslid$coords, ind = sm.sp.info$All, object.sym = T, replicate = NULL, side = NULL, land.pairs = sm.land.pairs)
sm.bY.unslid.ss <- sm.biY.unslid$symm.shape
cols2 <- c("Genus", "Species", "CatNum", "Rep")
sm.bY.unslid.info <- WriteMetadata(sm.bY.unslid.ss, cols2)
```

### Remove specimens added for user error measurement
First, remove the replicates from the main shape data for both small and large patch protocols. Create corresponding metadata tables. These data will be used for the majority of remaining analyses.
```{r}
main <- which(sm.bY.unslid.info$Rep == "non")
unslid.shape <- sm.bY.unslid.ss[, , main]
unslid.info <- sm.bY.unslid.info[main, ]
```

## Landmark variation heatmaps
We can use the `procrustes.var.plot()` function by Dr Thomas Guillerme to see how landmarks vary in the unslid dataset compared to the slid.
```{r}
# Find mean configuration
unslid.consensus <- (select.procrustes(unslid.shape, selector = mean))[[1]]

# Measure spherical coordinates differences from the mean
diff.from.mean <- coordinates.difference(coordinates = unslid.shape, reference = unslid.consensus, type = "spherical")
```

To test the hypothesis that each landmark's variation from the mean configuration is above expected, we can use the PC axes from a PCA.
```{r}
# Ordinate the data
unslid.twoD.shape <- two.d.array(unslid.shape)
unslid.ordination <- stats::prcomp(unslid.twoD.shape)

# Force the symmetric component of shape into class "gpagen" (required for variation.range)
unslid.gpagen.shape <- list()
unslid.gpagen.shape$coords <- unslid.shape
unslid.gpagen.shape$consensus <- unslid.consensus
class(unslid.gpagen.shape) <- "gpagen"

# Measure extremes of variation from mean on PC1
unslid.PC1.var <- variation.range(unslid.gpagen.shape, return.ID = FALSE, axis = 1, ordination = unslid.ordination, type = "spherical")
```

The next step is to make a heatmap showing the size of variation and direction of movement for each landmark in the dataset, comparing between the min and max specimens along PC1.
```{r}
# Wrap specimens on the tangent space
unslid.wrap.PCA <- plotTangentSpace(unslid.shape, verbose = FALSE)

# Select extreme specimens (the ones that make the warp-meshes)
un.hypothetical.1 <- unslid.wrap.PCA$pc.shapes[[1]]
un.hypothetical.2 <- unslid.wrap.PCA$pc.shapes[[2]]

# Plot the range of variation along PC1 using a heat color scheme
PC1.var.plot <- procrustes.var.plot(un.hypothetical.1, un.hypothetical.2, col = heat.colors, col.val = unslid.PC1.var[, "radius"], labels = F)
```

### Histogram of radius variation for unslid dataset
```{r}
# Plot the histogram of variation in landmarks for this protocol
var.range.unslid <- variation.range(unslid.gpagen.shape)
hist(var.range.unslid[, 1])
```

### Compare to slid 
```{r}
# Plot the histogram of variation in landmarks for this protocol
var.range <- variation.range(gpagen.shape)
hist(var.range[, 1])
```